---
title: "07_2_HDCA_heart_SCENIC_EN_18_and_regulon_enrichment"
author: "Raphaël Mauron"
output: html_document
editor_options: 
  chunk_output_type: console
---

# HDCA heart SCENIC EN18

Calculate Gene Regulatory Network and infer Transcription Factors activity (Regulons) with SCENIC.

Regulons enrichment in EN_7, EN_17, EN_18 in time.

NOTE: The path to load the data and the path to save the figures should be changed at your convenience.

Environment: **Docker**


## Settings

### Set library path

```{r setup, include=FALSE}
.libPaths("/home/rstudio/project/renv/library/R-4.3/aarch64-unknown-linux-gnu/")
knitr::opts_chunk$set(echo = TRUE)
```


### Load Libraries

```{r, message = FALSE}
library(Seurat)
library(SCENIC)
library(SingleCellExperiment)
library(ggplot2)
library(tidygraph)
library(ggraph)
library(igraph)
library(feather)
library(data.table)
```


### Set environment variables

```{r}
myfolder <- "/home/rstudio/project/output/HDCA_heart_sc_analysis_docker/"
```


### Load data

```{r}
EN <- readRDS(file = paste0(myfolder, "rds_objects/endothelial.rds"))
```


### Set seed

```{r}
set.seed(1)
```


## Sub-sample to specific cluster of interest

Prepare data_enrich for enrichment done later on.

```{r}
data_cell <- EN
cluster <- "18"
data_sub <- data_cell[,data_cell$clusters_subset %in% cluster]

to_keep_for_enrich <- c("7", "17", "18")
data_enrich <- data_cell[,data_cell$clusters_subset %in% to_keep_for_enrich]
```


### Cells per cluster per age group

```{r}
table(data_enrich$age_groups, data_enrich$clusters_subset)
```


### Add additional age groups to the metadata (age_group_refined)

```{r}
df <- data.frame(age = c(5.5, 6, 7, 8, 8.5, 9, 10, 11.5, 12, 13.25, 14), age_group = c("5-6", "5-6", "7-8", "7-8", "7-8", "9-14", "9-14", "9-14", "9-14", "9-14", "9-14"))
age_to_age_group <- setNames(df$age_group, nm = df$age)
data_enrich$age_group_refined <- factor(age_to_age_group[as.character(data_enrich$age)], levels = c("5-6", "7-8", "9-14"))
rm(df, age_to_age_group)
gc()
```


### Cells per cluster per new age group

```{r}
table(data_enrich$age_group_refined, data_enrich$clusters_subset)
```


### Get expression matrix

```{r}
exprMat <- GetAssayData(data_sub, assay = "RNA", slot = "counts")
exprMat <- as.matrix(exprMat)

rownames(exprMat)
dim(exprMat)

Idents(data_sub)
cellInfo <- data.frame(seuratCluster=Idents(data_sub))
```


### Initialize settings

database selected:
´´´wget https://resources.aertslab.org/cistarget/databases/old/homo_sapiens/hg19/refseq_r45/mc9nr/gene_based/hg19-500bp-upstream-7species.mc9nr.feather´´´

´´´wget https://resources.aertslab.org/cistarget/databases/old/homo_sapiens/hg19/refseq_r45/mc9nr/gene_based/hg19-tss-centered-10kb-7species.mc9nr.feather´´´


```{r}
org <- "hgnc"
dbDir <- "/home/rstudio/project/data/db_scenic" # RcisTarget databases location
dbs <- c("hg19-500bp-upstream-10species.mc9nr.feather",
         "hg19-tss-centered-10kb-7species.mc9nr.feather")
scenicOptions <- initializeScenic(org=org, dbDir=dbDir, dbs=dbs, nCores=10) #might give an error so need to rename motifAnnotations

motifAnnotations_hgnc <- motifAnnotations

scenicOptions <- initializeScenic(org=org, dbDir=dbDir, dbs=dbs, nCores=10) #rerun 
rm(motifAnnotations)

# Add cellInfo to scenicOptions object
scenicOptions@inputDatasetInfo$cellInfo <- cellInfo
```


### Co-expression network

```{r}
genesKept <- geneFiltering(exprMat, scenicOptions) #create list of markers from gene set that are found in databases provided

#filter expression matrix
exprMat_filtered <- exprMat[genesKept, ] #filter based on genes found in database
dim(exprMat_filtered)
exprMat_filtered <- exprMat[VariableFeatures(EN),] #keeps only the highly variable genes
dim(exprMat_filtered)

#run correlation
runCorrelation(exprMat_filtered, scenicOptions)

#run GENIE3
exprMat_filtered_log <- log2(exprMat_filtered+1) #check if already normalized in the exprMat_filtered
runGenie3(exprMat_filtered_log, scenicOptions)
```


#### Extract TF List from GENIE3

```{r}
genie_list <- loadInt(scenicOptions, "genie3ll")
table(genie_list$TF)
list_unique_TF <- c(sort(unique(genie_list$TF)))
list_unique_TF <- factor(list_unique_TF, levels = list_unique_TF)

# Save unique list_unique_TF
write.csv(list_unique_TF, file = paste0(myfolder, "scenic/markers_table/", "list_unique_TF","_EN_", cluster, ".tsv"))
```


#### Enrichment of unique TF in remaining clusters

```{r}
data_cell <- SetIdent(data_cell, value = "clusters_subset")
data_cell$clusters_subset <- factor(data_cell$clusters_subset, 
                                              levels = c("7","17", "18")) #To find most variable TF between THOSE clusters

# Identify all other clusters
all_clusters <- levels(data_cell$clusters_subset)
other_clusters <- setdiff(all_clusters, cluster)

# Run FindMarkers to compare the cluster of interest against all other clusters combined
markers_table <- FindMarkers(data_cell, 
                             ident.1 = cluster, 
                             ident.2 = other_clusters, 
                             test.use = "wilcox", 
                             features = list_unique_TF)

# Print the resulting table
markers_table$pct.diff <- markers_table$pct.1 - markers_table$pct.2

# Save markers_table
#write.csv(markers_table, file = paste0(myfolder, "scenic/markers_table/", "EN_", cluster, ".tsv"))

markers_table_to_keep <- markers_table %>%
  arrange(p_val) %>%
  filter(avg_log2FC > 0) %>%
  filter(p_val < 5e-7)
```


### Build and score the GRN

#### SCENIC_1

```{r}
scenicOptions <- runSCENIC_1_coexNetwork2modules(scenicOptions)
tfModules_asDF <- loadInt(scenicOptions, "tfModules_asDF")
```


#### SCENIC_2

```{r}
# Remove genes missing from RcisTarget databases
tfModules_asDF$TF <- as.character(tfModules_asDF$TF)
tfModules_asDF$Target <- as.character(tfModules_asDF$Target)
allTFs <- getDbTfs(scenicOptions)
tfModules_asDF <- tfModules_asDF[which(tfModules_asDF$TF %in% allTFs),]


genesInDb <- unique(unlist(lapply(getDatabases(scenicOptions), function(x)
  names(feather::feather_metadata(x)[["types"]]))))


geneInDb <- tfModules_asDF$Target %in% genesInDb
missingGene <- sort(unique(tfModules_asDF[which(!geneInDb),"Target"]))
if(length(missingGene)>0) 
  warning(paste0("Genes in co-expression modules not available in RcisTargetDatabases: ", 
                 paste(missingGene, collapse=", ")))
tfModules_asDF <- tfModules_asDF[which(geneInDb),]


# Targets with positive correlation
tfModules_Selected <- tfModules_asDF[which(tfModules_asDF$corr==1),]


# Add a column with the geneSet name (TF_method)
tfModules_Selected <- cbind(tfModules_Selected, geneSetName=paste(tfModules_Selected$TF, tfModules_Selected$method, sep="_"))
tfModules_Selected$geneSetName <- factor(as.character(tfModules_Selected$geneSetName))
allGenes <- unique(tfModules_Selected$Target)


# Split into tfModules (TF-modules, with several methods)
tfModules <- split(tfModules_Selected$Target, tfModules_Selected$geneSetName)


# Add TF to the gene set (used in the following steps, careful if editing)
tfModules <- setNames(lapply(names(tfModules), function(gsn) {
  tf <- strsplit(gsn, "_")[[1]][1]
  unique(c(tf, tfModules[[gsn]]))
}), names(tfModules))


# Keep gene sets with at least 'minGenes' genes
minGenes <- 20
tfModules <- tfModules[which(lengths(tfModules)>=minGenes)]


saveRDS(tfModules, file=getIntName(scenicOptions, "tfModules_forEnrichment"))
tfModules_forEnrichment <- loadInt(scenicOptions, "tfModules_forEnrichment")
```

```{r}
#RcisTarget
# 1.1 Calculate enrichment
motifs_AUC <- lapply(getDatabases(scenicOptions), function(rnkName) {
  ranking <- importRankings(rnkName, columns=allGenes)
  message("Scoring database: ", ranking@description)
  RcisTarget::calcAUC(tfModules, ranking, aucMaxRank=0.03*getNumColsInDB(ranking), nCores=9, verbose=TRUE)})

saveRDS(motifs_AUC, file=getIntName(scenicOptions, "motifs_AUC"))
```

```{r}
#1.2 Convert to table, filter by NES & add the TFs to which the motif is annotated
motifAnnot <- motifAnnotations_hgnc
motifEnrichment <- lapply(motifs_AUC, function(aucOutput)
{
  # Extract the TF of the gene-set name (i.e. MITF_w001):
  tf <- sapply(setNames(strsplit(rownames(aucOutput), "_"), rownames(aucOutput)), function(x) x[[1]])
  
  # Calculate NES and add motif annotation (provide tf in 'highlightTFs'):
  addMotifAnnotation(aucOutput, 
                     nesThreshold=3, digits=3, 
                     motifAnnot=motifAnnot,
                     motifAnnot_highConfCat=c("directAnnotation", "inferredBy_Orthology"),
                     motifAnnot_lowConfCat=c("inferredBy_MotifSimilarity",
                                               "inferredBy_MotifSimilarity_n_Orthology"), 
                     highlightTFs=tf)
})

# Merge both tables, adding a column that contains the 'motifDb'
motifEnrichment <- do.call(rbind, lapply(names(motifEnrichment), function(dbName){
  cbind(motifDb=dbName, motifEnrichment[[dbName]])
}))
saveRDS(motifEnrichment, file=getIntName(scenicOptions, "motifEnrichment_full"))
```

```{r}
#1.3 Keep only the motifs annotated to the initial TF
getIntName(scenicOptions)
motifEnrichment <- loadInt(scenicOptions, "motifEnrichment_full")

#instead, drop the once not in TF_highConf
motifEnrichment_selfMotifs <- motifEnrichment[which(motifEnrichment$TF_highConf != ""),, drop=FALSE]
motifEnrichment_selfMotifs <- motifEnrichment_selfMotifs[which(motifEnrichment_selfMotifs$AUC > 0.1),, drop=FALSE]
motifEnrichment_selfMotifs <- motifEnrichment_selfMotifs[which(motifEnrichment_selfMotifs$NES > 3.9),, drop=FALSE]
```


```{r}
#2. Prune targets
dbNames <- getDatabases(scenicOptions)
nCores <- 9

motifEnrichment_selfMotifs_wGenes <- lapply(names(dbNames), function(motifDbName){
  ranking <- importRankings(dbNames[motifDbName], columns=allGenes)
  addSignificantGenes(resultsTable=motifEnrichment_selfMotifs[motifEnrichment_selfMotifs$motifDb==motifDbName,],
                      geneSets=tfModules,
                      rankings=ranking,
                      maxRank=5000, method="aprox", nCores=nCores)
})

motifEnrichment_selfMotifs_wGenes <- rbindlist(motifEnrichment_selfMotifs_wGenes)
motifEnrichment_selfMotifs_wGenes[order(motifEnrichment_selfMotifs_wGenes$NES,decreasing=TRUE),][1:5,(1:ncol(motifEnrichment_selfMotifs_wGenes)-1), with=F]


motifEnrichment.asIncidList <- apply(motifEnrichment_selfMotifs_wGenes, 1, function(oneMotifRow) {
  genes <- strsplit(oneMotifRow["enrichedGenes"], ";")[[1]]
  oneMotifRow <- data.frame(rbind(oneMotifRow), stringsAsFactors=FALSE)
  data.frame(oneMotifRow[rep(1, length(genes)),c("NES", "motif", "highlightedTFs", "TFinDB")], genes, stringsAsFactors = FALSE)
})
motifEnrichment.asIncidList <- rbindlist(motifEnrichment.asIncidList)
colnames(motifEnrichment.asIncidList) <- c("NES", "motif", "TF", "annot", "gene")
motifEnrichment.asIncidList <- data.frame(motifEnrichment.asIncidList, stringsAsFactors = FALSE)


# Get targets for each TF, but keep info about best motif/enrichment
regulonTargetsInfo <- lapply(split(motifEnrichment.asIncidList, motifEnrichment.asIncidList$TF), function(tfTargets){
  # print(unique(tfTargets$TF))
  tfTable <- as.data.frame(do.call(rbind, lapply(split(tfTargets, tfTargets$gene), function(enrOneGene){
    highConfAnnot <- "**" %in% enrOneGene$annot
    enrOneGeneByAnnot <- enrOneGene
    if(highConfAnnot) enrOneGeneByAnnot <- enrOneGeneByAnnot[which(enrOneGene$annot == "**"),]
    bestMotif <- which.max(enrOneGeneByAnnot$NES)

    cbind(TF=unique(enrOneGene$TF), gene=unique(enrOneGene$gene), nMotifs=nrow(enrOneGene),
          bestMotif=as.character(enrOneGeneByAnnot[bestMotif,"motif"]), NES=as.numeric(enrOneGeneByAnnot[bestMotif,"NES"]),
          highConfAnnot=highConfAnnot)
  })), stringsAsFactors=FALSE)
  tfTable[order(tfTable$NES, decreasing = TRUE),]
})
rm(motifEnrichment.asIncidList)
regulonTargetsInfo <- rbindlist(regulonTargetsInfo)
colnames(regulonTargetsInfo) <- c("TF", "gene", "nMotifs", "bestMotif", "NES", "highConfAnnot")
```


#### Create modules based on the regulonTargetsInfo

```{r}
# Split the dataframe by TF
TF_modules <- split(regulonTargetsInfo, regulonTargetsInfo$TF)

# Extract target genes for each TF
TF_modules_list <- lapply(TF_modules, function(sub_df) sub_df$gene)
```


#### Export TF_modules_list

```{r}
regulonTargetsInfo_export <- regulonTargetsInfo[, c("TF", "gene")]
write.csv(regulonTargetsInfo_export, file = paste0(myfolder, "scenic/EN_enrichment/EN_modules.csv"), row.names = FALSE)
```


#### Use TF_modules_list

```{r}
my_TF <- "NFKB1" #"KLF4"
my_targets <- TF_modules_list[my_TF]
my_targets
```


#### Enrichment of Targets in clusters

```{r}
# Compute module score in clusters
data_sub_en <- data_enrich
data_sub_en <- AddModuleScore(data_sub_en, features = my_targets, name = "Enrichment_Score")

pdf(file = paste0(myfolder, "scenic/EN_enrichment/",my_TF, ".pdf"), width = 15, height = 10)
VlnPlot(data_sub_en, features = "Enrichment_Score1", group.by = "clusters_subset", split.by = "age_group_refined", pt.size = 0) +
  ggtitle(paste0("Enrichment of ", my_TF, " target genes"))
dev.off()
```


#### Enrichment of multiple TF

```{r}
TF_list <- c("NFKB1", "KLF4", "MECOM", "HAND2", "ZNF385D", "SALL3", "PROX1", "SOX6", "PRDM16", "FOXC2", "HMGA2", "MEIS2", "SMAD6", "HEY2", "LMO2", "NFATC1", "SP6", "ARID5A", "ARID5B", "RFX2", "GLIS3", "LEF1", "FOS", "MSX1", "SOX6", "FOXP2", "ESRRG", "MAF", "FREM1")
for (my_TF in TF_list){
  print(my_TF)
  
  my_targets <- TF_modules_list[my_TF]
  print(my_targets)
  
  data_sub_en <- data_enrich
  data_sub_en <- AddModuleScore(data_sub_en, features = my_targets, name = "Enrichment_Score")
  
  p <-   VlnPlot(data_sub_en, features = "Enrichment_Score1", group.by = "clusters_subset", split.by = "age_group_refined", pt.size = 0) +
    ggtitle(paste0("Enrichment of ", my_TF, " target genes"))
  
  pdf(file = paste0(myfolder, "scenic/EN_enrichment/", my_TF, ".pdf"), width = 15, height = 10)
  print(p)
  dev.off()
}
```


### Plot GRN

#### Select TF

```{r}
TF_for_regulons <- markers_table %>%
  arrange(p_val) %>%
  filter(avg_log2FC > 0) %>%
  filter(p_val < 5e-7)

TF_for_regulons <- rownames(TF_for_regulons)
```


#### Select GN

```{r}
list_unique_GN <- unique(regulonTargetsInfo$gene)

# Run FindMarkers to compare the cluster of interest against all other clusters combined
gene_table <- FindMarkers(data_cell, 
                             ident.1 = cluster, 
                             ident.2 = other_clusters, 
                             test.use = "wilcox", 
                             features = list_unique_GN)

# Print the resulting table
gene_table <- gene_table %>%
  top_n(1000, -p_val)
genes <- rownames(gene_table)
```


#### Keep regulons that are from TF_for_regulons

```{r}
# Filter rows based on the "TF" column
regulons <- regulonTargetsInfo[regulonTargetsInfo$TF %in% TF_for_regulons, ]

# Filter rows based on the "gene" column
regulons <- regulons[regulons$gene %in% genes, ]

unique(regulons$TF)
```


### Plot Gene Regulatory Networks

```{r}
net <- igraph::graph_from_data_frame(regulons, directed=T)
deg <- igraph::degree(net, mode = "all", normalized = TRUE)
bet = igraph::betweenness(net, normalized = TRUE)
tg <- tidygraph::as_tbl_graph(net) %>% 
  tidygraph::activate(nodes) %>% 
  dplyr::mutate(label=name,
		size = 2*deg)


set.seed(1)

layout <- layout.fruchterman.reingold(tg, niter=400)
unique_TF <- unique(regulons$TF)


# Fixed size for TF and fixed size for genes
p <- ggraph(tg, layout = layout) +
  geom_edge_link0(color = "grey20", alpha = 0.25) +
  geom_node_point(aes(size = ifelse(name %in% unique_TF, 1200, 400),
                      fill = ifelse(name %in% unique_TF, 'orange', 'turquoise'), 
                      color = 'black'), shape = 21, stroke = 0.3) +
  geom_node_text(aes(label = name, size = size*20), color = "black", repel = F) +
  scale_fill_manual(values = c('orange', 'turquoise')) +
  scale_color_manual(values = 'black') +
  theme_graph(background = "white") +
  theme(legend.position = "none")

#Save plot
pdf(paste0(myfolder, "scenic/figures/", "FB", "_", cluster, ".pdf"), width = 7, height = 7) #FB14 saved in 11x11
p
dev.off()
```


================================================================================
================================================================================
================================================================================
================================================================================
================================================================================







