---
title: "02_3_HDCA_heart_fibroblast_analysis"
author: "Julia Foyer, Ludvig Larsson & Raphaël Mauron"
output: html_document
editor_options: 
  chunk_output_type: console
---

# HDCA heart fibroblast analysis

In this script, a subset of the *high level* data generated in the script `01_HDCA_heart_QC_and_high_level_analysis.Rmd` is selected and re-clustered and analyzed accordingly. It is part of the *deep level* analysis.

NOTE: The path to load the data and the path to save the figures should be changed at your convenience.

Environment: **Docker**


## Settings

### Set library path

```{r setup, include=FALSE}
.libPaths("/home/rstudio/project/renv/library/R-4.3/aarch64-unknown-linux-gnu/")
knitr::opts_chunk$set(echo = TRUE)
```


### Load Libraries

```{r, message = FALSE}
library(Seurat)
library(Matrix)
library(dplyr)
library(harmony)
library(ggplot2)
library(patchwork)
library(openxlsx)
library(plotly)
```


### Set environment variables

```{r}
# Set the directory 
myfolder <- "/home/rstudio/project/output/HDCA_heart_sc_analysis_docker/"
```


### Load global functions

```{r}
source("/home/rstudio/project/code/global_functions.R")
```


### Load filtered data

```{r}
data <- readRDS(paste0(myfolder, "rds_objects/HDCA_heart_sc_full_extended.rds"))
```


### Set seed

```{r}
set.seed(1)
```


## Subset, process, clustering (Extended Fig. 13A)

* Keep all cells annotated as endothelial
* Calculate variable features
* Scale data while regressing out "nCount_RNA", "nFeature_RNA", "percent_ribo", "percent_mito", "percent_hb", "percent_hsp", "S.Score" and "G2M.Score
* Run PCA on scaled data (50 PCs)
* Run harmony (batch correction, 50 PCs)
* Compute UMAP (50 PCs)
* Construct knn adjacency matrix with hnsw_knn
* Cluster cells using cluster_louvain

```{r, message=FALSE}
# Subset the data to keep only fibroblasts
FB <- data[,data$cell.types == "OFT_FB"       |
            data$cell.types == "Int_FB"       |
            data$cell.types == "Valve_MC"     |
            data$cell.types == "EPDC"         |
            data$cell.types == "AnnFibr_FB"   |
            data$cell.types == "Peric_MC"     |
            data$cell.types == "PDE4Chigh_FB" |
            data$cell.types == "Prol_FB"      ]
           
# Process the data
FB <- NormalizeData(FB)

FB <- FindVariableFeatures(FB, nfeatures = 2000)

FB <- ScaleData(FB, do.scale = TRUE, do.center = TRUE, vars.to.regress = c("nCount_RNA", "nFeature_RNA", "percent_ribo", "percent_mito", "percent_hb", "percent_hsp", "S.Score", "G2M.Score"))

FB <- RunPCA(FB, npcs = 50, seed.use = 42)

set.seed(30)
FB <- RunHarmony(FB,
                   group.by.vars = c("sampleID"),
                   reduction = "pca",
                   project.dim = FALSE,
                   reduction.save = "harmony_subset")

FB <- RunUMAP(FB,
                dims = 1:50,
                reduction = "harmony_subset",
                n.neighbors = 10,
                metric = "cosine",
                min.dist = 0.3,
                spread = 1,
                repulsion.strength = 1,
                negative.sample.rate = 5,
                n.epochs = 100,
                reduction.key = "subset",
                reduction.name = "umap_subset",
                seed.use = 42)
gc()

# Compute k nearest neighbours
set.seed(1)
nn <- CreateNNAdjacencyMatrix(FB, reduction = "harmony_subset", k = 10, dims = 1:50)

# add the clusters to the metadata
cl <- LouvainClustering(nn = nn, seed = 1, resolution = 1.5)
FB$clusters_subset <- factor(cl)

pdf(paste0(myfolder, "fibroblast/fibroblast_subclusters.pdf"), width = 10, height = 10)
DimPlot(FB, reduction = "umap_subset", group.by = "clusters_subset", label = T) + labs(title = "Fibroblast subclusters") +
NoLegend()
dev.off()

gc()
```


## Add the annotation

The annotation was imported from an external .csv file. The annotation is added as a new metadata row `cell.types` by matching the cluster numbers.

```{r}
annotation <- read.csv2("/home/rstudio/project/data/metadata/HDCA_heart_SC_annotations_FB_240115.csv", header = 1, sep = ";")
cells_clusters <- setNames(annotation$cell_type, nm = annotation$cluster)
FB$cell.types <- cells_clusters[as.character(FB$clusters_subset)]
rm(annotation, cells_clusters)
gc()
```


## Differentially expressed genes (DEG) for all clusters (Supp Tab. 6)

```{r}
FB <- SetIdent(FB, value = "cell.types")
detable <- FindAllMarkers(FB, logfc.threshold = 0.25,
                          only.pos = TRUE, 
                          min.pct = 0.05)

# Only keep DEGs that have a p-value below 0.05.
detable <- detable[detable$p_val_adj < 0.01, ]

# Adding two extra columns giving the difference (and log difference) between pct1 and pct2.
# pct1 is the percentage of cells in the specific cluster that express this gene.
# pct2 is the percentage of cells in all other clusters that express this gene.
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2((detable$pct.1*99+1) / (detable$pct.2*99+1))

# Save the detable as a csv file.
write.csv2(detable, paste0(myfolder, "supplementary_tables/supplementary_table_6_DEG_FB.csv"))
```


### UMAP with annotation (Fig. 6A)

```{r}
values_to_keep <- c("7", "3", "18", "14", "5", "9", "19", "2", "12", "4", "13", "11", "17", "8", "1", "6", "15", "20")

data_FB <- FB[, FB$clusters_subset %in% values_to_keep]

pdf(file = paste0(myfolder, "fibroblast/fibroblast_clusters_remaining_annot.pdf"), width = 10, height = 10)
DimPlot(data_FB, reduction = "umap_subset", group.by = "cell.types", label = T) + NoLegend() + labs(title = "Fibroblast subclusters")
dev.off()
```


### DEG FB - top5 ordered (Supp Fig. 6A)

```{r}
values_to_keep <- c("7", "3", "18", "14", "5", "9", "19", "2", "12", "4", "13", "11", "17", "8", "1", "6", "15", "20")

data_FB <- FB[, FB$clusters_subset %in% values_to_keep]

# Select a new samples size for each cluster, to get groups that are equally (or almost equally) large.
sample_size <- table(data_FB$clusters_subset)
sample_size[sample_size > 150] <- 150
sample_size

# Randomly sample the cells that will be included when calculating DGE (same number of cells as selected above).
# Samples without replacement, so each cell only can be included once.
DGE_cells <- lapply(names(sample_size), function(x){ 
  set.seed(1)
  sample(colnames(data_FB)[data_FB$clusters_subset == x], size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# Select all data that relates to the sampled DGE cells.
DGE_DATA <- data_FB[, DGE_cells]

# reorder the cell types desired
DGE_DATA$cell.types <- factor(DGE_DATA$cell.types, 
                                              levels = c("EPDC_1", 
                                                         "EPDC_2", 
                                                         "Infl_FB", 
                                                         "FAP", 
                                                         "Int_FB_1", 
                                                         "Int_FB_2", 
                                                         "Int_FB_3", 
                                                         "Prol_FB_2", 
                                                         "Prol_FB_1", 
                                                         "VIC", 
                                                         "Valve_MC_1", 
                                                         "Valve_MC_2", 
                                                         "CALN1high_FB", 
                                                         "AnnFibr_FB^fg", 
                                                         "Adv_FB_1", 
                                                         "Adv_FB_2", 
                                                         "Peric_MC^fg", 
                                                         "PDE4Chigh_FB^fg"))

DGE_DATA <- SetIdent(DGE_DATA, value = "cell.types")

# The logfc threshold is set to 0.1 as compared to the default 0.25, which makes the detection of weaker signals possible (but it takes more computational power).
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 300,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05)

# Only keep DEG that have a p-value below 0.05.
detable <- detable[detable$p_val < 0.05, ]

# Adding two extra columns giving the difference (and log difference) between pct1 and pct2.
# pct1 is the percentage of cells in the specific cluster that express this gene.
# pct2 is the percentage of cells in all other clusters that express this gene.
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2((detable$pct.1*99+1) / (detable$pct.2*99+1))

# Pick the 40 genes with the lowest p value for each cluster, then 20 highest based on pct.diff, then 5 highest based on avg_log2FC.
detable %>%
  group_by(cluster) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(5, avg_log2FC) ->
  top5

pdf(paste0(myfolder, "fibroblast/DEG_top5_FB_ordered.pdf"), width = 10, height = 17)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top5 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()

gc()
rm(detable, DGE_DATA, top5, DGE_cells, sample_size)
```


### DEG FB selection of markers (Extended Fig. 13D)

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("MYH7", "MYH6", "MYBPC3", "TNNI1", "MYOZ2", "FABP3", "TMSB10", "S100A11", "GNG11", "FABP4", "VWF", "CDH5", "PECAM1", "LYVE1", "NPR3", "BMPER", "POSTN", "APCDD1", "TWIST1", "DDR2", "TCF21", "COL1A1", "MYH11", "ENPEP", "ITLN1", "PCSK1N", "MPZ", "C1QC", "JCHAIN", "CD3E", "KLRB1", "UBE2C", "H1-1")

FB <- SetIdent(FB, value = "clusters_subset")

p <- DotPlot(FB, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "fibroblast/dotplot_markers_FB.pdf"), width = 8, height = 8)
print(p)
dev.off()

rm(p)
gc()
```


### DEG FB - top10 ordered - FB_4, FB_11, FB_13 (Supp Fig. 6E)

```{r}
values_to_keep <- c("4", "11", "13")

data_FB <- FB[, FB$clusters_subset %in% values_to_keep]

# Select a new samples size for each cluster, to get groups that are equally (or almost equally) large.
sample_size <- table(data_FB$clusters_subset)
sample_size[sample_size > 150] <- 150
sample_size

# Randomly sample the cells that will be included when calculating DGE (same number of cells as selected above).
# Samples without replacement, so each cell only can be included once.
DGE_cells <- lapply(names(sample_size), function(x){ 
  set.seed(1)
  sample(colnames(data_FB)[data_FB$clusters_subset == x], size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# Select all data that relates to the sampled DGE cells.
DGE_DATA <- data_FB[, DGE_cells]

# reorder the cell types desired
DGE_DATA$cell.types <- factor(DGE_DATA$cell.types, levels = c("VIC", 
                                                         "Valve_MC_2", 
                                                         "Valve_MC_1"))

DGE_DATA <- SetIdent(DGE_DATA, value = "cell.types")

# The logfc threshold is set to 0.1 as compared to the default 0.25, which makes the detection of weaker signals possible (but it takes more computational power).
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 300,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05)

# Only keep DEG that have a p-value below 0.05.
detable <- detable[detable$p_val < 0.05, ]

# Adding two extra columns giving the difference (and log difference) between pct1 and pct2.
# pct1 is the percentage of cells in the specific cluster that express this gene.
# pct2 is the percentage of cells in all other clusters that express this gene.
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2((detable$pct.1*99+1) / (detable$pct.2*99+1))

# Pick the 40 genes with the lowest p value for each cluster, then 20 highest based on pct.diff, then 5 highest based on avg_log2FC.
detable %>%
  group_by(cluster) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(10, avg_log2FC) ->
  top10

pdf(paste0(myfolder, "fibroblast/DEG_top10_FB_ordered_4_11_13.pdf"), width = 4, height = 7.5)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top10 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
dev.off()

gc()
rm(detable, DGE_DATA, top10, DGE_cells, sample_size)
```


### FeaturePlot (Supp Fig. 6B)

```{r}
values_to_keep <- c("7", "3", "18", "14", "5", "9", "19", "2", "12", "4", "13", "11", "17", "8", "1", "6", "15", "20")

data_FB <- FB[, FB$clusters_subset %in% values_to_keep]

feature_list <- c("WT1", "PLA2G5", "MKI67", "TCF21", "CD34", "CLEC3B", "APCDD1", "ERBB4",  "PENK",  "ACTA2",  "CCL2", "FGF14", "PDGFRB")

for (feature in feature_list){
  p <-   FeaturePlot(data_FB, reduction = "umap_subset", features = feature, raster = FALSE, order = FALSE)

  pdf(paste0(myfolder, "fibroblast/feature_plots/", feature, ".pdf"), width = 10, height = 10)
  print(p)
  dev.off()
}
```


### Cell type proportion by age group in entire FB object (Supp Fig. 6C)

```{r}
age1 <- FB[,FB$age <= 6]
age2 <- FB[,FB$age == 7 | FB$age == 8 | FB$age == 8.5]
age3 <- FB[,FB$age == 9 | FB$age == 10 | FB$age == 11.5]
age4 <- FB[,FB$age >= 12]

FB_age <- c(age1, age2, age3, age4)
rm(age1, age2, age3, age4)

# create df by age grouping the number of cell per cluster
cell_by_cluster_entire_1 <- as.data.frame(table(FB_age[[1]]@meta.data[["clusters_subset"]]))
cell_by_cluster_entire_2 <- as.data.frame(table(FB_age[[2]]@meta.data[["clusters_subset"]]))
cell_by_cluster_entire_3 <- as.data.frame(table(FB_age[[3]]@meta.data[["clusters_subset"]]))
cell_by_cluster_entire_4 <- as.data.frame(table(FB_age[[4]]@meta.data[["clusters_subset"]]))

# add the proportion of each cell type
cell_by_cluster_entire_1$proportion <- cell_by_cluster_entire_1$Freq/sum(cell_by_cluster_entire_1$Freq)
cell_by_cluster_entire_2$proportion <- cell_by_cluster_entire_2$Freq/sum(cell_by_cluster_entire_2$Freq)
cell_by_cluster_entire_3$proportion <- cell_by_cluster_entire_3$Freq/sum(cell_by_cluster_entire_3$Freq)
cell_by_cluster_entire_4$proportion <- cell_by_cluster_entire_4$Freq/sum(cell_by_cluster_entire_4$Freq)

# export the df as .csv
write.csv2(cell_by_cluster_entire_1, paste0(myfolder, "fibroblast/cell_by_cluster_entire_1.csv"))
write.csv2(cell_by_cluster_entire_2, paste0(myfolder, "fibroblast/cell_by_cluster_entire_2.csv"))
write.csv2(cell_by_cluster_entire_3, paste0(myfolder, "fibroblast/cell_by_cluster_entire_3.csv"))
write.csv2(cell_by_cluster_entire_4, paste0(myfolder, "fibroblast/cell_by_cluster_entire_4.csv"))
```


## Export fibroblasts metadata and Seurat object

```{r}
saveRDS(FB[[]], paste0(myfolder, "/rds_objects/fibroblasts_metadata.rds"))
saveRDS(FB, paste0(myfolder, "/rds_objects/fibroblasts.rds"))
```

```{r}
FB <- readRDS(paste0(myfolder, "rds_objects/fibroblasts.rds"))
```


================================================================================
================================================================================
================================================================================
================================================================================
================================================================================








