---
title: "02_4_HDCA_heart_innervation_analysis"
author: "Raphaël Mauron"
output: html_document
editor_options: 
  chunk_output_type: console
---

# HDCA heart innervation analysis

In this script, a subset of the *high level* data generated in the script `01_HDCA_heart_QC_and_high_level_analysis.Rmd` is selected and re-clustered and analyzed accordingly. It is part of the *deep level* analysis.

NOTE: The path to load the data and the path to save the figures should be changed at your convenience.

Environment: **Docker**


## Settings

### Set library path

```{r setup, include=FALSE}
.libPaths("/home/rstudio/project/renv/library/R-4.3/aarch64-unknown-linux-gnu/")
knitr::opts_chunk$set(echo = TRUE)
```


### Load Libraries

```{r, message = FALSE}
library(Seurat)
library(Matrix)
library(dplyr)
library(harmony)
library(ggplot2)
library(patchwork)
library(openxlsx)
library(plotly)
```


### Set environment variables

```{r}
# Set the directory 
myfolder <- "/home/rstudio/project/output/HDCA_heart_sc_analysis_docker/"
```


### Load global functions

```{r}
source("/home/rstudio/project/code/global_functions.R")
```


### Load filtered data

```{r}
data <- readRDS(paste0(myfolder, "rds_objects/HDCA_heart_sc_full_extended.rds"))
```


### Set seed

```{r}
set.seed(1)
```


## Subset, process, clustering

* Keep all cells annotated as endothelial
* Calculate variable features
* Scale data while regressing out "nCount_RNA", "nFeature_RNA", "percent_ribo", "percent_mito", "percent_hb", "percent_hsp", "S.Score" and "G2M.Score
* Run PCA on scaled data (50 PCs)
* Run harmony (batch correction, 50 PCs)
* Compute UMAP (50 PCs)
* Construct knn adjacency matrix with hnsw_knn
* Cluster cells using cluster_louvain

```{r, message=FALSE}
# Subset the data to keep only endothelial cells
IN <- data[,data$high_level_clusters_removed == 25   | #25
            data$high_level_clusters_removed == 26 ] #26
           
# Remove all genes with 0 counts
IN <- IN[rowSums(GetAssayData(IN, slot = "counts")) > 0, ]

# Process the data
IN <- IN %>% 
  NormalizeData() %>% 
  FindVariableFeatures(nfeatures = 3000) %>% 
  ScaleData(vars.to.regress = c("nCount_RNA", "nFeature_RNA", 
                                "percent_ribo", "percent_mito", 
                                "percent_hb", "percent_hsp", 
                                "S.Score", "G2M.Score")) 

IN <- RunPCA(IN, npcs = 50, seed.use = 42)

set.seed(1)
IN <- RunHarmony(IN,
                 group.by.vars = c("sampleID"),
                 reduction = "pca",
                 project.dim = F,
                 reduction.save = "harmony_subset")

IN <- RunUMAP(IN,
              dims = 1:50,
              reduction = "harmony_subset",
              n.neighbors = 20, # Maybe 20 looks better, was set to 10
              metric = "cosine",
              min.dist = 0.3,
              spread = 1,
              repulsion.strength = 1,
              negative.sample.rate = 5,
              n.epochs = 100,
              reduction.key = "subset",
              reduction.name = "umap_subset",
              seed.use = 42)

gc()

# Compute k nearest neighbours
set.seed(1)
nn <- CreateNNAdjacencyMatrix(data = IN, reduction = "harmony_subset", k = 10, dims = 1:50)

# add the clusters to the metadata
cl <- LouvainClustering(nn = nn, seed = 1, resolution = 1) # change resolution
IN$clusters_subset <- factor(cl)

pdf(paste0(myfolder, "innervation/innervationl_subclusters.pdf"), width = 10, height = 10)
DimPlot(IN, reduction = "umap_subset", group.by = "clusters_subset", label = T) + labs(title = "Innervation subclusters")
dev.off()

gc()
```


### Subset good clusters and rebuild umap

```{r}
IN_subset <- IN[,
                  IN$clusters_subset == 1   |
                  IN$clusters_subset == 3   |
                  IN$clusters_subset == 4   |
                  IN$clusters_subset == 7   |
                  IN$clusters_subset == 8   |
                  IN$clusters_subset == 11   |
                  IN$clusters_subset == 12   |
                  IN$clusters_subset == 13   |
                  IN$clusters_subset == 17   |
                  IN$clusters_subset == 19   ]
IN_subset$clusters_subset <- factor(IN_subset$clusters_subset, levels = c("1", "3", "4", "7", "8", "11", "12", "13", "17", "19"))

IN_subset <- RunUMAP(IN_subset,
              dims = 1:50,
              reduction = "harmony_subset",
              n.neighbors = 20, # Maybe 20 looks better, was set to 10
              metric = "cosine",
              min.dist = 0.3,
              spread = 1,
              repulsion.strength = 1,
              negative.sample.rate = 5,
              n.epochs = 100,
              reduction.key = "subset",
              reduction.name = "umap_subset_sub",
              seed.use = 42)

pdf(paste0(myfolder, "innervation/innervationl_subclusters_selection.pdf"), width = 10, height = 10)
DimPlot(IN_subset, reduction = "umap_subset_sub", group.by = "clusters_subset", label = T)
dev.off()

```

```{r}
saveRDS(IN_subset, paste0(myfolder, "/rds_objects/innervation_selection.rds"))
```

## UMAP 3D

```{r}
IN <- RunUMAP(IN,
                dims = 1:50,
                reduction = "harmony_subset",
                n.neighbors = 10,
                metric = "cosine",
                min.dist = 0.3,
                spread = 1,
                repulsion.strength = 1,
                negative.sample.rate = 5,
                n.epochs = 100,
                reduction.key = "subset",
                reduction.name = "umap_subset_3d",
                seed.use = 42,
                n.components = 3)
gc()

df_IN <- data.frame(IN@reductions$umap_subset_3d@cell.embeddings)
df_IN <- data.frame(df_IN, seurat_clusters=IN$clusters_subset, sampleid = IN$sampleID , age = IN$age)
colnames(df_IN)[1:3] <- c("UMAP_1", "UMAP_2", "UMAP_3")
pal <- c(scales::hue_pal()(8), RColorBrewer::brewer.pal(9, "Set1"), RColorBrewer::brewer.pal(8, "Set2"))

p <- plot_ly(df_IN, 
             x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3,
             color = ~seurat_clusters,
             colors = pal,
             size = 0.5) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = "UMAP_1"), yaxix = list(title = "UMAP_2"), zaxis = list(title = "UMAP_3")))

htmlwidgets::saveWidget(p, paste0(myfolder, "innervation/umap3d.html"))
```


## Run DEG for all clusters

```{r, message=FALSE}
IN <- SetIdent(IN, value = "clusters_subset")
de_markers <- FindAllMarkers(IN, logfc.threshold = 0.25,
                             only.pos = TRUE, 
                             min.pct = 0.05)

openxlsx::write.xlsx(de_markers %>% filter(p_val_adj < 0.01),
                     file = paste0(myfolder, "innervation/detable_innervation.xlsx"))
```


### DE dotplot

```{r}
top20_markers <- de_markers %>% 
  group_by(cluster) %>% 
  slice_head(n = 20) %>% 
  pull(gene)

IN <- SetIdent(IN, value = "clusters_subset")

p <- DotPlot(IN, features = unique(top20_markers)) +
  coord_flip() +  # This flips the plot vertically
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  labs(x = "Cluster", y = "Genes")  # Flipped labels

pdf(paste0(myfolder, "innervation/dotplot_de_genes.pdf"), width = 10, height = 100)
print(p)
dev.off()

rm(p)
gc()
```



### Save images and data

```{r}
# IN subclusters projected on high level UMAP
p <- DimPlot(IN, reduction = "umap_oi", group.by = "clusters_subset", label = TRUE) + labs(title = "Subclustered innervation over the nonsubsetted UMAP")
pdf(file = paste0(myfolder, "innervation/clusters_innervation_big_umap.pdf"),
     width = 10, height = 10)
p
dev.off()

# High level IN clusters projected on the deep level IN UMAP or Rough clusters
p <- DimPlot(IN, reduction = "umap_subset", group.by = "clusters_louvain_oi", label = TRUE) + labs(title = "Rough clusters on the new innervation umap")
pdf(file = paste0(myfolder, "innervation/rough_clusters_innervation.pdf"),
     width = 10, height = 10)
p
dev.off()

# nCounts
pdf(paste0(myfolder, "innervation/nCounts_innervation.pdf"), width = 10, height = 10)
FeaturePlot(IN, reduction = "umap_subset", features = "nCount_RNA", raster = FALSE)
dev.off()

# nFeatures
pdf(paste0(myfolder, "innervation/nFeatures_innervation.pdf"), width = 10, height = 10)
FeaturePlot(IN, reduction = "umap_subset", features = "nFeature_RNA", raster = FALSE)
dev.off()
```


### Cell type proportion by age group in entire IN object

```{r}
age1 <- IN[,IN$age <= 6]
age2 <- IN[,IN$age == 7 | IN$age == 8 | IN$age == 8.5]
age3 <- IN[,IN$age == 9 | IN$age == 10 | IN$age == 11.5]
age4 <- IN[,IN$age >= 12]

IN_age <- c(age1, age2, age3, age4)
rm(age1, age2, age3, age4)

# create df by age grouping the number of cell per cluster
cell_by_cluster_entire_1 <- as.data.frame(table(IN_age[[1]]@meta.data[["clusters_subset"]]))
cell_by_cluster_entire_2 <- as.data.frame(table(IN_age[[2]]@meta.data[["clusters_subset"]]))
cell_by_cluster_entire_3 <- as.data.frame(table(IN_age[[3]]@meta.data[["clusters_subset"]]))
cell_by_cluster_entire_4 <- as.data.frame(table(IN_age[[4]]@meta.data[["clusters_subset"]]))

# add the proportion of each cell type
cell_by_cluster_entire_1$proportion <- cell_by_cluster_entire_1$Freq/sum(cell_by_cluster_entire_1$Freq)
cell_by_cluster_entire_2$proportion <- cell_by_cluster_entire_2$Freq/sum(cell_by_cluster_entire_2$Freq)
cell_by_cluster_entire_3$proportion <- cell_by_cluster_entire_3$Freq/sum(cell_by_cluster_entire_3$Freq)
cell_by_cluster_entire_4$proportion <- cell_by_cluster_entire_4$Freq/sum(cell_by_cluster_entire_4$Freq)

# export the df as .csv
write.csv2(cell_by_cluster_entire_1, paste0(myfolder, "innervation/cell_by_cluster_entire_1.csv"))
write.csv2(cell_by_cluster_entire_2, paste0(myfolder, "innervation/cell_by_cluster_entire_2.csv"))
write.csv2(cell_by_cluster_entire_3, paste0(myfolder, "innervation/cell_by_cluster_entire_3.csv"))
write.csv2(cell_by_cluster_entire_4, paste0(myfolder, "innervation/cell_by_cluster_entire_4.csv"))
```


### DEG rough IN - top20

From the cells in IN, plot the DotPlot on the High Level origin of those cell (rough)

```{r}
# Select a new samples size for each cluster, to get groups that are equally (or almost equally) large.
sample_size <- table(IN$clusters_subset)
sample_size[sample_size > 150] <- 150
sample_size

# Randomly sample the cells that will be included when calculating DGE (same number of cells as selected above).
# Samples without replacement, so each cell only can be included once.
DGE_cells <- lapply(names(sample_size), function(x){ 
  set.seed(1)
  sample(colnames(IN)[IN$clusters_subset == x], size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# Select all data that relates to the sampled DGE cells.
DGE_DATA <- data[, DGE_cells]

#Find differential expressed genes based on groups defined by "high_level_clusters_removed".
DGE_DATA <- SetIdent(DGE_DATA, value = "high_level_clusters_removed")

# The logfc threshold is set to 0.1 as compared to the default 0.25, which makes the detection of weaker signals possible (but it takes more computational power).
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 300,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05)

# Only keep DEG that have a p-value below 0.05.
detable <- detable[detable$p_val < 0.05, ]

# Adding two extra columns giving the difference (and log difference) between pct1 and pct2.
# pct1 is the percentage of cells in the specific cluster that express this gene.
# pct2 is the percentage of cells in all other clusters that express this gene.
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2((detable$pct.1*99+1) / (detable$pct.2*99+1))

# Save the detable as a csv file.
write.csv2(detable, paste0(myfolder, "innervation/detable_rough_clusters.csv"))

# Pick the 60 genes with the lowest p value for each age group, then 40 highest based on pct.diff, then 20 highest based on log.pct.diff.
detable %>%
  group_by(cluster) %>%
  top_n(-60, p_val) %>%
  top_n(40, pct.diff) %>%
  top_n(20, log.pct.diff) ->
  top20

# Genes get matched to one specific cluster, and the factor is leveled based on cluster number.
ord <- factor(sapply(unique(as.character(top20$gene)), function(x){niceRplots::getcluster(DGE_DATA, x, "high_level_clusters_removed")}))

pdf(paste0(myfolder, "innervation/DGE_rough_IN.pdf"), width = 15, height = length(ord)/6+1)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = unique(as.character(top20$gene))[order(as.numeric(as.character(ord)))]) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  #labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))
dev.off()

rm(detable, DGE_DATA, top20, DGE_cells, ord, sample_size)
```


### DEG IN - top5 ordered

```{r}
values_to_keep <- c("7", "3", "18", "14", "5", "9", "19", "2", "12", "4", "13", "11", "17", "8", "1", "6", "15", "20", "21")

data_IN <- IN[, IN$clusters_subset %in% values_to_keep]

# Select a new samples size for each cluster, to get groups that are equally (or almost equally) large.
sample_size <- table(data_IN$clusters_subset)
sample_size[sample_size > 150] <- 150
sample_size

# Randomly sample the cells that will be included when calculating DGE (same number of cells as selected above).
# Samples without replacement, so each cell only can be included once.
DGE_cells <- lapply(names(sample_size), function(x){ 
  set.seed(1)
  sample(colnames(data_IN)[data_IN$clusters_subset == x], size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# Select all data that relates to the sampled DGE cells.
DGE_DATA <- data_IN[, DGE_cells]

# reorder the cell types desired
DGE_DATA$clusters_subset <- factor(DGE_DATA$clusters_subset, 
                                              levels = c("7", 
                                                         "3", 
                                                         "18", 
                                                         "14", 
                                                         "5", 
                                                         "9", 
                                                         "19", 
                                                         "2", 
                                                         "12", 
                                                         "4", 
                                                         "13", 
                                                         "11", 
                                                         "17", 
                                                         "8", 
                                                         "1", 
                                                         "6", 
                                                         "15", 
                                                         "20", 
                                                         "21"))

#Find differential expressed genes based on groups defined by "clusters_subset_split".
DGE_DATA <- SetIdent(DGE_DATA, value = "clusters_subset")

# The logfc threshold is set to 0.1 as compared to the default 0.25, which makes the detection of weaker signals possible (but it takes more computational power).
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 300,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05)

# Only keep DEG that have a p-value below 0.05.
detable <- detable[detable$p_val < 0.05, ]

# Adding two extra columns giving the difference (and log difference) between pct1 and pct2.
# pct1 is the percentage of cells in the specific cluster that express this gene.
# pct2 is the percentage of cells in all other clusters that express this gene.
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2((detable$pct.1*99+1) / (detable$pct.2*99+1))

# Pick the 40 genes with the lowest p value for each cluster, then 20 highest based on pct.diff, then 5 highest based on avg_log2FC.
detable %>%
  group_by(cluster) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(5, avg_log2FC) ->
  top5

pdf(paste0(myfolder, "innervation/DEG_top5_IN_ordered.pdf"), width = 10, height = 17)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top5 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 0.5))
dev.off()

gc()
rm(detable, DGE_DATA, top5, DGE_cells, sample_size)
```


### DEG IN selection of markers

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("SOX10",
                "FOXD3",
                "PHOX2B", 
                "PHOX2A",
                "PRPH",
                "ASCL1",
                "CHGB",
                "CHGA",
                "PENK",
                "MBP",
                "PMP22",
                "HTR3A",
                "HTR3B",
                "TH"
                )

pdf(paste0(myfolder, "innervation/dotplot_markers_IN.pdf"), width = 8, height = 8)
DotPlot(IN, features = my_markers) +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete() +
  scale_y_discrete(limits = rev) +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))
dev.off()

rm(p)
gc()
```


### DEG IN selection of markers 2

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("SOX10", #for glia
                "FOXD3",
                "MBP",
                "MPZ",
                "PRPH", #for neurons
                "TH",
                "PHOX2B", 
                #"PHOX2A",
                #"ASCL1",
                #"CHGA",
                "PENK", #for chromaffin
                "PNMT",
                #"TH",
                #"MBP",
                #"CHGB",
                #"PMP22",
                #"HTR3A",
                #"HTR3B",
                "SST" #for VIP
                )

values_to_keep <- c("7", "3", "18", "14", "5", "9", "19", "2", "12", "4", "13", "11", "17", "8", "1", "6", "15", "20", "21")

data_IN <- IN[, IN$clusters_subset %in% values_to_keep]
data_IN$clusters_subset <- factor(data_IN$clusters_subset, level=values_to_keep)

data_IN <- SetIdent(data_IN, value = "clusters_subset")

pdf(paste0(myfolder, "innervation/dotplot_markers_IN_2.pdf"), width = 8, height = 8)
DotPlot(data_IN, features = my_markers) +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete() +
  scale_y_discrete(limits = rev) +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))
dev.off()

gc()
```


### DEG IN-4, 7, 12, 11, 8, 19, 17, 3, 13, 1 - top5 ordered

```{r}
values_to_keep <- c("4", "7", "12", "11", "8", "19", "17", "3", "13", "1")

data_IN <- IN[, IN$clusters_subset %in% values_to_keep]

# Select a new samples size for each cluster, to get groups that are equally (or almost equally) large.
sample_size <- table(data_IN$clusters_subset)
sample_size[sample_size > 150] <- 150
sample_size

# Randomly sample the cells that will be included when calculating DGE (same number of cells as selected above).
# Samples without replacement, so each cell only can be included once.
DGE_cells <- lapply(names(sample_size), function(x){ 
  set.seed(1)
  sample(colnames(data_IN)[data_IN$clusters_subset == x], size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# Select all data that relates to the sampled DGE cells.
DGE_DATA <- data_IN[, DGE_cells]

# reorder the cell types desired
DGE_DATA$clusters_subset <- factor(DGE_DATA$clusters_subset, 
                                              levels = c("4", "7", "12", "11", "8", "19", "17", "3", "13", "1"))

#Find differential expressed genes based on groups defined by "clusters_subset_split".
DGE_DATA <- SetIdent(DGE_DATA, value = "clusters_subset")

# The logfc threshold is set to 0.1 as compared to the default 0.25, which makes the detection of weaker signals possible (but it takes more computational power).
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 300,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05)

# Only keep DEG that have a p-value below 0.05.
detable <- detable[detable$p_val < 0.05, ]

# Adding two extra columns giving the difference (and log difference) between pct1 and pct2.
# pct1 is the percentage of cells in the specific cluster that express this gene.
# pct2 is the percentage of cells in all other clusters that express this gene.
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2((detable$pct.1*99+1) / (detable$pct.2*99+1))

# Pick the 40 genes with the lowest p value for each cluster, then 20 highest based on pct.diff, then 5 highest based on avg_log2FC.
detable %>%
  group_by(cluster) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(5, avg_log2FC) ->
  top5

pdf(paste0(myfolder, "innervation/DEG_top5_IN_sel_cl.pdf"), width = 6, height = 10)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top5 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))
dev.off()

gc()
rm(detable, DGE_DATA, top5, DGE_cells, sample_size)
```


### DEG IN selection of cl, spe markers

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("SOX10", "FOXD3", "MPZ", "PMP22", "MBP", "ASCL1", "PHOX2B", "PHOX2A", "HTR3A", "PRPH", "CHAT", "SST", "TH", "DBH", "PNMT", "CHGA", "CHGB", "PENK", "EPAS1", "COX4I2", "NDUFS2", "HIGD1C")

values_to_keep <- c("4", "7", "12", "11", "8", "19", "17", "3", "13", "1")

data_IN <- IN[, IN$clusters_subset %in% values_to_keep]
data_IN$clusters_subset <- factor(data_IN$clusters_subset, level=values_to_keep)

data_IN <- SetIdent(data_IN, value = "clusters_subset")

pdf(paste0(myfolder, "innervation/dotplot_markers_IN_cl_ord_3.pdf"), width = 7, height = 4)
DotPlot(data_IN, features = my_markers) +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete() +
  scale_y_discrete(limits = rev) +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))
dev.off()

gc()
```


### FeaturePlot

```{r}
feature_list <- c("TH", "PENK", "CHGB", "CHAT", "ELAVL4", "STMN2",
                  "SOX10", "FOXD3", "MBP", "DHH",
                  "PHOX2B", "PHOX2A", "ASCL1",
                  "PRRX1", "PRRX2", "TWIST1", "TWIST2",
                  "DCX", "SOX2", "PRPH", "ISL1",
                  "MKI67", "CENPA", "TOP2A",
                  "FLI1", "PECAM1")

for (feature in feature_list){
  p <-   FeaturePlot(IN, reduction = "umap_subset", features = feature, raster = FALSE, order = TRUE)

  pdf(paste0(myfolder, "innervation/feature_plots/", feature, ".pdf"), width = 10, height = 10)
  print(p)
  dev.off()
}
```


## Export innervation metadata and Seurat object

```{r}
saveRDS(IN[[]], paste0(myfolder, "/rds_objects/innervation_metadata.rds"))
saveRDS(IN, paste0(myfolder, "/rds_objects/innervation.rds"))
```

```{r}
IN <- readRDS(paste0(myfolder, "rds_objects/innervation.rds"))
```


================================================================================
================================================================================
================================================================================
================================================================================
================================================================================







