---
title: "11_HDCA_heart_gene_sets_enrichment"
author: "Raphaël Mauron"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Enrichment of heart markers in SC data

Pathological gene sets enrichment analysis.

NOTE: The path to load the data and the path to save the figures should be changed at your convenience.

Environment: **Docker**

## Settings

### Set library path


```{r setup, include=FALSE}
.libPaths("/home/rstudio/project/renv/library/R-4.3/aarch64-unknown-linux-gnu/")
knitr::opts_chunk$set(echo = TRUE)
```


### Set environment variables

```{r}
# Set the directory 
myfolder <- "/home/rstudio/project/output/HDCA_heart_sc_analysis_docker/"
```


### Load Seurat object

```{r, message = FALSE}
data <- readRDS(paste0(myfolder, "rds_objects/HDCA_heart_sc_full_extended.rds"))
CM <- readRDS(file = paste0(myfolder, "rds_objects/cardiomyocytes02.rds"))
EN <- readRDS(file = paste0(myfolder, "rds_objects/endothelial.rds"))
FB <- readRDS(paste0(myfolder, "rds_objects/fibroblasts.rds"))
IN <- readRDS(paste0(myfolder, "rds_objects/innervation.rds"))
```


### Load libraries

```{r, message = FALSE}
library(Seurat)
library(Matrix)
library(dplyr)
library(data.table)
library(tibble)
library(openxlsx)
library(patchwork)
library(tidyr)
library(ggplot2)
```



## Gene Set Enrichment Analysis (GSEA)

### Gene sets

```{r}
# Keep list in GeneSets that corresponds to ST clusters of interest
# GeneSets <- list(c("MEIS2", "ISL1", "LGR5", "PDGFRA", "BMP4"),
#                  c("SOX4", "HES4", "CTHRC1", "IGFBP2"))

GeneSets_path <- read.csv2( "/home/rstudio/project/data/metadata/pathological_gene_enrichment.csv")
gene_list <- as.list(GeneSets_path)
gene_list <- lapply(gene_list, function(vec) vec[vec != ""])
print(gene_list)
rm(GeneSets_path)
gc()
```


### For CM

```{r}
# Compute module score on single-cell data
CM <- AddModuleScore(CM, features = GeneSets, name = "Enrichment_Score")

# Rename the Enrichment_Score meta.data
# Create new_column_names using the list of numbers
new_column_names <- c("CVP", "FVP")
  
# Use a loop to rename the columns based on the list of indices and names
i_start <- as.numeric(dim(CM@meta.data)[2]-length(GeneSets)+1)
i_end <- as.numeric(dim(CM@meta.data)[2])
i_column <- as.numeric(dim(CM@meta.data)[2]-length(GeneSets))

for (i in i_start:i_end) {
  colnames(CM@meta.data)[i] <- new_column_names[i - i_column]
}

# Plot violin plots
# If you have 10 clusters, the scores should now be named Enrichment_Score1, Enrichment_Scoree2, ..., Enrichment_Score10
CM_sub <- CM[, CM$clusters_subset_split %in% c("6",  "1", "10", "8-1", "8-3", "17", "8-2", "18", "4-2",  "4-1", "3", "14",  "9",  "12", "19-2",  "19-1", "7",  "2", "5")]
selected_clusters <- c("6",  "1", "10", "8-1", "8-3", "17", "8-2", "18", "4-2",  "4-1", "3", "14",  "9",  "12", "19-2",  "19-1", "7",  "2", "5")
CM_sub$clusters_subset_split <- factor(CM_sub$clusters_subset_split, levels = selected_clusters)

# Plot
pdf(file = paste0(myfolder, "ST_enrichment/Violin_enrichment_CVP_FVP_CM.pdf"), width = 10, height = 3)
VlnPlot(CM_sub, features = new_column_names, group.by = "clusters_subset_split", pt.size = 0) +
  plot_layout(ncol = 2)
dev.off()
```


### For EN

```{r}
# Compute module score on single-cell data
EN <- AddModuleScore(EN, features = GeneSets, name = "Enrichment_Score")

# Rename the Enrichment_Score meta.data
# Create new_column_names using the list of numbers
new_column_names <- c("CVP", "FVP")

# Use a loop to rename the columns based on the list of indices and names
i_start <- as.numeric(dim(EN@meta.data)[2]-length(GeneSets)+1)
i_end <- as.numeric(dim(EN@meta.data)[2])
i_column <- as.numeric(dim(EN@meta.data)[2]-length(GeneSets))

for (i in i_start:i_end) {
  colnames(EN@meta.data)[i] <- new_column_names[i - i_column]
}

# Plot violin plots
# If you have 10 clusters, the scores should now be named Enrichment_Score1, Enrichment_Scoree2, ..., Enrichment_Score10
EN_sub <- EN[, EN$clusters_subset %in% c("1", "12", "3", "6", "18", "17", "7", "19", "15", "13", "10", "2", "5", "21", "16", "20", "11", "14")]
selected_clusters <- c("1", "12", "3", "6", "18", "17", "7", "19", "15", "13", "10", "2", "5", "21", "16", "20", "11", "14")
EN$clusters_subset <- factor(EN_sub$clusters_subset, levels = selected_clusters)

# Plot
pdf(file = paste0(myfolder, "ST_enrichment/Violin_enrichment_CVP_FVP_EN.pdf"), width = 10, height = 3)
VlnPlot(EN_sub, features = new_column_names, group.by = "clusters_subset", pt.size = 0) + 
  plot_layout(ncol = 2)
dev.off()
```


### For FB

```{r}
# Compute module score on single-cell data
FB <- AddModuleScore(FB, features = GeneSets, name = "Enrichment_Score")

# Rename the Enrichment_Score meta.data
# Create new_column_names using the list of numbers
new_column_names <- c("CVP", "FVP")

# Use a loop to rename the columns based on the list of indices and names
i_start <- as.numeric(dim(FB@meta.data)[2]-length(GeneSets)+1)
i_end <- as.numeric(dim(FB@meta.data)[2])
i_column <- as.numeric(dim(FB@meta.data)[2]-length(GeneSets))

for (i in i_start:i_end) {
  colnames(FB@meta.data)[i] <- new_column_names[i - i_column]
}

# Plot violin plots
# If you have 10 clusters, the scores should now be named Enrichment_Score1, Enrichment_Scoree2, ..., Enrichment_Score10
FB_sub <- FB[, FB$clusters_subset %in% c("7", "3", "18", "14", "5", "9", "19", "2", "12", "4", "13", "11", "17", "8", "1", "6", "15", "20")]
selected_clusters <- c("7", "3", "18", "14", "5", "9", "19", "2", "12", "4", "13", "11", "17", "8", "1", "6", "15", "20")
FB$clusters_subset <- factor(FB_sub$clusters_subset, levels = selected_clusters)

# Plot
pdf(file = paste0(myfolder, "ST_enrichment/Violin_enrichment_CVP_FVP_FB.pdf"), width = 10, height = 3)
VlnPlot(FB_sub, features = new_column_names, group.by = "clusters_subset", pt.size = 0) + 
  plot_layout(ncol = 2)
dev.off()
```


### For IN

```{r}
# Compute module score on single-cell data
IN <- AddModuleScore(IN, features = GeneSets, name = "Enrichment_Score")

# Rename the Enrichment_Score meta.data
# Create new_column_names using the list of numbers
new_column_names <- c("CVP", "FVP")

# Use a loop to rename the columns based on the list of indices and names
i_start <- as.numeric(dim(IN@meta.data)[2]-length(GeneSets)+1)
i_end <- as.numeric(dim(IN@meta.data)[2])
i_column <- as.numeric(dim(IN@meta.data)[2]-length(GeneSets))

for (i in i_start:i_end) {
  colnames(IN@meta.data)[i] <- new_column_names[i - i_column]
}

# Plot violin plots
# If you have 10 clusters, the scores should now be named Enrichment_Score1, Enrichment_Scoree2, ..., Enrichment_Score10
IN_sub <- IN[, IN$clusters_subset %in% c("4", "7", "12", "11", "8", "19", "17", "3", "13", "1")]
selected_clusters <- c("4", "7", "12", "11", "8", "19", "17", "3", "13", "1")
IN$clusters_subset <- factor(IN_sub$clusters_subset, levels = selected_clusters)

# Plot
pdf(file = paste0(myfolder, "ST_enrichment/Violin_enrichment_CVP_FVP_IN.pdf"), width = 10, height = 3)
VlnPlot(IN_sub, features = new_column_names, group.by = "clusters_subset", pt.size = 0) + 
  plot_layout(ncol = 2)
dev.off()
```




### For CM+EN+FB+IN

```{r}
# Subset CM clusters of interest (exclude:  11, 13, 15, 16, 20, 22)
subset_values <- c("1", "2", "3", "4-1", "4-2", "5", "6", "7", "8-1", "8-2", "8-3", "9", "10", "12", "14", "17", "18", "19-1", "19-2", "21")
CM_subset <- CM[, CM$clusters_subset_split %in% subset_values]
# reorder the cell types desired
CM_subset$clusters_subset_split <- factor(CM_subset$clusters_subset_split, levels = c("6", "1", "10", "8-1", "8-3", "17", "8-2", "18", "4-2", "4-1", "3", "14", "9", "12", "21", "19-2", "19-1", "7", "2", "5"))
rm(CM)
gc()

# Subset EN clusters of interest (exclude:  11, 13, 15, 16, 20, 22)
subset_values <- c("1", "12", "3", "6", "18", "17", "7", "19", "15", "13", "10", "2", "5", "21", "16", "20", "11", "14", "22")
EN_subset <- EN[, EN$clusters_subset %in% subset_values]
# reorder the cell types desired
EN_subset$clusters_subset <- factor(EN_subset$clusters_subset, levels = c("1", "12", "3", "6", "18", "17", "7", "19", "15", "13", "10", "2", "5", "21", "16", "20", "11", "14", "22"))
rm(EN)
gc()

# Subset FB clusters of interest
subset_values <- c("7", "3", "18", "14", "5", "9", "19", "2", "12", "4", "13", "11", "17", "8", "1", "6", "15", "20", "21")
FB_subset <- FB[, FB$clusters_subset %in% subset_values]

# reorder the cell types desired
FB_subset$clusters_subset <- factor(FB_subset$clusters_subset, levels = c("7", "3", "18", "14", "5", "9", "19", "2", "12", "4", "13", "11", "17", "8", "1", "6", "15", "20", "21"))
rm(FB)
gc()

# Subset IN clusters of interest
subset_values <- c("4", "7", "12", "11", "8", "19", "17", "3", "13", "1")
IN_subset <- IN[, IN$clusters_subset %in% subset_values]

# reorder the cell types desired
IN_subset$clusters_subset <- factor(IN_subset$clusters_subset, levels = c("4", "7", "12", "11", "8", "19", "17", "3", "13", "1"))
rm(IN)
gc()

# Subset HL clsuters of interest
subset_values <- c("6", "8", "10", "15", "18_1", "20", "22", "25", "26", "29", "34")
HL_subset <- data[, data$high_level_clusters %in% subset_values]

# reorder the cell types desired
HL_subset$high_level_clusters <- factor(HL_subset$high_level_clusters, levels = c("6", "8", "10", "15", "18_1", "20", "22", "25", "26", "29", "34"))
rm(data)
gc()

# Create a similarly named meta.data row for each object
CM_subset$final_clusters <- paste0("CM_", CM_subset$clusters_subset_split)
EN_subset$final_clusters <- paste0("EN_", EN_subset$clusters_subset)
FB_subset$final_clusters <- paste0("FB_", FB_subset$clusters_subset)
IN_subset$final_clusters <- paste0("IN_", IN_subset$clusters_subset)
HL_subset$final_clusters <- paste0("HL_", HL_subset$high_level_clusters)


# Merge the objects and keep the factor level of each of the object
Merged_object <- merge(CM_subset, y = c(EN_subset, FB_subset, IN_subset, HL_subset))

Merged_object$final_clusters <- factor(Merged_object$final_clusters, 
                               levels = c(paste0("CM_", levels(CM_subset$clusters_subset_split)),
                                          paste0("EN_", levels(EN_subset$clusters_subset)),
                                          paste0("FB_", levels(FB_subset$clusters_subset)),
                                          paste0("IN_", levels(IN_subset$clusters_subset)),
                                          paste0("HL_", levels(HL_subset$high_level_clusters))))

rm(CM_subset, EN_subset, FB_subset, IN_subset, HL_subset, subset_values)
gc()
```


```{r}
# Compute module score on single-cell data
Merged_object <- AddModuleScore(Merged_object, features = gene_list, name = "Enrichment_Score")


# Get list name to change the column names in meta.data
# Rename the Enrichment_Score meta.data
#new_column_names <- c("CVP", "FVP")
new_column_names <- names(gene_list)

# Use a loop to rename the columns based on the list of indices and names
i_start <- as.numeric(dim(Merged_object@meta.data)[2]-length(gene_list)+1)
i_end <- as.numeric(dim(Merged_object@meta.data)[2])
i_column <- as.numeric(dim(Merged_object@meta.data)[2]-length(gene_list))

for (i in i_start:i_end) {
  colnames(Merged_object@meta.data)[i] <- new_column_names[i - i_column]
}
```


```{r}
# Plot patchwork for every pathology
pdf(file = paste0(myfolder, "ST_enrichment/Violin_enrichment_CVP_FVP_merged.pdf"), 
     width = 35, height = 3)
VlnPlot(Merged_object, features = new_column_names, group.by = "final_clusters", pt.size = 0) + 
  plot_layout(ncol = 2)
dev.off()

# Plot new figures for each pathology
for (condition in new_column_names) {
  p <- VlnPlot(Merged_object, features = condition, group.by = "final_clusters", pt.size = 0)
  pdf(file = paste0(myfolder, "ST_enrichment/Violin_enrichment_merged_gene_sets_", condition, ".pdf"), width = 20, height = 3)
  print(p + geom_hline(yintercept = 0) + theme(legend.position = "none"))
  dev.off()
}
```


### Heatmap visualisation

```{r}
mat <- as.data.frame(Merged_object$final_clusters)
# mat <- cbind(mat, Merged_object$CVP)
# mat <- cbind(mat, Merged_object$FVP)

for (i in i_start:i_end) {
  mat <- cbind(mat, Merged_object@meta.data[,i])
}

colnames(mat) <- c("clusters", new_column_names)

mat_averages <- mat %>%
  group_by(clusters) %>%
  summarize(across(where(is.numeric), ~ mean(.x, na.rm = TRUE), .names = "avg_{col}"))
print(mat_averages)

pal <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "Reds"))(25)

mat_averages_long <- mat_averages %>%
  pivot_longer(
    cols = starts_with("avg_"), # Use starts_with to select all columns that were averaged
    names_to = "variable",
    values_to = "value"
  )

#plot
p <- ggplot(mat_averages_long, aes(variable, clusters, fill = value)) + 
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  labs(fill = "Enrichment \n score")

pdf(file = paste0(myfolder, "ST_enrichment/Heatmap_enrichment_pathology_merged.pdf"), 
     width = 4, height = 13)
p
dev.off()
```


### Heatmap visualisation (annotated) (Supp. Fig. 8B)

```{r}
mat <- as.data.frame(Merged_object$final_clusters)
mat <- cbind(mat, Merged_object$cell.types)

for (i in i_start:i_end) {
  mat <- cbind(mat, Merged_object@meta.data[,i])
}

colnames(mat) <- c("clusters", "celltype", new_column_names)
dim(mat)

# Separate the numeric and non-numeric columns
numeric_cols <- mat %>% select(where(is.numeric))
non_numeric_cols <- mat %>% select(clusters, celltype)

# Combine numeric columns with clusters for grouping
mat_with_clusters <- cbind(non_numeric_cols, numeric_cols)

# Group by clusters and calculate averages for numeric columns
mat_averages <- mat_with_clusters %>%
  group_by(clusters) %>%
  summarize(across(where(is.numeric), ~ mean(.x, na.rm = TRUE), .names = "avg_{col}"))

# Combine the non-numeric columns back with the averages
# Keep the unique combinations of clusters and celltype
non_numeric_distinct <- non_numeric_cols %>% distinct(clusters, celltype)
mat_averages <- left_join(non_numeric_distinct, mat_averages, by = "clusters")

pal <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "Reds"))(25)

mat_averages_long <- mat_averages %>%
  pivot_longer(
    cols = starts_with("avg_"), # Use starts_with to select all columns that were averaged
    names_to = "variable",
    values_to = "value"
  ) %>%
  arrange(clusters) %>%
  mutate(celltype_ordered = factor(celltype, levels = unique(celltype[order(clusters)])))

# Create the heatmap plot
p <- ggplot(mat_averages_long, aes(variable, celltype_ordered, fill = value)) + 
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90),
        axis.title.y = element_blank()) +
  labs(fill = "Enrichment \n score")

pdf(file = paste0(myfolder, "ST_enrichment/Heatmap_enrichment_pathology_merged_annot.pdf"), 
     width = 4.5, height = 13)
p
dev.off()
```


## On coarse-grained clusters

### For data

```{r}
# Compute module score on single-cell data
data <- AddModuleScore(data, features = gene_list, name = "Enrichment_Score")


# Get list name to change the column names in meta.data
# Rename the Enrichment_Score meta.data
new_column_names <- names(gene_list)

# Use a loop to rename the columns based on the list of indices and names
i_start <- as.numeric(dim(data@meta.data)[2]-length(gene_list)+1)
i_end <- as.numeric(dim(data@meta.data)[2])
i_column <- as.numeric(dim(data@meta.data)[2]-length(gene_list))

for (i in i_start:i_end) {
  colnames(data@meta.data)[i] <- new_column_names[i - i_column]
}
```


### Heatmap visualisation (annotated) (Supp. Fig. 8A)

```{r}
mat <- as.data.frame(data$high_level_clusters_removed)
mat <- cbind(mat, data$cell.types)

for (i in i_start:i_end) {
  mat <- cbind(mat, data@meta.data[,i])
}

colnames(mat) <- c("clusters", "celltype", new_column_names)
dim(mat)

# Separate the numeric and non-numeric columns
numeric_cols <- mat %>% select(where(is.numeric))
non_numeric_cols <- mat %>% select(clusters, celltype)

# Combine numeric columns with clusters for grouping
mat_with_clusters <- cbind(non_numeric_cols, numeric_cols)

# Group by clusters and calculate averages for numeric columns
mat_averages <- mat_with_clusters %>%
  group_by(clusters) %>%
  summarize(across(where(is.numeric), ~ mean(.x, na.rm = TRUE), .names = "avg_{col}"))

# Combine the non-numeric columns back with the averages
# Keep the unique combinations of clusters and celltype
non_numeric_distinct <- non_numeric_cols %>% distinct(clusters, celltype)
mat_averages <- left_join(non_numeric_distinct, mat_averages, by = "clusters")

pal <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "Reds"))(25)

mat_averages_long <- mat_averages %>%
  pivot_longer(
    cols = starts_with("avg_"), # Use starts_with to select all columns that were averaged
    names_to = "variable",
    values_to = "value"
  ) %>%
  arrange(clusters) %>%
  mutate(celltype_ordered = factor(celltype, levels = unique(celltype[order(clusters)])))

# Create the heatmap plot
p <- ggplot(mat_averages_long, aes(variable, celltype_ordered, fill = value)) + 
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90),
        axis.title.y = element_blank()) +
  labs(fill = "Enrichment \n score")

pdf(file = paste0(myfolder, "ST_enrichment/Heatmap_enrichment_pathology_HL_annot.pdf"), 
     width = 4.5, height = 8)
p
dev.off()
```

================================================================================
================================================================================
================================================================================
================================================================================
================================================================================






