---
title: "ST_HDCA_alignment_metrics"
author: "Raphaël Mauron"
output: html_document
editor_options: 
  chunk_output_type: console
---

# HDCA heart ST alignement metrics

This script presents the process for computing non-negative matrix factorisation.

NOTE: The path to load the data and the path to save the figures should be changed at your convenience.

Environment: **Conda: r-semla**

## Set up and loading material

### Knit setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Load library

```{r}
library(semla)
library(tidyr)
library(RcppML)
library(ggplot2)
library(openxlsx)
library(RcppML)
library(singlet)
library(tidyverse)
```


### Set working directory

```{r}
myfolder <- "~/hdca_DevHeart/output/HDCA_heart_sc_analysis_docker/ST_alignment_metrics/"
```


### Load `Seurat` object with ST data

```{r}
ST <- readRDS("~/hdca_DevHeart/data/ST/220114_STdata_harmony_semla.rds")
ST_clustering <- readRDS("~/hdca_DevHeart/data/ST/HDCA_heart_clustering.rds")
```


### Update image paths

```{r}
image_files <- list.files(path = "~/hdca_DevHeart/data/ST", 
                          pattern = "tissue_hires_image.png", 
                          full.names = TRUE, 
                          recursive = TRUE)
ST@tools$Staffli@imgs <- image_files
```


### Load H&E images

```{r}
ST <- LoadImages(ST, image_height = 1.5e3)
```


### Add section ID's to the ST object

```{r}
ST$sample_id <- paste0("section_", GetCoordinates(ST)$sampleID)
```


## Age specific cell type co-localization

```{r}
ST_meta_data <- openxlsx::read.xlsx("~/hdca_DevHeart/data/ST/HDCA_heart_ST_sections_overview_chambers_present.xlsx")
ST_meta_data$age <- as.numeric(gsub("[^0-9.]", "", ST_meta_data$Age))

ST_meta_data <- select(ST_meta_data, Section, age)
ST_meta_data$sample_id <- paste0("section_", ST_meta_data$Section)

ST_meta_data$age_groups <- cut(
  ST_meta_data$age,
  breaks = c(-Inf, 8, 10, Inf),
  labels = c("w6-w7", "w8-w9", "w10-w12"),
  right = FALSE
)

ST_meta_data
```


### Re-order sample by setting factors

```{r}
ST$sample_id <- factor(ST$sample_id, levels = unique(ST_meta_data$sample_id))
```


### Age group 

```{r}
#append age groups to ST_clustering
age_groups <- setNames(ST_meta_data$age_groups, nm = ST_meta_data$sample_id)
ST_clustering$age_groups <- age_groups[as.character(ST_clustering$sample_id)]

table(ST_clustering$sample_id, ST_clustering$age_groups)

# Create groups based on both clusters and ages and save to the object as a factor.
groups <- paste0(ST_clustering$seurat_clusters_annot, '_', ST_clustering$age_groups)
ST_clustering$subgroups <- factor(groups, levels = mixedsort(unique(groups)))

table(ST_clustering$sample_id, ST_clustering$subgroups)
table(ST_clustering$subgroups)

write.csv2(table(ST_clustering$subgroups), "~/hdca_DevHeart/output/HDCA_heart_sc_analysis_docker/Temporal_ST_clusters.csv")
```


### Temporal DEG on ST clusters

```{r}
DefaultAssay(ST_clustering) <- "SCT"

# reorder the cell types desired
ST_clustering$subgroups <- factor(ST_clustering$subgroups, levels = sort(unique(ST_clustering$subgroups)))

ST_clustering <- SetIdent(ST_clustering, value = "subgroups")

#ST_clustering <- FindVariableFeatures(ST_clustering, nfeatures = 4000)
table(ST_clustering$subgroups)

# The logfc threshold is set to 0.1 as compared to the default 0.25, which makes the detection of weaker signals possible (but it takes more computational power).
ST_clustering <- PrepSCTFindMarkers(ST_clustering)

detable <- FindAllMarkers(ST_clustering,
                          only.pos = TRUE,
                          max.cells.per.ident = 300,
                          logfc.threshold = .1,
                          assay = "SCT",
                          min.pct = 0.05)

# Only keep DEGs that have a p-value below 0.05.
detable <- detable[detable$p_val < 0.05, ]

# Adding two extra columns giving the difference (and log difference) between pct1 and pct2.
# pct1 is the percentage of cells in the specific cluster that express this gene.
# pct2 is the percentage of cells in all other clusters that express this gene.
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2((detable$pct.1*99+1) / (detable$pct.2*99+1))

# Save the detable as a csv file.
#write.csv2(detable, "~/hdca_DevHeart/output/HDCA_heart_sc_analysis_docker/banksy/detable_res_09.csv")


# Pick the 40 genes with the lowest p value for each group, then 20 highest based on pct.diff, then 5 highest based on avg log2fc.
detable %>%
  group_by(cluster) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(5, avg_log2FC) ->
  top5

pdf(paste0("~/hdca_DevHeart/output/HDCA_heart_sc_analysis_docker/Temporal_DEG_top5_ST_clusters.pdf"), width = 20, height = 35)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(ST_clustering, features = top5 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))
dev.off()

rm(detable)
```


## Group technica replicates

```{r}
replicates_num <- list(c("5", "6"), 
                   c("7", "8"),
                   c("13", "14", "15"),
                   c("35", "36"),
                   c("37", "38"),
                   c("9", "10", "11", "12"),
                   c("16", "18", "19"),
                   c("17", "20", "21"),
                   c("22", "23"),
                   c("24"),
                   c("33", "34"),
                   c("1", "2"),
                   c("3", "4"),
                   c("25", "26", "30"),
                   c("27", "28", "29"),
                   c("31", "32"))
replicates_num
```


### Load deconvolution results

```{r}
# HL results
stereoscope <- read.table("~/hdca_DevHeart/data/ST/stereoscope/W.2023-09-25133054.715121_hl.tsv", check.names = FALSE)
annotation_HL <- read.csv2("~/hdca_DevHeart/data/metadata/HDCA_heart_SC_annotations_HL_240115.csv", header = 1, sep = ";")
# Create a named vector for renaming columns
rename_vector <- setNames(annotation_HL$cell_type, annotation_HL$cluster)

# Rename the columns of the stereoscope dataframe
stereoscope <- stereoscope %>%
  rename_with(~ rename_vector[as.character(.)], everything())
# Find shares genes
all(colnames(ST) == rownames(stereoscope))

stereoscope.assay <- CreateAssayObject(data = stereoscope %>% t())
ST[["stereoscope"]] <- stereoscope.assay

DefaultAssay(ST) <- "stereoscope"
```

```{r}
#check namse of ST objects
rownames(ST)
```


### Comparison between replicates

#### Cell type composition per sections (38 section)

```{r}
# Extract sample identifiers from the barcodes
split_stereoscope <- split(stereoscope, as.integer(str_extract(rownames(stereoscope), "\\d+$")))

celltype_df <- t(split_stereoscope[["1"]]) #like celltype_df

# Create an empty dataframe with appropriate column names
results <- data.frame(matrix(nrow = nrow(celltype_df), ncol = length(split_stereoscope)))
colnames(results) <- names(split_stereoscope)
rownames(results) <- rownames(celltype_df)

# Populate the dataframe with 0 values
results[] <- 0

for (section in names(split_stereoscope)) {
  print(section)

  filtered_data <- t(split_stereoscope[[section]])

  # Get the average celltype composition for that subset. (divide by number of barcode to get a total score of 1)
  results[[section]] <- rowSums(filtered_data) / dim(filtered_data)[2]
}

results <- results |>
  rownames_to_column(var = "Cell_Type")

write.csv2(results, paste0(myfolder, "celltype_comp_sections.csv"))
```

```{r}
# Convert data from wide to long format for plotting
data_long <- tidyr::gather(results, key = "Section", value = "Proportion", -Cell_Type)
n <- length(results$Cell_Type)

assign_group <- function(section, replicates) {
for (i in seq_along(replicates)) {
  if (as.character(section) %in% replicates[[i]]) {
    return(i)
  }
}
return(NA)
}

data_long$Group <- sapply(data_long$Section, assign_group, replicates = replicates_num)

# Check the modified DataFrame
print(data_long)

desired_order <- c(5,6,7,8,13,14,15,35,36,37,38,1,2,27,28,29,31,32,3,4,25,26,30,9,10,11,12,33,34,16,18,19,22,23,17,20,21,24)
data_long$Section <- factor(data_long$Section, levels = desired_order)


# custom color
set.seed(42)
#color_palette <- c("#9E3D22", "#B5153A", "#27763D", "#7C4D79", "#D9452C", "#209d56", "#f3c300", "#875692", "#417aaf", "#008856", "#e68fac", "#fb6843", "#604e97", "#f6a600", "#d13770", "#a93214", "#8db600", "#e25822", "#1f77b4", "#ff8e2b", "#2ca02c", "#9467bd", "#7b7b7b", "#38cca5", "#17becf", "#00439a", "#ffbb78", "#98df8a",  "#9edae5", "#393b79", "#637939", "#7b4173", "#5254a3", "#a55194", "#b5cf6b", "#ffcc00", "#32a852", "#ff5733", "#ff8c00", "#9acd32")
color_palette <- c("#DE7E18FF", "#875692", "#003C30FF", "#DFC27DFF", "#f6a600", "#8C510AFF", "#6AAC91FF", "#209d56", "#244800", "#01665EFF", "#B8E186FF", "#80CDC1FF", "#417aaf", "#D9452C", "#3bd9d4", "#01665EFF", "#DE77AEFF", "#1f77b4", "#35978FFF", "#A53010FF", "#C51B7DFF", "#92AA4CFF", "#e25822", "#8E0152FF", "#728653FF", "#20ae00", "#DFC27DFF", "#ffbb78", "#f3c300", "#ff8e2b", "#9E3D22", "#B5153A", "#27763D")


#colors <- sample(viridis(n))
colors <- sample(color_palette, size = n, replace = FALSE)

# Plot with custom colors
plts <- ggplot(data_long, aes(x = Section, y = Proportion, fill = Cell_Type)) +
  geom_bar(stat = "identity") +
  labs(title = "Stacked Boxplot of Cell Types Proportions by Section",
      x = NULL,
      y = "Cell Type proportion",
      fill = "Cell Type") +
  scale_fill_manual(values = colors) +
  theme_classic() +
  theme(panel.background = element_rect(fill = "white",
                                          colour = "white",
                                          size = 0,
                                          linetype = "solid"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(size = 12, color = "black"),
          axis.text.y = element_text(size = 11, color = "black"),
          aspect.ratio = 1)


pdf(paste0(myfolder, "celltype_comp_sections.pdf"), width = 20, height = 20)
plts
dev.off()
```


```{r}
# Plot with custom colors, grouping by Group
plts <- ggplot(data_long, aes(x = factor(Section), y = Proportion, fill = Cell_Type)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Group, scales = "free_x") +
  labs(title = "Stacked Boxplot of Cell Types Proportions by Section Groups",
       x = NULL,
       y = "Cell Type proportion",
       fill = "Cell Type") +
  scale_fill_manual(values = colors) +
  theme_classic() +
  theme(panel.background = element_rect(fill = "white",
                                        colour = "white",
                                        size = 0,
                                        linetype = "solid"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 11, color = "black"),
        aspect.ratio = 1)

pdf(paste0(myfolder, "celltype_comp_grouped.pdf"), width = 20, height = 20)
plts
dev.off()
```





## Group technica replicates (17)

```{r}
replicates_num_clust <- list(c("5", "6"), 
                             c("7", "8"),
                             
                             c("1", "2"),
                             c("3", "4"),
                             c("28", "29"),
                             c("31", "32"),
                             
                             c("10", "11"),
                             c("24"),
                             c("33", "34"))
replicates_num_clust
```


#### Cluster composition per sections (17 section)

```{r}
cluster_distribution <- table(ST_clustering$seurat_clusters_annot, ST_clustering$sample_id) |> as.matrix()
cluster_sum <- colSums(cluster_distribution)

# Create an empty dataframe with appropriate column names
cluster_props <- data.frame(matrix(nrow = nrow(cluster_distribution), ncol = ncol(cluster_distribution)))
colnames(cluster_props) <- colnames(cluster_distribution)
rownames(cluster_props) <- rownames(cluster_distribution)

# Populate the dataframe with 0 values
cluster_props[] <- 0


for (section in colnames(cluster_distribution)) {
  print(section)
  print(cluster_sum[[section]])
  
  section_sub <- cluster_distribution[, section]
  section_sub_props <- section_sub/cluster_sum[[section]] * 100
  
  cluster_props[[section]] <- section_sub_props
}

cluster_props
colSums(cluster_props)
```

```{r}
cluster_props <- cluster_props |>
  rownames_to_column(var = "Clusters")

write.csv2(cluster_props, paste0(myfolder, "cluster_comp_sections.csv"))

# Convert data from wide to long format for plotting
data_long <- tidyr::gather(cluster_props, key = "Section", value = "Proportion", -Clusters)
n <- length(unique(data_long$Clusters))

assign_group <- function(section, replicates) {
for (i in seq_along(replicates)) {
  if (as.character(section) %in% replicates[[i]]) {
    return(i)
  }
}
return(NA)
}

data_long <- data_long %>%
  mutate(Section = str_remove(Section, "section_"))


data_long$Group <- sapply(data_long$Section, assign_group, replicates = replicates_num_clust)

desired_order <- c(5, 6, 7, 8, 1, 2, 28, 29, 31, 32, 3, 4, 10, 11, 33, 34, 24)
data_long$Section <- factor(data_long$Section, levels = desired_order)


# custom color
set.seed(42)
#color_palette <- c("#9E3D22", "#B5153A", "#27763D", "#7C4D79", "#D9452C", "#209d56", "#f3c300", "#875692", "#417aaf", "#008856", "#e68fac", "#fb6843", "#604e97", "#f6a600", "#d13770", "#a93214", "#8db600", "#e25822", "#1f77b4", "#ff8e2b", "#2ca02c", "#9467bd", "#7b7b7b", "#38cca5", "#17becf", "#00439a", "#ffbb78", "#98df8a",  "#9edae5", "#393b79", "#637939", "#7b4173", "#5254a3", "#a55194", "#b5cf6b", "#ffcc00", "#32a852", "#ff5733", "#ff8c00", "#9acd32")
color_palette <- c("#DE7E18FF", "#875692", "#003C30FF", "#DFC27DFF", "#f6a600", "#8C510AFF", "#6AAC91FF", "#209d56", "#244800", "#01665EFF", "#B8E186FF", "#80CDC1FF", "#417aaf", "#D9452C", "#3bd9d4", "#01665EFF", "#DE77AEFF", "#1f77b4", "#35978FFF", "#A53010FF", "#C51B7DFF", "#92AA4CFF", "#e25822", "#8E0152FF", "#728653FF", "#20ae00", "#DFC27DFF", "#ffbb78", "#f3c300", "#ff8e2b", "#9E3D22", "#B5153A", "#27763D")


#colors <- sample(viridis(n))
colors <- sample(color_palette, size = n, replace = FALSE)

# Plot with custom colors
plts <- ggplot(data_long, aes(x = Section, y = Proportion, fill = Clusters)) +
  geom_bar(stat = "identity") +
  labs(title = "Stacked Boxplot of Clusters percentage by Section",
      x = NULL,
      y = "Clusters proportion",
      fill = "Clusters") +
  scale_fill_manual(values = colors) +
  theme_classic() +
  theme(panel.background = element_rect(fill = "white",
                                          colour = "white",
                                          size = 0,
                                          linetype = "solid"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(size = 12, color = "black"),
          axis.text.y = element_text(size = 11, color = "black"),
          aspect.ratio = 1)


pdf(paste0(myfolder, "cluster_comp_sections.pdf"), width = 20, height = 20)
plts
dev.off()
```

```{r}
# Plot with custom colors, grouping by Group
plts <- ggplot(data_long, aes(x = factor(Section), y = Proportion, fill = Clusters)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Group, scales = "free_x") +
  labs(title = "Stacked Boxplot of Cell Types Proportions by Section Groups",
       x = NULL,
       y = "Clusters proportion",
       fill = "Clusters") +
  scale_fill_manual(values = colors) +
  theme_classic() +
  theme(panel.background = element_rect(fill = "white",
                                        colour = "white",
                                        size = 0,
                                        linetype = "solid"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 11, color = "black"),
        aspect.ratio = 1)

pdf(paste0(myfolder, "cluster_comp_grouped.pdf"), width = 20, height = 20)
plts
dev.off()
```


#### Banksy regions composition per sections (17 section)

## Load BANKSY regions

```{r}
banksy_regions <- read.csv2("~/hdca_DevHeart/output/HDCA_heart_sc_analysis_docker/banksy/metadata.csv", header = 1, sep = ",") %>%
  mutate(X = sub("_section", "", X)) %>%
  select(c("X", "clust_Harmony_BANKSY_k50_res0.8", "clust_Harmony_BANKSY_k50_res0.9"))

banksy_regions$sample <- sub(".*_(\\d+)$", "\\1", banksy_regions$X)
split_data <- split(banksy_regions$X, banksy_regions$sample)
list_of_lists <- lapply(split_data, function(barcodes) {
  sub("_[0-9]+$", "", barcodes)
})
list_of_lists <- list_of_lists[order(as.numeric(names(list_of_lists)))]
```



### Annotate section with corresponding sampleID

```{r}
#resolution 0.9
regions <- setNames(banksy_regions$clust_Harmony_BANKSY_k50_res0.9, nm = banksy_regions$X)
ST$banksy_09 <- regions[as.character(colnames(ST))]
```

```{r}
# sanity check
table(ST$sample_id, ST$banksy_09)
```

```{r}
# annotation banksy regions
banksy_num <- c(7, 6, 1, 2, 4, 5, 11, 13, 9, 10, 3, 14, 8, 15, 12)
banksy_annot <- c("LA", "RA", "LV_C", "RV_C_1", "LV_T", "RV_T", "EP", "CV-LYV", "AVP-PM", "V", "OFT-GV", "ICGP", "B", "RV_C_2", "NA")

banksy_09_annot <- setNames(banksy_annot, nm = banksy_num)
ST$banksy_09_annot <- banksy_09_annot[as.character(ST$banksy_09)]
```


```{r}
region_distribution <- table(ST$banksy_09_annot, ST$sample_id) |> as.matrix()

region_distribution <- region_distribution[, colnames(region_distribution) %in% c("section_5", "section_6", "section_7", "section_8", "section_1", "section_2", "section_3", "section_4", "section_28", "section_29", "section_31", "section_32", "section_10", "section_11", "section_24", "section_33", "section_34")]


region_sum <- colSums(region_distribution)

# Create an empty dataframe with appropriate column names
region_props <- data.frame(matrix(nrow = nrow(region_distribution), ncol = ncol(region_distribution)))
colnames(region_props) <- colnames(region_distribution)
rownames(region_props) <- rownames(region_distribution)

# Populate the dataframe with 0 values
region_props[] <- 0


for (section in colnames(region_distribution)) {
  print(section)
  print(region_sum[[section]])
  
  section_sub <- region_distribution[, section]
  section_sub_props <- section_sub/region_sum[[section]] * 100
  
  region_props[[section]] <- section_sub_props
}

region_props
colSums(region_props)
```


```{r}
region_props <- region_props |>
  rownames_to_column(var = "Regions")

write.csv2(region_props, paste0(myfolder, "regions_banksy_sections.csv"))

# Convert data from wide to long format for plotting
data_long <- tidyr::gather(region_props, key = "Section", value = "Proportion", -Regions)
n <- length(unique(data_long$Regions))

assign_group <- function(section, replicates) {
for (i in seq_along(replicates)) {
  if (as.character(section) %in% replicates[[i]]) {
    return(i)
  }
}
return(NA)
}

data_long <- data_long %>%
  mutate(Section = str_remove(Section, "section_"))


data_long$Group <- sapply(data_long$Section, assign_group, replicates = replicates_num_clust)

desired_order <- c(5, 6, 7, 8, 1, 2, 28, 29, 31, 32, 3, 4, 10, 11, 33, 34, 24)
data_long$Section <- factor(data_long$Section, levels = desired_order)


# custom color
set.seed(42)
#color_palette <- c("#9E3D22", "#B5153A", "#27763D", "#7C4D79", "#D9452C", "#209d56", "#f3c300", "#875692", "#417aaf", "#008856", "#e68fac", "#fb6843", "#604e97", "#f6a600", "#d13770", "#a93214", "#8db600", "#e25822", "#1f77b4", "#ff8e2b", "#2ca02c", "#9467bd", "#7b7b7b", "#38cca5", "#17becf", "#00439a", "#ffbb78", "#98df8a",  "#9edae5", "#393b79", "#637939", "#7b4173", "#5254a3", "#a55194", "#b5cf6b", "#ffcc00", "#32a852", "#ff5733", "#ff8c00", "#9acd32")
color_palette <- c("#DE7E18FF", "#875692", "#003C30FF", "#DFC27DFF", "#f6a600", "#8C510AFF", "#6AAC91FF", "#209d56", "#244800", "#01665EFF", "#B8E186FF", "#80CDC1FF", "#417aaf", "#D9452C", "#3bd9d4", "#01665EFF", "#DE77AEFF", "#1f77b4", "#35978FFF", "#A53010FF", "#C51B7DFF", "#92AA4CFF", "#e25822", "#8E0152FF", "#728653FF", "#20ae00", "#DFC27DFF", "#ffbb78", "#f3c300", "#ff8e2b", "#9E3D22", "#B5153A", "#27763D")


#colors <- sample(viridis(n))
colors <- sample(color_palette, size = n, replace = FALSE)

# Plot with custom colors
plts <- ggplot(data_long, aes(x = Section, y = Proportion, fill = Regions)) +
  geom_bar(stat = "identity") +
  labs(title = "Stacked Boxplot of Regions percentage by Section",
      x = NULL,
      y = "Regions proportion",
      fill = "Regions") +
  scale_fill_manual(values = colors) +
  theme_classic() +
  theme(panel.background = element_rect(fill = "white",
                                          colour = "white",
                                          size = 0,
                                          linetype = "solid"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(size = 12, color = "black"),
          axis.text.y = element_text(size = 11, color = "black"),
          aspect.ratio = 1)


pdf(paste0(myfolder, "regions_banksy_comp_sections.pdf"), width = 20, height = 20)
plts
dev.off()
```

```{r}
# Plot with custom colors, grouping by Group
plts <- ggplot(data_long, aes(x = factor(Section), y = Proportion, fill = Regions)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Group, scales = "free_x") +
  labs(title = "Stacked Boxplot of Regions Proportions by Section Groups",
       x = NULL,
       y = "Regions proportion",
       fill = "Regions") +
  scale_fill_manual(values = colors) +
  theme_classic() +
  theme(panel.background = element_rect(fill = "white",
                                        colour = "white",
                                        size = 0,
                                        linetype = "solid"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 11, color = "black"),
        aspect.ratio = 1)

pdf(paste0(myfolder, "regions_banksy_comp_grouped.pdf"), width = 20, height = 20)
plts
dev.off()
```










