---
title: "ST_HDCA_banksy"
author: "Raphaël Mauron"
output: html_document
editor_options: 
  chunk_output_type: console
---

# HDCA heart ST BANKSY

NOTE: The path to load the data and the path to save the figures should be changed at your convenience.

Environment: **Nexus server**

We follow pipeline from https://prabhakarlab.github.io/Banksy/articles/multi-sample.html


## Set up and loading material

### Knit setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
.libPaths()
.libPaths("/home/raphael.mauron/R/x86_64-pc-linux-gnu-library/4.4/")

lapply(.libPaths(), list.files)
```


### Load library

```{r}
library(semla)
library(tidyr)
library(RcppML)
library(ggplot2)
library(openxlsx)
library(tidyverse)
library(SpatialExperiment)
library(Banksy)
library(Seurat)
library(SeuratData)
library(gridExtra)
library(pals)
library(dplyr)
library(SummarizedExperiment)
library(cowplot)
library(ggspavis)
library(scater)
library(data.table)
library(harmony)

library(SeuratWrappers)
# Kelly palette for visualization
mypal <- kelly()[-1]
```


### Set working directory

```{r}
myfolder <- "/srv/home/raphael.mauron/projects/HDCA_heart/results/"
```


### Load data

```{r}
section_list <- readRDS("/srv/home/raphael.mauron/projects/HDCA_heart/data/ST_section_list_processed_nmf.rds")
```


### Convert semla object to SpatialExperiment

adapted from: https://github.com/drighelli/SpatialExperiment/issues/115

```{r}
## Function
seurat_to_spe <- function(seu, sample_id, img_id) {
    ## Convert to SCE
    sce <- Seurat::as.SingleCellExperiment(seu)
    
    ## Extract spatial coordinates
    spatialCoords <- as.matrix(
        seu@tools$Staffli@meta_data[, c("pxl_col_in_fullres", "pxl_row_in_fullres")])
    
    ## Extract and process image data
    img <- SpatialExperiment::SpatialImage(
        x = as.raster(seu@tools$Staffli@imgs))
    
    # imgData <- DataFrame(
    #     sample_id = sample_id,
    #     image_id = img_id,
    #     data = I(list(img)),
    #     scaleFactor = seu@tools[["Staffli"]]@scalefactors[["tissue_lowres_scalef"]])
    
    # Convert to SpatialExperiment
    spe <- SpatialExperiment(
        assays = assays(sce),
        rowData = rowData(sce),
        colData = colData(sce),
        metadata = metadata(sce),
        reducedDims = reducedDims(sce),
        altExps = altExps(sce),
        sample_id = sample_id,
        spatialCoords = spatialCoords
        #imgData = imgData
    )
    
    # Trim the data
    imgData(spe) <- NULL
    assay(spe, "logcounts") <- NULL
    reducedDims(spe) <- NULL
    altExps(spe) <- NULL
    rowData(spe) <- NULL
    
    # indicate all spots are on the tissue
    spe$in_tissue <- 1
    spe$sample_id <- sample_id
    
    # add colnames
    colnames(spe) <- paste0(colnames(spe), "_", spe$sample_id)
    invisible(gc())
    
    # Return Spatial Experiment object
    spe
}

```


### Apply transformation

```{r}
spe_list <- lapply(names(section_list), function(i) {
    seurat_to_spe(seu = section_list[[i]], sample_id = i, img_id = i)
})
```

```{r}
# check that the nmf embeddings have been removed
#spe_list[[1]]@int_colData@listData[["altExps"]]@listData[["Spatial"]]@se@int_colData@listData[["reducedDims"]]
```


## Run BANKSY
Following the multi sample vignette: https://prabhakarlab.github.io/Banksy/articles/multi-sample.html


## Subset data

### From 38 sections to 17 selected sections (same kept for initial clustering (spe_list_17))

```{r}
section_17 <- c("section_1", "section_2", "section_3", "section_4", "section_5", "section_6", "section_7", "section_8", "section_10", "section_11", "section_24", "section_28", "section_29", "section_31", "section_32", "section_33", "section_34")

# Function to subset a single SpatialExperiment object
subset_spe <- function(spe, section_17) {
  # Ensure the object is a SpatialExperiment
  if (!"SpatialExperiment" %in% class(spe)) {
    stop("The object is not a SpatialExperiment")
  }
  # Access colData and subset based on sample_id
  col_data <- colData(spe)
  subset_indices <- col_data$sample_id %in% section_17
  
  # Return the subsetted SpatialExperiment object
  return(spe[, subset_indices])
}

# Apply the subsetting function to each SpatialExperiment object in the list
subsetted_spe_list <- lapply(spe_list, subset_spe, section_17 = section_17)

# Filter out any empty SpatialExperiment objects
spe_list_17 <- subsetted_spe_list[sapply(subsetted_spe_list, function(spe) ncol(spe) > 0)]

#change names of objects
spe_list_17 <- setNames(spe_list_17, nm = section_17)

rm(subsetted_spe_list)
gc()
```


```{r}
# 38 sections
spe <- do.call(cbind, spe_list)
unique(spe$sample_id)

#' Stagger spatial coordinates
locs <- spatialCoords(spe)
locs <- cbind(locs, sample_id = factor(spe$sample_id))
locs_dt <- data.table(locs)
colnames(locs_dt) <- c("sdimx", "sdimy", "group")
locs_dt[, sdimx := sdimx - min(sdimx), by = group]
global_max <- max(locs_dt$sdimx) * 1.5
locs_dt[, sdimx := sdimx + group * global_max]
locs <- as.matrix(locs_dt[, 1:2])
rownames(locs) <- colnames(spe)
spatialCoords(spe) <- locs

# 17 sections
spe_17 <- spe[, spe$sample_id %in% section_17]
sample_names <- unique(spe_17$sample_id)
```

```{r}
# check sample ids in new objects
unique(spe$sample_id)
unique(spe_17$sample_id)
```



#### 17


```{r}
#' Get HVGs
seu_17 <- as.Seurat(spe_17, data = NULL)
seu_17 <- FindVariableFeatures(seu_17, nfeatures = 2000)

#' Normalize data
scale_factor <- median(colSums(assay(spe_17, "counts")))
seu_17 <- NormalizeData(seu_17, scale.factor = scale_factor, normalization.method = "RC")

#' Add data to SpatialExperiment and subset to HVGs
aname <- "normcounts"
assay(spe_17, aname) <- GetAssayData(seu_17)
spe_17 <- spe_17[VariableFeatures(seu_17),]
```

```{r}
compute_agf <- TRUE #For characterising neighborhoods, BANKSY computes the weighted neighborhood mean (H_0) and the azimuthal Gabor filter (H_1)
k_geom <- 6 #first order neighbours per visium spot
spe_17 <- computeBanksy(spe_17, assay_name = aname, compute_agf = compute_agf, k_geom = k_geom)
```

```{r}
# lambda <- 0.2 #small lambda --> cell type detection
lambda <- 0.8 #high lambda --> region detection
npcs <- 20
use_agf <- TRUE
spe_17 <- runBanksyPCA(spe_17, use_agf = use_agf, lambda = lambda, npcs = npcs, seed = 1000)
```

```{r}
set.seed(1000)
harmony_embedding <- HarmonyMatrix(
    data_mat = reducedDim(spe_17, "PCA_M1_lam0.8"),
    meta_data = colData(spe_17),
    vars_use = c("sample_id"),
    do_pca = FALSE,
    max.iter.harmony = 20,
    verbose = FALSE
)
reducedDim(spe_17, "Harmony_BANKSY") <- harmony_embedding
```

```{r}
spe_17 <- runBanksyUMAP(spe_17, use_agf = TRUE, lambda = lambda, npcs = npcs)
spe_17 <- runBanksyUMAP(spe_17, dimred = "Harmony_BANKSY")
```

```{r}
plot_grid(
    plotReducedDim(spe_17, "UMAP_M1_lam0.8", 
                   point_size = 0.1,
                   point_alpha = 0.5,
                   color_by = "sample_id") +
        theme(legend.position = "none"),
    plotReducedDim(spe_17, "UMAP_Harmony_BANKSY", 
                   point_size = 0.1,
                   point_alpha = 0.5,
                   color_by = "sample_id") +
        theme(legend.title = element_blank()) +
        guides(colour = guide_legend(override.aes = list(size = 5, alpha = 1))),
    nrow = 1,
    rel_widths = c(1, 1.2)
)
```

```{r}
# resolution used for Leiden graph-based clustering
spe_17 <- clusterBanksy(spe_17, dimred = "Harmony_BANKSY", resolution = 0.5,  k_neighbors = 50, seed = 1000)
spe_17 <- clusterBanksy(spe_17, dimred = "Harmony_BANKSY", resolution = 0.6,  k_neighbors = 50, seed = 1000)
spe_17 <- clusterBanksy(spe_17, dimred = "Harmony_BANKSY", resolution = 0.7,  k_neighbors = 50, seed = 1000)
spe_17 <- clusterBanksy(spe_17, dimred = "Harmony_BANKSY", resolution = 0.8,  k_neighbors = 50, seed = 1000)
spe_17 <- clusterBanksy(spe_17, dimred = "Harmony_BANKSY", resolution = 0.9,  k_neighbors = 50, seed = 1000)
spe_17 <- clusterBanksy(spe_17, dimred = "Harmony_BANKSY", resolution = 1,  k_neighbors = 50, seed = 1000) #too high
spe_17 <- clusterBanksy(spe_17, dimred = "Harmony_BANKSY", resolution = 1.2,  k_neighbors = 50, seed = 1000)

spe_17 <- connectClusters(spe_17)

# check number of regions detected per resolutions
unique(spe_17$clust_Harmony_BANKSY_k50_res0.5)
unique(spe_17$clust_Harmony_BANKSY_k50_res0.6)
unique(spe_17$clust_Harmony_BANKSY_k50_res0.7)
unique(spe_17$clust_Harmony_BANKSY_k50_res0.8)
unique(spe_17$clust_Harmony_BANKSY_k50_res0.9)
unique(spe_17$clust_Harmony_BANKSY_k50_res1)
unique(spe_17$clust_Harmony_BANKSY_k50_res1.2)


# length
length(unique(spe_17$clust_Harmony_BANKSY_k50_res0.5))
length(unique(spe_17$clust_Harmony_BANKSY_k50_res0.6))
length(unique(spe_17$clust_Harmony_BANKSY_k50_res0.7))
length(unique(spe_17$clust_Harmony_BANKSY_k50_res0.8))
length(unique(spe_17$clust_Harmony_BANKSY_k50_res0.9))
length(unique(spe_17$clust_Harmony_BANKSY_k50_res1))
length(unique(spe_17$clust_Harmony_BANKSY_k50_res1.2))
```

```{r}
#select the resolution you want to plot regions from:
clusterNames(spe_17)
cnm <- clusterNames(spe_17)[4]

sample_names <- unique(spe_17$sample_id)
spatial_plots <- lapply(sample_names, function(snm) {
    x <- spe_17[, spe_17$sample_id == snm]
    df <- cbind.data.frame(clust=colData(x)[[cnm]], spatialCoords(x))
    ggplot(df, aes(x=sdimy, y=sdimx, col=clust)) +
        geom_point(size = 1) +
        scale_color_manual(values = pals::kelly()[-1]) +
        theme_classic() +
        theme(
            legend.position = "none",
            axis.text.x=element_blank(),
            axis.text.y=element_blank(),
            axis.ticks=element_blank(),
            axis.title.x=element_blank(),
            axis.title.y=element_blank()) +
        labs(title = snm) +
        coord_equal()
})

title <- ggdraw() + draw_label(cnm, size = 10, hjust = 0, x = 0.05)
plot_grid(title, plotlist = spatial_plots, ncol = 6, byrow = FALSE)
```


### Save object

```{r}
saveRDS(spe_17, file = paste0(myfolder, "spe_17_banksy.rds"))
```



### Export Banksy matrix

```{r}
# export metadata of SpatialExperiment object
metadata <- spe_17@colData
dim(metadata)

write.csv(metadata, file = paste0(myfolder, "metadata.csv"))
```



```{r}
sessionInfo()
sessionInfo() %>% capture.output(file=paste0("/home/raphael.mauron/banksy/session_info_banksy.txt"))
```

