---
title: "ST_HDCA_NMF"
author: "Raphaël Mauron"
output: html_document
editor_options: 
  chunk_output_type: console
---

# HDCA heart ST NMF

This script presents the process for computing non-negative matrix factorisation.

NOTE: The path to load the data and the path to save the figures should be changed at your convenience.

Environment: **Conda: r-semla**

## Set up and loading material

### Knit setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Load library

```{r}
library(semla)
library(tidyr)
library(RcppML)
library(ggplot2)
library(openxlsx)
library(RcppML)
library(singlet)
library(tidyverse)
```


### Set working directory

```{r}
myfolder <- "~/hdca_DevHeart/output/HDCA_heart_sc_analysis_docker/ST_nmf/"
```


### Load `Seurat` object with ST data

```{r}
ST <- readRDS("~/hdca_DevHeart/data/ST/220114_STdata_harmony_semla.rds")
ST_clustering <- readRDS("~/hdca_DevHeart/data/ST/HDCA_heart_clustering.rds")
```


### Update image paths

```{r}
image_files <- list.files(path = "~/hdca_DevHeart/data/ST", 
                          pattern = "tissue_hires_image.png", 
                          full.names = TRUE, 
                          recursive = TRUE)
ST@tools$Staffli@imgs <- image_files
```


### Load H&E images

```{r}
ST <- LoadImages(ST, image_height = 1.5e3)
```


### Add section ID's to the ST object

```{r}
ST$sample_id <- paste0("section_", GetCoordinates(ST)$sampleID)
```


## Age specific cell type co-localization

```{r}
ST_meta_data <- openxlsx::read.xlsx("~/hdca_DevHeart/data/ST/HDCA_heart_ST_sections_overview_chambers_present.xlsx")
ST_meta_data$age <- as.numeric(gsub("[^0-9.]", "", ST_meta_data$Age))

ST_meta_data <- select(ST_meta_data, Section, age)
ST_meta_data$sample_id <- paste0("section_", ST_meta_data$Section)
ST_meta_data
```


### Re-order sample by setting factors

```{r}
ST$sample_id <- factor(ST$sample_id, levels = unique(ST_meta_data$sample_id))
```


## NMF

### All sections together (not using this one)

#### Check if object okay for NMF

```{r}
object <- ST
```


```{r}
length(VariableFeatures(object))
```


#### Run NMF

```{r}
set.seed(42)
object <- RunNMF(object, assay = "Spatial") #run NMF
```


#### Save object

```{r}
saveRDS(object,
        file = paste0("~/hdca_DevHeart/output/HDCA_heart_sc_analysis_docker/", "rds_objects/", "ST_processed_nmf", ".rds"))
```

```{r, eval = FALSE}
object <- readRDS(paste0("~/hdca_DevHeart/output/HDCA_heart_sc_analysis_docker/", "rds_objects/", "ST_processed_nmf", ".rds"))
```


#### Check optimal number of factors k

```{r}
k <- ncol(object@reductions$nmf@feature.loadings)
k
```

```{r}
plot.dir <- paste0(c(myfolder, "nmf_sections", "k"), collapse = "/")
if(dir.exists(plot.dir)) {
} else {
  dir.create(plot.dir, recursive = TRUE)
}
```


#### Print factorization rank

```{r}
pdf(paste0(plot.dir, "/RankPlot_nmf", ".pdf"), width = 25, height = 12)
print(RankPlot(object))
dev.off()
```


#### Spatial visualization of nmf

```{r}
plot.dir <- paste0(c(myfolder, "nmf_sections", "nmf"), collapse = "/")
if(dir.exists(plot.dir)) {
} else {
  dir.create(plot.dir, recursive = TRUE)
}
```

```{r}
numbsection <- dim(object@tools$Staffli@image_info)[1]

for (section in 1:numbsection) {
  point_size <- 1.5
  p <- MapFeatures(object,
          section_number = section,
          label_by = "sample_id",
          features = paste0("NMF_", 1:k),
          override_plot_dims = FALSE,
          pt_size = point_size,
          colors = viridis::magma(n = 11, direction = -1)) &
      theme(plot.title = element_blank())
  pdf(paste0(plot.dir, "/", unique(object$sample_id)[section], "_nmf_", ".pdf"), width = 30, height = 40)
  print(p)
  dev.off()
}
```


#### Spatial visualization (1 nmf accross every sections)

```{r}
plot.dir <- paste0(c(myfolder, "nmf_across_all_sections"), collapse = "/")
if(dir.exists(plot.dir)) {
} else {
  dir.create(plot.dir, recursive = TRUE)
}
```


```{r}
for (nmf in 1:k) {
  point_size <- 1.5
  p <- MapFeatures(object,
          label_by = "sample_id",
          features = paste0("NMF_", nmf),
          override_plot_dims = FALSE,
          pt_size = point_size,
          colors = viridis::magma(n = 11, direction = -1)) &
      theme(plot.title = element_blank())
  pdf(paste0(plot.dir, "/", "nmf_", nmf, ".pdf"), width = 30, height = 40)
  print(p)
  dev.off()
}
```


#### Plot Feature loadings

```{r}
plot.dir <- paste0(c(myfolder, "nmf_sections", "features_loading"), collapse = "/")
if(dir.exists(plot.dir)) {
} else {
  dir.create(plot.dir, recursive = TRUE)
}
```

```{r}
for (factor in 1:k) {
  p <- PlotFeatureLoadings(object,
                      dims = factor,
                      reduction = "nmf",
                      nfeatures = 30,
                      mode = "dotplot",
                      fill = "darkmagenta",
                      pt_size = 3)
  pdf(paste0(plot.dir, "/factor_", factor, "_loading", ".pdf"), width = 7, height = 8)
  print(p)
  dev.off()
}
```


### All sections together // k = 20

#### Check if object okay for NMF

```{r}
object_20 <- ST
```


```{r}
length(VariableFeatures(object_20))
```


#### Run NMF

```{r}
set.seed(42)
object_20 <- RunNMF(object_20, k = 20, assay = "Spatial") # run NMF
```


#### Save object

```{r}
saveRDS(object_20,
        file = paste0("~/hdca_DevHeart/output/HDCA_heart_sc_analysis_docker/", "rds_objects/", "ST_processed_20_nmf", ".rds"))
```

```{r, eval = FALSE}
object_20 <- readRDS(paste0("~/hdca_DevHeart/output/HDCA_heart_sc_analysis_docker/", "rds_objects/", "ST_processed_20_nmf", ".rds"))
```


#### Check optimal number of factors k

```{r}
k <- ncol(object_20@reductions$nmf@feature.loadings)
k
```

```{r}
plot.dir <- paste0(c(myfolder, "20_nmf_sections", "k"), collapse = "/")
if(dir.exists(plot.dir)) {
} else {
  dir.create(plot.dir, recursive = TRUE)
}
```


#### Print factorization rank

```{r}
pdf(paste0(plot.dir, "/RankPlot_nmf", ".pdf"), width = 25, height = 12)
print(RankPlot(object_20))
dev.off()
```


#### Spatial visualization of nmf

```{r}
plot.dir <- paste0(c(myfolder, "20_nmf_sections_redo", "nmf"), collapse = "/")
if(dir.exists(plot.dir)) {
} else {
  dir.create(plot.dir, recursive = TRUE)
}
```

```{r}
numbsection <- dim(object_20@tools$Staffli@image_info)[1]

for (section in 1:numbsection) {
  point_size <- 1.5
  p <- MapFeatures(object_20,
          section_number = section,
          label_by = "sample_id",
          features = paste0("NMF_", 1:k),
          override_plot_dims = FALSE,
          pt_size = point_size,
          colors = viridis::magma(n = 11, direction = -1)) &
      theme(plot.title = element_blank())
  pdf(paste0(plot.dir, "/", unique(object_20$sample_id)[section], "_nmf_", ".pdf"), width = 30, height = 40)
  print(p)
  dev.off()
}
```


#### Spatial visualization (1 nmf accross every sections)

```{r}
plot.dir <- paste0(c(myfolder, "20_nmf_across_all_sections_redo"), collapse = "/")
if(dir.exists(plot.dir)) {
} else {
  dir.create(plot.dir, recursive = TRUE)
}
```


```{r}
for (nmf in 1:k) {
  point_size <- 1.5
  p <- MapFeatures(object_20,
          label_by = "sample_id",
          features = paste0("NMF_", nmf),
          override_plot_dims = FALSE,
          pt_size = point_size,
          colors = viridis::magma(n = 11, direction = -1)) &
      theme(plot.title = element_blank())
  pdf(paste0(plot.dir, "/", "nmf_", nmf, ".pdf"), width = 30, height = 40)
  print(p)
  dev.off()
}
```


#### Plot Feature loadings

```{r}
plot.dir <- paste0(c(myfolder, "20_nmf_sections_redo", "features_loading"), collapse = "/")
if(dir.exists(plot.dir)) {
} else {
  dir.create(plot.dir, recursive = TRUE)
}
```

```{r}
for (factor in 1:k) {
  p <- PlotFeatureLoadings(object_20,
                      dims = factor,
                      reduction = "nmf",
                      nfeatures = 30,
                      mode = "dotplot",
                      fill = "darkmagenta",
                      pt_size = 3)
  pdf(paste0(plot.dir, "/factor_", factor, "_loading", ".pdf"), width = 7, height = 8)
  print(p)
  dev.off()
}
```


================================================================================
================================================================================
================================================================================
================================================================================
================================================================================










