---
title: "ST_HDCA_alignment_metrics"
author: "Raphaël Mauron"
output: html_document
editor_options: 
  chunk_output_type: console
---

# HDCA heart ST spatial ccc

This script presents the process for computing non-negative matrix factorisation.

NOTE: The path to load the data and the path to save the figures should be changed at your convenience.

Environment: **Conda: r-semla**

## Set up and loading material

### Knit setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Load library

```{r}
library(semla)
library(tidyr)
library(RcppML)
library(ggplot2)
library(openxlsx)
library(RcppML)
library(singlet)
library(tidyverse)
library(viridis)
library(pheatmap)
library(tibble)
library(colorjam)
library(patchwork)
library(plotly)
library(kableExtra)
```


### Set working directory

```{r}
myfolder <- "~/hdca_DevHeart/output/HDCA_heart_sc_analysis_docker/ST_spatial_ccc/"
```


### Load `Seurat` object with ST data

```{r}
ST <- readRDS("~/hdca_DevHeart/data/ST/220114_STdata_harmony_semla.rds")
ST_clustering <- readRDS("~/hdca_DevHeart/data/ST/HDCA_heart_clustering.rds")
```


### Update image paths

```{r}
image_files <- list.files(path = "~/hdca_DevHeart/data/ST", 
                          pattern = "tissue_hires_image.png", 
                          full.names = TRUE, 
                          recursive = TRUE)
ST@tools$Staffli@imgs <- image_files
```


### Load H&E images

```{r}
ST <- LoadImages(ST, image_height = 1.5e3)
```


### Add section ID's to the ST object

```{r}
ST$sample_id <- paste0("section_", GetCoordinates(ST)$sampleID)
```


## Age specific cell type co-localization

```{r}
ST_meta_data <- openxlsx::read.xlsx("~/hdca_DevHeart/data/ST/HDCA_heart_ST_sections_overview_chambers_present.xlsx")
ST_meta_data$age <- as.numeric(gsub("[^0-9.]", "", ST_meta_data$Age))

ST_meta_data <- select(ST_meta_data, Section, age)
ST_meta_data$sample_id <- paste0("section_", ST_meta_data$Section)

ST_meta_data$age_groups <- cut(
  ST_meta_data$age,
  breaks = c(-Inf, 8, 10, Inf),
  labels = c("w6-w7", "w8-w9", "w10-w12"),
  right = FALSE
)

ST_meta_data
```


### Re-order sample by setting factors

```{r}
ST$sample_id <- factor(ST$sample_id, levels = unique(ST_meta_data$sample_id))
```


## Spatial CCC

```{r}
DefaultAssay(ST) <- "SCT"
MapFeatures(ST,
            features = "MYH6",
            slot = "data", #data
            pt_size = 1) +
    patchwork::plot_layout(guides = "collect") &
    theme(plot.subtitle = element_blank()) &
    ThemeLegendRight()
```


```{r}
# parameters
ligand_receptor_pairs <- list(
  c("NPPA", "NPR1"),
  c("NPPA", "NPR3"),
  c("NPPB", "NPR1"),
  c("NPPB", "NPR2"),
  c("NPPB", "NPR3"),
  c("NPPC", "NPR2"),
  c("NPPC", "NPR3"),
  c("OSTN", "NPR3"),
  c("SEMA3C", "PLXNA2"),
  c("SEMA3C", "PLXNA4"),
  c("SEMA3D", "PLXNA2"),
  c("SEMA3D", "PLXNA4"),
  c("SEMA6A", "PLXNA2"),
  c("SEMA6A", "PLXNA4"),
  c("SEMA6B", "PLXNA2"),
  c("SEMA6B", "PLXNA4"),
  c("SEMA3A", "PLXNA4"),
  c("SEMA3A", "NRP1"))
section_number_list <- c(5, 2, 11, 24)
object <- ST
pt_size <- 1
#features <- c()

#loop through all the pairs and all the section required
for (features in ligand_receptor_pairs) {
  print(features)
  for (section_number in section_number_list) {
    print(section_number)
    
    # the function
    exprVals <- FetchData(ST, vars = features, layer = "data")
    exprVals_rescaled <- exprVals |> as.data.frame()
    
    # Calculate the product of cell state proportion 1 and cell state proportion 2 for each barcode
    exprVals_rescaled$product <- apply(exprVals_rescaled[, 1:length(features)], 1, prod)
    
    # Calculate the geometric mean (in R= exp(mean(log(x))))
    exprVals_rescaled$geom_mean <- exprVals_rescaled$product ** (1 / length(features))
    
    # Colocalisation score is the geom_mean multiplied by the number of events in the calculation
    exprVals_rescaled$coloc_score <- exprVals_rescaled$geom_mean * length(features)
    
    #Also calculate the sum of the events on the spot. Get an idea of the proportion of the spot is taken by those specific events
    exprVals_rescaled$sum <- apply(exprVals_rescaled[, 1:length(features)], 1, sum)
    
    # Fetch the coordinate of each read to plot in 2D, subset to coordinate corresponding to the sectionID selected
    coor <- GetCoordinates(object)
    coor2 <- coor[coor$sampleID == section_number, ]
    rm(coor)
    
    # Subset exprVals_rescaled
    exprVals_rescaled <- exprVals_rescaled[rownames(exprVals_rescaled) %in% coor2$barcode, ]
    
    # Create colors for coloc_score
    coloc_score_normalized <- scales::rescale(exprVals_rescaled$coloc_score, to = c(0, 1))
    
    # Generate colors using the normalized coloc_score values
    cols <- scales::gradient_n_pal(colours = viridis::turbo(n = 11))(coloc_score_normalized)
    #cols <- scales::gradient_n_pal(colours = viridis::turbo(n = 11))(exprVals_rescaled$coloc_score)
    exprVals_rescaled$cols <- cols
    
    
    # Create ggplot object to plot
    gg <- bind_cols(exprVals_rescaled, coor2)
    
    # Get maximum dimension from section_number to plot
    full_width <- object@tools[["Staffli"]]@image_info[["full_width"]][section_number]
    full_height <- object@tools[["Staffli"]]@image_info[["full_height"]][section_number]
    
    score <- gg$coloc_score
    
    # Plot coloc_score in 2D
    plts <- ggplot(gg, aes(pxl_col_in_fullres, pxl_row_in_fullres)) +
      geom_point(color = cols, size = pt_size, aes(fill = score), show.legend = TRUE) +
    
      scale_fill_gradientn(colours = viridis::turbo(n = 11), #use color
        breaks = c(round(min(score, na.rm = TRUE), digits = 2),
                   round((max(score, na.rm = TRUE) * 4 / 4), digits = 2),
                   round((max(score, na.rm = TRUE) * 3 / 4), digits = 2), #add 4 ticks based on the respective cell_proportion
                   round((max(score, na.rm = TRUE) * 2 / 4), digits = 2),
                   round((max(score, na.rm = TRUE) * 1 / 4), digits = 2)),
        limits = c(round(min(score, na.rm = TRUE), digits = 2),
                   round(max(score, na.rm = TRUE), digits = 2))
      ) +
    
      scale_y_reverse(limits = c(full_height, 0)) +
      scale_x_continuous(limits = c(0, full_width)) +
    
      labs(x = NULL, y = NULL) +
    
      theme(panel.background = element_rect(fill = "white",
                                            colour = "white",
                                            size = 0,
                                            linetype = "solid"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            axis.text = NULL,
            axis.text.x = element_blank(),
            axis.text.y = element_blank(),
            axis.ticks = element_blank(),
            aspect.ratio = 1) +
    
      ggtitle(label = paste(features, collapse = " , "),
              subtitle = "Colocalisation")
    
    #plts
    
    pdf(file = paste0(myfolder, features[1], "_", features[2], "_section_", section_number, ".pdf"), width = 6, height = 6)
    print(plts)
    dev.off()
    
  }
}
```


## The code

```{r}
exprVals <- FetchData(ST, vars = features, layer = "data")
exprVals_rescaled <- exprVals |> as.data.frame()

# Calculate the product of cell state proportion 1 and cell state proportion 2 for each barcode
exprVals_rescaled$product <- apply(exprVals_rescaled[, 1:length(features)], 1, prod)

# Calculate the geometric mean (in R= exp(mean(log(x))))
exprVals_rescaled$geom_mean <- exprVals_rescaled$product ** (1 / length(features))

# Colocalisation score is the geom_mean multiplied by the number of events in the calculation
exprVals_rescaled$coloc_score <- exprVals_rescaled$geom_mean * length(features)

#Also calculate the sum of the events on the spot. Get an idea of the proportion of the spot is taken by those specific events
exprVals_rescaled$sum <- apply(exprVals_rescaled[, 1:length(features)], 1, sum)

# Fetch the coordinate of each read to plot in 2D, subset to coordinate corresponding to the sectionID selected
coor <- GetCoordinates(object)
coor2 <- coor[coor$sampleID == section_number, ]
rm(coor)

# Subset exprVals_rescaled
exprVals_rescaled <- exprVals_rescaled[rownames(exprVals_rescaled) %in% coor2$barcode, ]

# Create colors for coloc_score
coloc_score_normalized <- scales::rescale(exprVals_rescaled$coloc_score, to = c(0, 1))

# Generate colors using the normalized coloc_score values
cols <- scales::gradient_n_pal(colours = viridis::turbo(n = 11))(coloc_score_normalized)
#cols <- scales::gradient_n_pal(colours = viridis::turbo(n = 11))(exprVals_rescaled$coloc_score)
exprVals_rescaled$cols <- cols


# Create ggplot object to plot
gg <- bind_cols(exprVals_rescaled, coor2)

# Get maximum dimension from section_number to plot
full_width <- object@tools[["Staffli"]]@image_info[["full_width"]][section_number]
full_height <- object@tools[["Staffli"]]@image_info[["full_height"]][section_number]

score <- gg$coloc_score

# Plot coloc_score in 2D
plts <- ggplot(gg, aes(pxl_col_in_fullres, pxl_row_in_fullres)) +
  geom_point(color = cols, size = pt_size, aes(fill = score), show.legend = TRUE) +

  scale_fill_gradientn(colours = viridis::turbo(n = 11), #use color
    breaks = c(round(min(score, na.rm = TRUE), digits = 2),
               round((max(score, na.rm = TRUE) * 4 / 4), digits = 2),
               round((max(score, na.rm = TRUE) * 3 / 4), digits = 2), #add 4 ticks based on the respective cell_proportion
               round((max(score, na.rm = TRUE) * 2 / 4), digits = 2),
               round((max(score, na.rm = TRUE) * 1 / 4), digits = 2)),
    limits = c(round(min(score, na.rm = TRUE), digits = 2),
               round(max(score, na.rm = TRUE), digits = 2))
  ) +

  scale_y_reverse(limits = c(full_height, 0)) +
  scale_x_continuous(limits = c(0, full_width)) +

  labs(x = NULL, y = NULL) +

  theme(panel.background = element_rect(fill = "white",
                                        colour = "white",
                                        size = 0,
                                        linetype = "solid"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = NULL,
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        aspect.ratio = 1) +

  ggtitle(label = paste(features, collapse = " , "),
          subtitle = "Colocalisation")

plts


```











