---
title: "10_HDCA_heart_reviewer_request"
author: "Raphaël Mauron"
output: html_document
editor_options: 
  chunk_output_type: console
---

# HDCA heart Reviewer requests
- Temporal DEG for single-cell
- DEG with ISS gene panel
- Subclustering analysis HL_20
- Pathway enrichment analysis
- CCC analysis from selection of LR pairs

NOTE: The path to load the data and the path to save the figures should be changed at your convenience.

Environment: **Docker**


## Settings

### Set library path

```{r setup, include=FALSE}
.libPaths("/home/rstudio/project/renv/library/R-4.3/aarch64-unknown-linux-gnu/")
knitr::opts_chunk$set(echo = TRUE)
```


### Set environment variables

```{r}
# Set the directory 
myfolder <- "/home/rstudio/project/output/HDCA_heart_sc_analysis_docker/"
```


### Load Libraries

```{r, message = FALSE}
library(Seurat)
library(Matrix)
library(future)
library(RcppHNSW)
library(igraph)
library(future.apply)
library(niceRplots)
library(DoubletFinder)
library(ggplot2)
library(ROCR)
library(KernSmooth)
library(dplyr)
library(patchwork)
library(harmony)
library(gprofiler2)
library(plotly)
library(gtools)
library(openxlsx)
library(dplyr)
library(tibble)
```


### Load global functions

```{r}
source("/home/rstudio/project/code/global_functions.R")
source("/home/rstudio/project/code/LR.R")
```


### Set seed

```{r}
set.seed(1)
```


## Load data

```{r}
data <- readRDS(paste0(myfolder, "rds_objects/HDCA_heart_sc_full_extended.rds"))
```


## Temporal analysis (single-cell)

### DGE (Extended Fig. 2B, 4B)

```{r}
# set seed for reproducibility when sampling
set.seed(1)

# Create groups based on both clusters and ages and save to the object as a factor.
#groups <- paste0(data$high_level_clusters_removed, '_', as.character(data$age_groups))
groups_annot <- paste0(data$cell.types, '_', as.character(data$age_groups))
sort(unique(groups_annot))

#data$subgroups <- factor(groups, levels = mixedsort(unique(groups)))
data$subgroups_annot <- factor(groups_annot, levels = mixedsort(unique(groups_annot)))
sort(unique(data$subgroups_annot))

# take only 150 cells per cluster to speed up calculation
sample_size <- table(data$subgroups_annot)
sample_size[ sample_size > 150 ] <- 150
DGE_cells <- lapply(names(sample_size), function(x){
  set.seed(1)
  sample( colnames(data) [ data$subgroups_annot == x ] , size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# subset based on previous step, removing NA & setting Identity to these groups (cluster_age)
DGE_DATA <- data[, DGE_cells]
DGE_DATA <- DGE_DATA[, !DGE_DATA$subgroups_annot %in% c("HL_excl_1_12-14", "HL_excl_1_5-6", "HL_excl_1_7-8", "HL_excl_1_9-11", "HL_excl_2_12-14", "HL_excl_2_5-6", "HL_excl_2_7-8", "HL_excl_2_9-11", "HL_excl_3_12-14", "HL_excl_3_5-6", "HL_excl_3_7-8", "HL_excl_3_9-11", "HL_excl_4_12-14", "HL_excl_4_5-6", "HL_excl_4_7-8", "HL_excl_4_9-11")]
sort(unique(DGE_DATA$subgroups_annot))

DGE_DATA$subgroups_annot <- factor(DGE_DATA$subgroups_annot, levels = c(
  "Mat_vCM_5-6", "Mat_vCM_7-8", "Mat_vCM_9-11", "Mat_vCM_12-14",
  "Mat_aCM_5-6", "Mat_aCM_7-8", "Mat_aCM_9-11", "Mat_aCM_12-14",
  "MetAct_vCM_1_5-6", "MetAct_vCM_1_7-8", "MetAct_vCM_1_9-11", "MetAct_vCM_1_12-14",
  "MetAct_vCM_2_5-6", "MetAct_vCM_2_7-8", "MetAct_vCM_2_9-11", "MetAct_vCM_2_12-14",
  "MetAct_aCM_5-6", "MetAct_aCM_7-8", "MetAct_aCM_9-11", "MetAct_aCM_12-14",
  "Prol_CM_5-6", "Prol_CM_7-8", "Prol_CM_9-11", "Prol_CM_12-14",
  "Immat_CM_5-6", "Immat_CM_7-8", "Immat_CM_9-11", "Immat_CM_12-14",
  "TMSB10high_C_1_5-6", "TMSB10high_C_1_7-8", "TMSB10high_C_1_9-11", "TMSB10high_C_1_12-14",
  "TMSB10high_C_2_5-6", "TMSB10high_C_2_7-8", "TMSB10high_C_2_9-11", "TMSB10high_C_2_12-14",
  "MacroVasc_EC_5-6", "MacroVasc_EC_7-8", "MacroVasc_EC_9-11", "MacroVasc_EC_12-14",
  "MicroVasc_EC_5-6", "MicroVasc_EC_7-8", "MicroVasc_EC_9-11", "MicroVasc_EC_12-14",
  "PDE4Chigh_EC_5-6", "PDE4Chigh_EC_7-8", "PDE4Chigh_EC_9-11", "PDE4Chigh_EC_12-14",
  "LEC_5-6", "LEC_7-8", "LEC_9-11", "LEC_12-14",
  "Endoc_EC_5-6", "Endoc_EC_7-8", "Endoc_EC_9-11", "Endoc_EC_12-14",
  "EndocCush_EC_5-6", "EndocCush_EC_7-8", "EndocCush_EC_9-11", "EndocCush_EC_12-14",
  "Valve_MC_5-6", "Valve_MC_7-8", "Valve_MC_9-11", "Valve_MC_12-14",
  "PDE4Chigh_FB_5-6", "PDE4Chigh_FB_7-8", "PDE4Chigh_FB_9-11", "PDE4Chigh_FB_12-14",
  "Int_FB_5-6", "Int_FB_7-8", "Int_FB_9-11", "Int_FB_12-14",
  "Peric_MC_5-6", "Peric_MC_7-8", "Peric_MC_9-11", "Peric_MC_12-14",
  "OFT_FB_5-6", "OFT_FB_7-8", "OFT_FB_9-11", "OFT_FB_12-14",
  "EPDC_5-6", "EPDC_7-8", "EPDC_9-11", "EPDC_12-14",
  "AnnFibr_FB_5-6", "AnnFibr_FB_7-8", "AnnFibr_FB_9-11", "AnnFibr_FB_12-14",
  "Prol_FB_5-6", "Prol_FB_7-8", "Prol_FB_9-11", "Prol_FB_12-14",
  "OFT_SMC_5-6", "OFT_SMC_7-8", "OFT_SMC_9-11", "OFT_SMC_12-14",
  "CA_SMC_5-6", "CA_SMC_7-8", "CA_SMC_9-11", "CA_SMC_12-14",
  "PC_5-6", "PC_7-8", "PC_9-11", "PC_12-14",
  "EpC_5-6", "EpC_7-8", "EpC_9-11", "EpC_12-14",
  "NB-N_5-6", "NB-N_7-8", "NB-N_9-11", "NB-N_12-14",
  "SCP-GC_5-6", "SCP-GC_7-8", "SCP-GC_9-11", "SCP-GC_12-14",
  "MyC_5-6", "MyC_7-8", "MyC_9-11", "MyC_12-14",
  "LyC_5-6", "LyC_7-8", "LyC_9-11", "LyC_12-14"
))
DGE_DATA <- SetIdent(DGE_DATA, value = DGE_DATA@meta.data$subgroups_annot)
sort(unique(DGE_DATA$subgroups_annot))

# find markers accross every category
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 150,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05,
                          return.thresh = 1)

# add metrics and save table
detable <- detable[ detable$p_val < 0.05,  ]
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2( (detable$pct.1+1) / (detable$pct.2+1) )
write.csv2(detable, paste0(myfolder, "detable_for_each_cluster_split_by_age.csv"))

# Save results as a dot plot
detable$groups <- paste0(detable$cell_cluster,detable$cluster)

detable %>%
  group_by(groups) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(5, avg_log2FC) ->
  top5

# save
pdf(paste0(myfolder, "DGE_temporal_v2_2.pdf"), width = 60, height = 100)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top5 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()
```


### DGE - order ISS genes & clusters

```{r}
# specific list of marker to project the dotplot on
iss_gene <- read.csv(file = paste0(myfolder, "supplementary_tables/extended_table_01_ISS_gene_list.csv"))
my_markers <- iss_gene$Gene.list

# set seed for reproducibility when sampling
set.seed(1)

# Create groups based on both clusters and ages and save to the object as a factor.
#groups <- paste0(data$high_level_clusters_removed, '_', as.character(data$age_groups))
groups_annot <- paste0(data$cell.types, '_', as.character(data$age_groups))
sort(unique(groups_annot))

#data$subgroups <- factor(groups, levels = mixedsort(unique(groups)))
data$subgroups_annot <- factor(groups_annot, levels = mixedsort(unique(groups_annot)))
sort(unique(data$subgroups_annot))

# take only 150 cells per cluster to speed up calculation
sample_size <- table(data$subgroups_annot)
sample_size[ sample_size > 150 ] <- 150
DGE_cells <- lapply(names(sample_size), function(x){
  set.seed(1)
  sample( colnames(data) [ data$subgroups_annot == x ] , size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# subset based on previous step, removing NA & setting Identity to these groups (cluster_age)
DGE_DATA <- data[, DGE_cells]
DGE_DATA <- DGE_DATA[, !DGE_DATA$subgroups_annot %in% c("HL_excl_1_12-14", "HL_excl_1_5-6", "HL_excl_1_7-8", "HL_excl_1_9-11", "HL_excl_2_12-14", "HL_excl_2_5-6", "HL_excl_2_7-8", "HL_excl_2_9-11", "HL_excl_3_12-14", "HL_excl_3_5-6", "HL_excl_3_7-8", "HL_excl_3_9-11", "HL_excl_4_12-14", "HL_excl_4_5-6", "HL_excl_4_7-8", "HL_excl_4_9-11")]
sort(unique(DGE_DATA$subgroups_annot))


DGE_DATA$subgroups_annot <- factor(DGE_DATA$subgroups_annot, levels = c(
  "Mat_vCM_5-6", "Mat_vCM_7-8", "Mat_vCM_9-11", "Mat_vCM_12-14",
  "Mat_aCM_5-6", "Mat_aCM_7-8", "Mat_aCM_9-11", "Mat_aCM_12-14",
  "MetAct_vCM_1_5-6", "MetAct_vCM_1_7-8", "MetAct_vCM_1_9-11", "MetAct_vCM_1_12-14",
  "MetAct_vCM_2_5-6", "MetAct_vCM_2_7-8", "MetAct_vCM_2_9-11", "MetAct_vCM_2_12-14",
  "MetAct_aCM_5-6", "MetAct_aCM_7-8", "MetAct_aCM_9-11", "MetAct_aCM_12-14",
  "Prol_CM_5-6", "Prol_CM_7-8", "Prol_CM_9-11", "Prol_CM_12-14",
  "Immat_CM_5-6", "Immat_CM_7-8", "Immat_CM_9-11", "Immat_CM_12-14",
  "TMSB10high_C_1_5-6", "TMSB10high_C_1_7-8", "TMSB10high_C_1_9-11", "TMSB10high_C_1_12-14",
  "TMSB10high_C_2_5-6", "TMSB10high_C_2_7-8", "TMSB10high_C_2_9-11", "TMSB10high_C_2_12-14",
  "MacroVasc_EC_5-6", "MacroVasc_EC_7-8", "MacroVasc_EC_9-11", "MacroVasc_EC_12-14",
  "MicroVasc_EC_5-6", "MicroVasc_EC_7-8", "MicroVasc_EC_9-11", "MicroVasc_EC_12-14",
  "PDE4Chigh_EC_5-6", "PDE4Chigh_EC_7-8", "PDE4Chigh_EC_9-11", "PDE4Chigh_EC_12-14",
  "LEC_5-6", "LEC_7-8", "LEC_9-11", "LEC_12-14",
  "Endoc_EC_5-6", "Endoc_EC_7-8", "Endoc_EC_9-11", "Endoc_EC_12-14",
  "EndocCush_EC_5-6", "EndocCush_EC_7-8", "EndocCush_EC_9-11", "EndocCush_EC_12-14",
  "Valve_MC_5-6", "Valve_MC_7-8", "Valve_MC_9-11", "Valve_MC_12-14",
  "PDE4Chigh_FB_5-6", "PDE4Chigh_FB_7-8", "PDE4Chigh_FB_9-11", "PDE4Chigh_FB_12-14",
  "Int_FB_5-6", "Int_FB_7-8", "Int_FB_9-11", "Int_FB_12-14",
  "Peric_MC_5-6", "Peric_MC_7-8", "Peric_MC_9-11", "Peric_MC_12-14",
  "OFT_FB_5-6", "OFT_FB_7-8", "OFT_FB_9-11", "OFT_FB_12-14",
  "EPDC_5-6", "EPDC_7-8", "EPDC_9-11", "EPDC_12-14",
  "AnnFibr_FB_5-6", "AnnFibr_FB_7-8", "AnnFibr_FB_9-11", "AnnFibr_FB_12-14",
  "Prol_FB_5-6", "Prol_FB_7-8", "Prol_FB_9-11", "Prol_FB_12-14",
  "OFT_SMC_5-6", "OFT_SMC_7-8", "OFT_SMC_9-11", "OFT_SMC_12-14",
  "CA_SMC_5-6", "CA_SMC_7-8", "CA_SMC_9-11", "CA_SMC_12-14",
  "PC_5-6", "PC_7-8", "PC_9-11", "PC_12-14",
  "EpC_5-6", "EpC_7-8", "EpC_9-11", "EpC_12-14",
  "NB-N_5-6", "NB-N_7-8", "NB-N_9-11", "NB-N_12-14",
  "SCP-GC_5-6", "SCP-GC_7-8", "SCP-GC_9-11", "SCP-GC_12-14",
  "MyC_5-6", "MyC_7-8", "MyC_9-11", "MyC_12-14",
  "LyC_5-6", "LyC_7-8", "LyC_9-11", "LyC_12-14"
))
DGE_DATA <- SetIdent(DGE_DATA, value = DGE_DATA@meta.data$subgroups_annot)
sort(unique(DGE_DATA$subgroups_annot))

p <- DotPlot(DGE_DATA, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DGE_temporal_ISS_markers.pdf"), width = 40, height = 30)
print(p)
dev.off()

rm(p)
gc()
```


### DGE - selection of HL clusters

```{r}
# set seed for reproducibility when sampling
set.seed(1)

# Create groups based on both clusters and ages and save to the object as a factor.
groups <- paste0(data$high_level_clusters_removed, '_', as.character(data$age_groups))
data$subgroups <- factor(groups, levels = mixedsort(unique(groups)))

# take only 150 cells per cluster to speed up calculation
sample_size <- table(data$subgroups)
sample_size[ sample_size > 150 ] <- 150
DGE_cells <- lapply(names(sample_size), function(x){
  set.seed(1)
  sample( colnames(data) [ data$subgroups == x ] , size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# subset based on previous step, removing NA & setting Identity to these groups (cluster_age)
DGE_DATA <- data[, DGE_cells]
DGE_DATA <- DGE_DATA[, DGE_DATA$subgroups %in% c("1_5-6", "1_7-8", "1_9-11", "1_12-14",
                                                 "8_5-6", "8_7-8", "8_9-11", "8_12-14",
                                                 "10_5-6", "10_7-8", "10_9-11", "10_12-14",
                                                 "13_5-6", "13_7-8", "13_9-11", "13_12-14",
                                                 "20_5-6", "20_7-8", "20_9-11", "20_12-14",
                                                 "22_5-6", "22_7-8", "22_9-11", "22_12-14"
                                                 )]
DGE_DATA <- SetIdent(DGE_DATA, value = DGE_DATA@meta.data$subgroups)

# find markers accross every category
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 150,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05,
                          return.thresh = 1)

# add metrics and save table
detable <- detable[ detable$p_val < 0.05,  ]
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2( (detable$pct.1+1) / (detable$pct.2+1) )
write.csv2(detable, paste0(myfolder, "detable_selection_HL_split_by_age.csv"))

# Save results as a dot plot
detable$groups <- paste0(detable$cell_cluster,detable$cluster)

detable %>%
  group_by(groups) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(10, avg_log2FC) ->
  top10

# save
pdf(paste0(myfolder, "DGE_temporal_selection_HL.pdf"), width = 15, height = 40)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top10 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()
```


### DGE - selection of HL clusters (HL_10)

```{r}
# set seed for reproducibility when sampling
set.seed(1)

# Create groups based on both clusters and ages and save to the object as a factor.
groups <- paste0(data$high_level_clusters_removed, '_', as.character(data$age_groups))
data$subgroups <- factor(groups, levels = mixedsort(unique(groups)))

# take only 150 cells per cluster to speed up calculation
sample_size <- table(data$subgroups)
sample_size[ sample_size > 150 ] <- 150
DGE_cells <- lapply(names(sample_size), function(x){
  set.seed(1)
  sample( colnames(data) [ data$subgroups == x ] , size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# subset based on previous step, removing NA & setting Identity to these groups (cluster_age)
DGE_DATA <- data[, DGE_cells]
DGE_DATA <- DGE_DATA[, DGE_DATA$subgroups %in% c("10_5-6", "10_7-8", "10_9-11", "10_12-14")]
DGE_DATA <- SetIdent(DGE_DATA, value = DGE_DATA@meta.data$subgroups)

# find markers accross every category
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 150,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05,
                          return.thresh = 1)

# add metrics and save table
detable <- detable[ detable$p_val < 0.05,  ]
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2( (detable$pct.1+1) / (detable$pct.2+1) )
#write.csv2(detable, paste0(myfolder, "detable_selection_HL_split_by_age.csv"))

# Save results as a dot plot
detable$groups <- paste0(detable$cell_cluster,detable$cluster)

detable %>%
  group_by(groups) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(10, avg_log2FC) ->
  top10

# save
pdf(paste0(myfolder, "DGE_temporal_selection_HL10.pdf"), width = 5, height = 10)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top10 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()
```


### DGE - selection of HL clusters (HL_8)

```{r}
# set seed for reproducibility when sampling
set.seed(1)

# Create groups based on both clusters and ages and save to the object as a factor.
groups <- paste0(data$high_level_clusters_removed, '_', as.character(data$age_groups))
data$subgroups <- factor(groups, levels = mixedsort(unique(groups)))

# take only 150 cells per cluster to speed up calculation
sample_size <- table(data$subgroups)
sample_size[ sample_size > 150 ] <- 150
DGE_cells <- lapply(names(sample_size), function(x){
  set.seed(1)
  sample( colnames(data) [ data$subgroups == x ] , size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# subset based on previous step, removing NA & setting Identity to these groups (cluster_age)
DGE_DATA <- data[, DGE_cells]
DGE_DATA <- DGE_DATA[, DGE_DATA$subgroups %in% c("8_5-6", "8_7-8", "8_9-11", "8_12-14")]
DGE_DATA <- SetIdent(DGE_DATA, value = DGE_DATA@meta.data$subgroups)

# find markers accross every category
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 150,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05,
                          return.thresh = 1)

# add metrics and save table
detable <- detable[ detable$p_val < 0.05,  ]
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2( (detable$pct.1+1) / (detable$pct.2+1) )
#write.csv2(detable, paste0(myfolder, "detable_selection_HL_split_by_age.csv"))

# Save results as a dot plot
detable$groups <- paste0(detable$cell_cluster,detable$cluster)

detable %>%
  group_by(groups) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(10, avg_log2FC) ->
  top10

# save
pdf(paste0(myfolder, "DGE_temporal_selection_HL8.pdf"), width = 5, height = 10)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top10 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()
```


### DGE - selection of HL clusters (HL_22)

```{r}
# set seed for reproducibility when sampling
set.seed(1)

# Create groups based on both clusters and ages and save to the object as a factor.
groups <- paste0(data$high_level_clusters_removed, '_', as.character(data$age_groups))
data$subgroups <- factor(groups, levels = mixedsort(unique(groups)))

# take only 150 cells per cluster to speed up calculation
sample_size <- table(data$subgroups)
sample_size[ sample_size > 150 ] <- 150
DGE_cells <- lapply(names(sample_size), function(x){
  set.seed(1)
  sample( colnames(data) [ data$subgroups == x ] , size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# subset based on previous step, removing NA & setting Identity to these groups (cluster_age)
DGE_DATA <- data[, DGE_cells]
DGE_DATA <- DGE_DATA[, DGE_DATA$subgroups %in% c("22_5-6", "22_7-8", "22_9-11", "22_12-14")]
DGE_DATA <- SetIdent(DGE_DATA, value = DGE_DATA@meta.data$subgroups)

# find markers accross every category
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 150,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05,
                          return.thresh = 1)

# add metrics and save table
detable <- detable[ detable$p_val < 0.05,  ]
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2( (detable$pct.1+1) / (detable$pct.2+1) )
#write.csv2(detable, paste0(myfolder, "detable_selection_HL_split_by_age.csv"))

# Save results as a dot plot
detable$groups <- paste0(detable$cell_cluster,detable$cluster)

detable %>%
  group_by(groups) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(10, avg_log2FC) ->
  top10

# save
pdf(paste0(myfolder, "DGE_temporal_selection_HL22.pdf"), width = 5, height = 10)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top10 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()
```


### DGE - selection of HL clusters (HL_23)

```{r}
# set seed for reproducibility when sampling
set.seed(1)

# Create groups based on both clusters and ages and save to the object as a factor.
groups <- paste0(data$high_level_clusters_removed, '_', as.character(data$age_groups))
data$subgroups <- factor(groups, levels = mixedsort(unique(groups)))

# take only 150 cells per cluster to speed up calculation
sample_size <- table(data$subgroups)
sample_size[ sample_size > 150 ] <- 150
DGE_cells <- lapply(names(sample_size), function(x){
  set.seed(1)
  sample( colnames(data) [ data$subgroups == x ] , size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# subset based on previous step, removing NA & setting Identity to these groups (cluster_age)
DGE_DATA <- data[, DGE_cells]
DGE_DATA <- DGE_DATA[, DGE_DATA$subgroups %in% c("23_5-6", "23_7-8", "23_9-11", "23_12-14")]
DGE_DATA <- SetIdent(DGE_DATA, value = DGE_DATA@meta.data$subgroups)

# find markers accross every category
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 150,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05,
                          return.thresh = 1)

# add metrics and save table
detable <- detable[ detable$p_val < 0.05,  ]
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2( (detable$pct.1+1) / (detable$pct.2+1) )
#write.csv2(detable, paste0(myfolder, "detable_selection_HL_split_by_age.csv"))

# Save results as a dot plot
detable$groups <- paste0(detable$cell_cluster,detable$cluster)

detable %>%
  group_by(groups) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(10, avg_log2FC) ->
  top10

# save
pdf(paste0(myfolder, "DGE_temporal_selection_HL23.pdf"), width = 5, height = 10)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top10 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()
```


### DGE - selection of HL clusters (HL_20)

```{r}
# set seed for reproducibility when sampling
set.seed(1)

# Create groups based on both clusters and ages and save to the object as a factor.
groups <- paste0(data$high_level_clusters_removed, '_', as.character(data$age_groups))
data$subgroups <- factor(groups, levels = mixedsort(unique(groups)))

# take only 150 cells per cluster to speed up calculation
sample_size <- table(data$subgroups)
sample_size[ sample_size > 150 ] <- 150
DGE_cells <- lapply(names(sample_size), function(x){
  set.seed(1)
  sample( colnames(data) [ data$subgroups == x ] , size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# subset based on previous step, removing NA & setting Identity to these groups (cluster_age)
DGE_DATA <- data[, DGE_cells]
DGE_DATA <- DGE_DATA[, DGE_DATA$subgroups %in% c("20_5-6", "20_7-8", "20_9-11", "20_12-14")]
DGE_DATA <- SetIdent(DGE_DATA, value = DGE_DATA@meta.data$subgroups)

# find markers accross every category
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 150,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05,
                          return.thresh = 1)

# add metrics and save table
detable <- detable[ detable$p_val < 0.05,  ]
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2( (detable$pct.1+1) / (detable$pct.2+1) )
#write.csv2(detable, paste0(myfolder, "detable_selection_HL_split_by_age.csv"))

# Save results as a dot plot
detable$groups <- paste0(detable$cell_cluster,detable$cluster)

detable %>%
  group_by(groups) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(10, avg_log2FC) ->
  top10

# save
pdf(paste0(myfolder, "DGE_temporal_selection_HL20.pdf"), width = 5, height = 10)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top10 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()
```


### DGE - selection of HL clusters (HL_1)

```{r}
# set seed for reproducibility when sampling
set.seed(1)

# Create groups based on both clusters and ages and save to the object as a factor.
groups <- paste0(data$high_level_clusters_removed, '_', as.character(data$age_groups))
data$subgroups <- factor(groups, levels = mixedsort(unique(groups)))

# take only 150 cells per cluster to speed up calculation
sample_size <- table(data$subgroups)
sample_size[ sample_size > 150 ] <- 150
DGE_cells <- lapply(names(sample_size), function(x){
  set.seed(1)
  sample( colnames(data) [ data$subgroups == x ] , size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# subset based on previous step, removing NA & setting Identity to these groups (cluster_age)
DGE_DATA <- data[, DGE_cells]
DGE_DATA <- DGE_DATA[, DGE_DATA$subgroups %in% c("1_5-6", "1_7-8", "1_9-11", "1_12-14")]
DGE_DATA <- SetIdent(DGE_DATA, value = DGE_DATA@meta.data$subgroups)

# find markers accross every category
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 150,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05,
                          return.thresh = 1)

# add metrics and save table
detable <- detable[ detable$p_val < 0.05,  ]
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2( (detable$pct.1+1) / (detable$pct.2+1) )
#write.csv2(detable, paste0(myfolder, "detable_selection_HL_split_by_age.csv"))

# Save results as a dot plot
detable$groups <- paste0(detable$cell_cluster,detable$cluster)

detable %>%
  group_by(groups) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(10, avg_log2FC) ->
  top10

# save
pdf(paste0(myfolder, "DGE_temporal_selection_HL1.pdf"), width = 5, height = 10)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top10 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()
```


### DGE normal - selection of HL clusters (HL_8, HL_10, HL_22) (Extended Fig. 2C)

```{r}
# set seed for reproducibility when sampling
set.seed(1)

# take only 150 cells per cluster to speed up calculation
sample_size <- table(data$high_level_clusters_removed)
sample_size[ sample_size > 150 ] <- 150
DGE_cells <- lapply(names(sample_size), function(x){
  set.seed(1)
  sample( colnames(data) [ data$high_level_clusters_removed == x ] , size = sample_size[x])
  })
DGE_cells <- unlist(DGE_cells)

# subset based on previous step, removing NA & setting Identity to these groups (cluster_age)
DGE_DATA <- data[, DGE_cells]
DGE_DATA <- DGE_DATA[, DGE_DATA$high_level_clusters_removed %in% c("8", "10", "22")]
DGE_DATA <- SetIdent(DGE_DATA, value = DGE_DATA@meta.data$high_level_clusters_removed)
unique(DGE_DATA$high_level_clusters_removed)

# find markers accross every category
detable <- FindAllMarkers(DGE_DATA,
                          only.pos = TRUE,
                          max.cells.per.ident = 150,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05,
                          return.thresh = 1)

# add metrics and save table
detable <- detable[ detable$p_val < 0.05,  ]
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2( (detable$pct.1+1) / (detable$pct.2+1) )
#write.csv2(detable, paste0(myfolder, "detable_selection_HL_split_by_age.csv"))

# Save results as a dot plot
detable$groups <- paste0(detable$cell_cluster,detable$cluster)

detable %>%
  group_by(groups) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(10, avg_log2FC) ->
  top10

# save
pdf(paste0(myfolder, "DGE_selection_HL8_10_22.pdf"), width = 4.5, height = 8)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(DGE_DATA, features = top10 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()
```


### DGE - special genes temporal in EN clusters

```{r}
EN <- readRDS(file = paste0(myfolder, "rds_objects/endothelial.rds"))

# specific list of marker to project the dotplot on
my_markers <- c("RALYL", "CACNA2D1", "NRG3", "POLQ", "ARL15", "EYS", "PLD5", "KCNMB2-AS", "NRXN3", "TMEM132D", "KCNQ5", "AC015574.1", "AL357507.1", "DOCK4", "ARHGEF28", "RGS6")

values_to_keep <- c("1", "12", "3", "6", "18", "17", "7", "19", "15", "13", "10", "2", "5", "21", "16", "20", "11", "14")

data_EN <- EN[, EN$clusters_subset %in% values_to_keep]

# Create groups based on both clusters and ages and save to the object as a factor.
groups <- paste0(data_EN$clusters_subset, '_', as.character(data_EN$age_groups))
data_EN$subgroups <- factor(groups, levels = mixedsort(unique(groups)))

data_EN <- SetIdent(data_EN, value = "subgroups")

p <- DotPlot(data_EN, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DGE_temporal_spe_genes_EN_cl.pdf"), width = 30, height = 5)
print(p)
dev.off()

rm(p)
gc()
```


### DGE - special genes in EN clusters

```{r}
EN <- readRDS(file = paste0(myfolder, "rds_objects/endothelial.rds"))

# specific list of marker to project the dotplot on
my_markers <- c("NPR3", "APCDD1", "GJA5", "HEY1", "RGCC", "APLNR", "CCL21", "TMSB4X", "PDE4C", "ACKR1", "ACKR4", "NRP1", "BMPR2", "ICAM1", "VCAM1", "SELE", "SELP", "VWF")

values_to_keep <- c("1", "12", "3", "6", "18", "17", "7", "19", "15", "13", "10", "2", "5", "21", "16", "20", "11", "14")

data_EN <- EN[, EN$clusters_subset %in% values_to_keep]

# # Create groups based on both clusters and ages and save to the object as a factor.
# groups <- paste0(data_EN$clusters_subset, '_', as.character(data_EN$age_groups))
# data_EN$subgroups <- factor(groups, levels = mixedsort(unique(groups)))

data_EN$clusters_subset <- factor(data_EN$clusters_subset, levels = values_to_keep)
data_EN <- SetIdent(data_EN, value = "clusters_subset")

p <- DotPlot(data_EN, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DGE_spe_genes_EN_cl_r2q1.pdf"), width = 10, height = 5)
print(p)
dev.off()

rm(p)
gc()
```


### DGE - special genes in CM clusters

```{r}
CM <- readRDS(file = paste0(myfolder, "rds_objects/cardiomyocytes02.rds"))

# specific list of marker to project the dotplot on
my_markers <- c("TENM1", "TENM2", "TENM3", "TENM4", "ADGRL1", "ADGRL2", "ADGRL3", "ADGRL4")

values_to_keep <- c("19-2", "19-1", "12", "9", "14")

data_CM <- CM[, CM$clusters_subset_split %in% values_to_keep]

data_CM$clusters_subset_split <- factor(data_CM$clusters_subset_split, levels = values_to_keep)
data_CM <- SetIdent(data_CM, value = "clusters_subset_split")

p <- DotPlot(data_CM, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DGE_spe_genes_CM_R2Q2.pdf"), width = 5, height = 4)
print(p)
dev.off()

rm(p)
gc()
```


### FeaturePlots - special in HL_15 clusters (Extended Fig. 3C)

```{r}
data_HL <- data[, data$high_level_clusters_removed %in% "15"]
data_HL <- SetIdent(data_HL, value = "high_level_clusters_removed")

p1 <- FeaturePlot(data_HL, reduction = "umap_oi", features = 'FLT3', order = TRUE)
p2 <- FeaturePlot(data_HL, reduction = "umap_oi", features = 'CX3CR1', order = TRUE)
p3 <- FeaturePlot(data_HL, reduction = "umap_oi", features = 'CSF1R', order = TRUE)

pdf(paste0(myfolder, "complementary/umap_HL15_FLT3.pdf"), width = 5, height = 5)
print(p1)
dev.off()

pdf(paste0(myfolder, "complementary/umap_HL15_CX3CR1.pdf"), width = 5, height = 5)
print(p2)
dev.off()

pdf(paste0(myfolder, "complementary/umap_HL15_CSF1R.pdf"), width = 5, height = 5)
print(p3)
dev.off()

rm(p)
gc()
```


### FeaturePlots - special in HL_29 clusters (Extended Fig. 3C)

```{r}
data_HL <- data[, data$high_level_clusters_removed %in% "29"]
data_HL <- SetIdent(data_HL, value = "high_level_clusters_removed")

p1 <- FeaturePlot(data_HL, reduction = "umap_oi", features = 'JCHAIN', order = TRUE)
p2 <- FeaturePlot(data_HL, reduction = "umap_oi", features = 'CD3E', order = TRUE)
p3 <- FeaturePlot(data_HL, reduction = "umap_oi", features = 'KLRB1', order = TRUE)

pdf(paste0(myfolder, "complementary/umap_HL29_JCHAIN.pdf"), width = 5, height = 5)
print(p1)
dev.off()

pdf(paste0(myfolder, "complementary/umap_HL29_CD3E.pdf"), width = 5, height = 5)
print(p2)
dev.off()

pdf(paste0(myfolder, "complementary/umap_HL29_KLRB1.pdf"), width = 5, height = 5)
print(p3)
dev.off()

rm(p)
gc()
```


## ISS DEG (HL/DL)

```{r}
#load file containing kept cells and DL clusters
annotation_split <- read.csv(file = paste0("/home/rstudio/project/output/HDCA_heart_sc_analysis_docker/stereoscope/annotation_split.csv"))

#load ISS gene list
iss_gene <- read.csv(file = paste0(myfolder, "supplementary_tables/extended_table_01_ISS_gene_list.csv"))
```


### Keep only cells present in annotation_split

```{r}
data_kept_cells <- data[, colnames(data) %in% annotation_split$ID]
```

```{r}
#sanity check
identical(colnames(data_kept_cells), sort(annotation_split$ID))
```


## Add DL annotation

```{r}
annot_dl <- setNames(annotation_split$clusters_subset_split, nm = annotation_split$ID)
data_kept_cells$annot_dl <- factor(annot_dl[as.character(colnames(data_kept_cells))], levels = unique(annotation_split$clusters_subset_split))


#sanity check
table(data_kept_cells$annot_dl, data_kept_cells$high_level_clusters_removed)

rm(annot_dl, annotation_split)
```


### Compute DEG for ISS gene over HL clusters

```{r}
# specific list of marker to project the dotplot on
my_markers <- iss_gene$Gene.list

# remove undesired cells
cells_to_remove <- data$cell.types == "HL_excl_1" | 
                   data$cell.types == "HL_excl_2" | 
                   data$cell.types == "HL_excl_3" | 
                   data$cell.types == "HL_excl_4"

data_HL <- data[,!cells_to_remove]
rm(cells_to_remove)

# reorder the cell types desired
data_HL$high_level_clusters_removed <- factor(data_HL$cell.types, 
                                              levels = c("Mat_vCM",
                                                         "Mat_aCM",
                                                         "MetAct_vCM_1",
                                                         "MetAct_vCM_2",
                                                         "MetAct_aCM",
                                                         "Prol_CM",
                                                         "Immat_CM",
                                                         "TMSB10high_C_1",
                                                         "TMSB10high_C_2",
                                                         "MacroVasc_EC",
                                                         "MicroVasc_EC",
                                                         "PDE4Chigh_EC",
                                                         "LEC",
                                                         "Endoc_EC",
                                                         "EndocCush_EC",
                                                         "Valve_MC",
                                                         "PDE4Chigh_FB",
                                                         "Int_FB",
                                                         "Peric_MC",
                                                         "OFT_FB",
                                                         "EPDC",
                                                         "AnnFibr_FB",
                                                         "Prol_FB",
                                                         "OFT_SMC",
                                                         "CA_SMC",
                                                         "PC",
                                                         "EpC",
                                                         "NB-N",
                                                         "SCP-GC",
                                                         "MyC",
                                                         "LyC"))

data_HL <- SetIdent(data_HL, value = "high_level_clusters_removed")

p <- DotPlot(data_HL, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "HL_dotplot_ISS_genes.pdf"), width = 11, height = 24)
  print(p)
dev.off()

rm(data_HL)
gc()
```


### Compute DEG for ISS gene over DL clusters

```{r}
# specific list of marker to project the dotplot on
my_markers <- iss_gene$Gene.list

# reorder the cell types desired
data_kept_cells$annot_dl <- factor(data_kept_cells$annot_dl, 
                                   levels = c("CM_1",   "CM_2",   "CM_3",   "CM_4-1", "CM_4-2", "CM_5",   "CM_6",   "CM_7",   "CM_8-1", "CM_8-2", "CM_8-3", "CM_9",   "CM_10", "CM_12",  "CM_14",  "CM_17",  "CM_18",  "CM_19-1", "CM_19-2",
                                              "EN_1",  "EN_2",  "EN_3",  "EN_5",  "EN_6",  "EN_7",  "EN_10", "EN_11", "EN_12", "EN_13", "EN_14", "EN_15", "EN_16", "EN_17", "EN_18", "EN_19", "EN_20", "EN_21",
                                              "FB_1",  "FB_2",  "FB_3",  "FB_4",  "FB_5",  "FB_6",  "FB_7",  "FB_8",  "FB_9",  "FB_11", "FB_12", "FB_13", "FB_14", "FB_15", "FB_17", "FB_18", "FB_19", "FB_20",
                                              "IN_1",  "IN_3",  "IN_4",  "IN_7",  "IN_8",  "IN_11", "IN_12", "IN_13", "IN_17", "IN_19",
                                              "HL_6",    "HL_8",    "HL_10",   "HL_15",   "HL_18_1", "HL_20",   "HL_22",   "HL_29",   "HL_34"
                                              ))

data_kept_cells <- SetIdent(data_kept_cells, value = "annot_dl")

p <- DotPlot(data_kept_cells, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_ISS_genes.pdf"), width = 24, height = 24)
  print(p)
dev.off()
```

```{r}
#export annotation DL
write.csv(data_kept_cells$annot_dl, file = paste0(myfolder, "annotation_dl_extra_metadata.csv"))
```


### Specific gene over all DL clusters - CPCS

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("SHOX2", "PRRX1", "TBX18", "ZNF385B", "TENM2", "TENM3", "TENM4", "GDF10", "SLIT2",
                "RCAN1", "IRX1", "IRX2", "IRX5", "BRINP3", "SEMA3A", "IRX3", "CSMD1", "PLXNA4",
                "HCN1", "HCN4", "CACNA1D", "CACNA1G")

# reorder the cell types desired
data_kept_cells$annot_dl <- factor(data_kept_cells$annot_dl, 
                                   levels = c("CM_1",   "CM_2",   "CM_3",   "CM_4-1", "CM_4-2", "CM_5",   "CM_6",   "CM_7",   "CM_8-1", "CM_8-2", "CM_8-3", "CM_9",   "CM_10", "CM_12",  "CM_14",  "CM_17",  "CM_18",  "CM_19-1", "CM_19-2",
                                              "EN_1",  "EN_2",  "EN_3",  "EN_5",  "EN_6",  "EN_7",  "EN_10", "EN_11", "EN_12", "EN_13", "EN_14", "EN_15", "EN_16", "EN_17", "EN_18", "EN_19", "EN_20", "EN_21",
                                              "FB_1",  "FB_2",  "FB_3",  "FB_4",  "FB_5",  "FB_6",  "FB_7",  "FB_8",  "FB_9",  "FB_11", "FB_12", "FB_13", "FB_14", "FB_15", "FB_17", "FB_18", "FB_19", "FB_20",
                                              "IN_1",  "IN_3",  "IN_4",  "IN_7",  "IN_8",  "IN_11", "IN_12", "IN_13", "IN_17", "IN_19",
                                              "HL_6",    "HL_8",    "HL_10",   "HL_15",   "HL_18_1", "HL_20",   "HL_22",   "HL_29",   "HL_34"
                                              ))

data_kept_cells <- SetIdent(data_kept_cells, value = "annot_dl")

p <- DotPlot(data_kept_cells, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_CPCS_marker_genes.pdf"), width = 20, height = 5)
  print(p)
dev.off()
```


### Specific gene over all DL clusters - ox sensing

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("EPAS1", "COX4I2", "HIGD1C")

# reorder the cell types desired
data_kept_cells$annot_dl <- factor(data_kept_cells$annot_dl, 
                                   levels = c("CM_1",   "CM_2",   "CM_3",   "CM_4-1", "CM_4-2", "CM_5",   "CM_6",   "CM_7",   "CM_8-1", "CM_8-2", "CM_8-3", "CM_9",   "CM_10", "CM_12",  "CM_14",  "CM_17",  "CM_18",  "CM_19-1", "CM_19-2",
                                              "EN_1",  "EN_2",  "EN_3",  "EN_5",  "EN_6",  "EN_7",  "EN_10", "EN_11", "EN_12", "EN_13", "EN_14", "EN_15", "EN_16", "EN_17", "EN_18", "EN_19", "EN_20", "EN_21",
                                              "FB_1",  "FB_2",  "FB_3",  "FB_4",  "FB_5",  "FB_6",  "FB_7",  "FB_8",  "FB_9",  "FB_11", "FB_12", "FB_13", "FB_14", "FB_15", "FB_17", "FB_18", "FB_19", "FB_20",
                                              "IN_1",  "IN_3",  "IN_4",  "IN_7",  "IN_8",  "IN_11", "IN_12", "IN_13", "IN_17", "IN_19",
                                              "HL_6",    "HL_8",    "HL_10",   "HL_15",   "HL_18_1", "HL_20",   "HL_22",   "HL_29",   "HL_34"
                                              ))

data_kept_cells <- SetIdent(data_kept_cells, value = "annot_dl")

p <- DotPlot(data_kept_cells, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_ox-sensing_marker_genes.pdf"), width = 20, height = 5)
  print(p)
dev.off()
```

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("EPAS1", "COX4I2", "HIGD1C", "RGS5", "RGS4", "NDUFA4L2")

# reorder the cell types desired
data_kept_cells$annot_dl <- factor(data_kept_cells$annot_dl, 
                                   levels = c("CM_1",   "CM_2",   "CM_3",   "CM_4-1", "CM_4-2", "CM_5",   "CM_6",   "CM_7",   "CM_8-1", "CM_8-2", "CM_8-3", "CM_9",   "CM_10", "CM_12",  "CM_14",  "CM_17",  "CM_18",  "CM_19-1", "CM_19-2",
                                              "EN_1",  "EN_2",  "EN_3",  "EN_5",  "EN_6",  "EN_7",  "EN_10", "EN_11", "EN_12", "EN_13", "EN_14", "EN_15", "EN_16", "EN_17", "EN_18", "EN_19", "EN_20", "EN_21",
                                              "FB_1",  "FB_2",  "FB_3",  "FB_4",  "FB_5",  "FB_6",  "FB_7",  "FB_8",  "FB_9",  "FB_11", "FB_12", "FB_13", "FB_14", "FB_15", "FB_17", "FB_18", "FB_19", "FB_20",
                                              "IN_1",  "IN_3",  "IN_4",  "IN_7",  "IN_8",  "IN_11", "IN_12", "IN_13", "IN_17", "IN_19",
                                              "HL_6",    "HL_8",    "HL_10",   "HL_15",   "HL_18_1", "HL_20",   "HL_22",   "HL_29",   "HL_34"
                                              ))

data_kept_cells <- SetIdent(data_kept_cells, value = "annot_dl")

p <- DotPlot(data_kept_cells, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_ox-sensing_marker_genes_v2.pdf"), width = 20, height = 4)
  print(p)
dev.off()
```


### Specific gene over all DL clusters - CVP-FVP

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("MEIS2", "ISL1", "LGR5", "PDGFRA", "BMP4")

# reorder the cell types desired
data_kept_cells$annot_dl <- factor(data_kept_cells$annot_dl, 
                                   levels = c("CM_1",   "CM_2",   "CM_3",   "CM_4-1", "CM_4-2", "CM_5",   "CM_6",   "CM_7",   "CM_8-1", "CM_8-2", "CM_8-3", "CM_9",   "CM_10", "CM_12",  "CM_14",  "CM_17",  "CM_18",  "CM_19-1", "CM_19-2",
                                              "EN_1",  "EN_2",  "EN_3",  "EN_5",  "EN_6",  "EN_7",  "EN_10", "EN_11", "EN_12", "EN_13", "EN_14", "EN_15", "EN_16", "EN_17", "EN_18", "EN_19", "EN_20", "EN_21",
                                              "FB_1",  "FB_2",  "FB_3",  "FB_4",  "FB_5",  "FB_6",  "FB_7",  "FB_8",  "FB_9",  "FB_11", "FB_12", "FB_13", "FB_14", "FB_15", "FB_17", "FB_18", "FB_19", "FB_20",
                                              "IN_1",  "IN_3",  "IN_4",  "IN_7",  "IN_8",  "IN_11", "IN_12", "IN_13", "IN_17", "IN_19",
                                              "HL_6",    "HL_8",    "HL_10",   "HL_15",   "HL_18_1", "HL_20",   "HL_22",   "HL_29",   "HL_34"
                                              ))

data_kept_cells <- SetIdent(data_kept_cells, value = "annot_dl")

p <- DotPlot(data_kept_cells, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_CVP_marker_genes.pdf"), width = 20, height = 4)
  print(p)
dev.off()
```

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("SOX4", "HES4", "CTHRC1", "IGFBP2")

# reorder the cell types desired
data_kept_cells$annot_dl <- factor(data_kept_cells$annot_dl, 
                                   levels = c("CM_1",   "CM_2",   "CM_3",   "CM_4-1", "CM_4-2", "CM_5",   "CM_6",   "CM_7",   "CM_8-1", "CM_8-2", "CM_8-3", "CM_9",   "CM_10", "CM_12",  "CM_14",  "CM_17",  "CM_18",  "CM_19-1", "CM_19-2",
                                              "EN_1",  "EN_2",  "EN_3",  "EN_5",  "EN_6",  "EN_7",  "EN_10", "EN_11", "EN_12", "EN_13", "EN_14", "EN_15", "EN_16", "EN_17", "EN_18", "EN_19", "EN_20", "EN_21",
                                              "FB_1",  "FB_2",  "FB_3",  "FB_4",  "FB_5",  "FB_6",  "FB_7",  "FB_8",  "FB_9",  "FB_11", "FB_12", "FB_13", "FB_14", "FB_15", "FB_17", "FB_18", "FB_19", "FB_20",
                                              "IN_1",  "IN_3",  "IN_4",  "IN_7",  "IN_8",  "IN_11", "IN_12", "IN_13", "IN_17", "IN_19",
                                              "HL_6",    "HL_8",    "HL_10",   "HL_15",   "HL_18_1", "HL_20",   "HL_22",   "HL_29",   "HL_34"
                                              ))

data_kept_cells <- SetIdent(data_kept_cells, value = "annot_dl")

p <- DotPlot(data_kept_cells, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_FVP_marker_genes.pdf"), width = 20, height = 4)
  print(p)
dev.off()
```


### Specific gene over all DL clusters - R2Q2

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("TENM1", "TENM2", "TENM3", "TENM4", "ADGRL1", "ADGRL2", "ADGRL3", "ADGRL4", "LRRC4C", "NTNG1", "GDF10", "NRXN1", "NRXN3", "NECTIN2", "CLSTN1", "CLSTN2", "NXPH1", "SNAP25", "STX1A", "VAMP2")

# reorder the cell types desired
data_kept_cells$annot_dl <- factor(data_kept_cells$annot_dl, 
                                   levels = c("CM_1",   "CM_2",   "CM_3",   "CM_4-1", "CM_4-2", "CM_5",   "CM_6",   "CM_7",   "CM_8-1", "CM_8-2", "CM_8-3", "CM_9",   "CM_10", "CM_12",  "CM_14",  "CM_17",  "CM_18",  "CM_19-1", "CM_19-2",
                                              "EN_1",  "EN_2",  "EN_3",  "EN_5",  "EN_6",  "EN_7",  "EN_10", "EN_11", "EN_12", "EN_13", "EN_14", "EN_15", "EN_16", "EN_17", "EN_18", "EN_19", "EN_20", "EN_21",
                                              "FB_1",  "FB_2",  "FB_3",  "FB_4",  "FB_5",  "FB_6",  "FB_7",  "FB_8",  "FB_9",  "FB_11", "FB_12", "FB_13", "FB_14", "FB_15", "FB_17", "FB_18", "FB_19", "FB_20",
                                              "IN_1",  "IN_3",  "IN_4",  "IN_7",  "IN_8",  "IN_11", "IN_12", "IN_13", "IN_17", "IN_19",
                                              "HL_6",    "HL_8",    "HL_10",   "HL_15",   "HL_18_1", "HL_20",   "HL_22",   "HL_29",   "HL_34"
                                              ))

data_kept_cells <- SetIdent(data_kept_cells, value = "annot_dl")

p <- DotPlot(data_kept_cells, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_R2Q2_marker_genes.pdf"), width = 20, height = 5)
  print(p)
dev.off()
```


### Specific gene over all DL clusters - R2Q4

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("MESP1", "NKX2-5", "KIT", "KITLG", "CD34", "ABCG2", "ENG", "ALCAM", "GATA4", "MEF2C", "PECAM1", "PTPRC", "KDR", "CD44", "THY1", "CD46", "ITGB1", "MAB21L2", "ETV2", "FUT4", "SOX4", "HES4", "CTHRC1", "IGFBP2", "MEIS2", "LGR5", "BMP4", "ISL1", "PDGFRA")

# reorder the cell types desired
data_kept_cells$annot_dl <- factor(data_kept_cells$annot_dl, 
                                   levels = c("CM_1",   "CM_2",   "CM_3",   "CM_4-1", "CM_4-2", "CM_5",   "CM_6",   "CM_7",   "CM_8-1", "CM_8-2", "CM_8-3", "CM_9",   "CM_10", "CM_12",  "CM_14",  "CM_17",  "CM_18",  "CM_19-1", "CM_19-2",
                                              "EN_1",  "EN_2",  "EN_3",  "EN_5",  "EN_6",  "EN_7",  "EN_10", "EN_11", "EN_12", "EN_13", "EN_14", "EN_15", "EN_16", "EN_17", "EN_18", "EN_19", "EN_20", "EN_21",
                                              "FB_1",  "FB_2",  "FB_3",  "FB_4",  "FB_5",  "FB_6",  "FB_7",  "FB_8",  "FB_9",  "FB_11", "FB_12", "FB_13", "FB_14", "FB_15", "FB_17", "FB_18", "FB_19", "FB_20",
                                              "IN_1",  "IN_3",  "IN_4",  "IN_7",  "IN_8",  "IN_11", "IN_12", "IN_13", "IN_17", "IN_19",
                                              "HL_6",    "HL_8",    "HL_10",   "HL_15",   "HL_18_1", "HL_20",   "HL_22",   "HL_29",   "HL_34"
                                              ))

data_kept_cells <- SetIdent(data_kept_cells, value = "annot_dl")

p <- DotPlot(data_kept_cells, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_R2Q4_marker_genes.pdf"), width = 20, height = 7)
  print(p)
dev.off()
```


### Specific gene over all DL clusters - R2Q4 - v2

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("MESP1", "NKX2-5", "FUT4", "KDR", "PDGFRA", "HCN4", "ISL1", "WT1", "TBX18", "SCX", "SEMA3D", "ADGRL2", "MEIS2", "LGR5", "BMP4", "SOX4", "HES4", "CTHRC1", "IGFBP2", "SALL1", "HOPX1", "FOXA2", "GFRA2", "KIT", "CD24")

# reorder the cell types desired
data_kept_cells$annot_dl <- factor(data_kept_cells$annot_dl, 
                                   levels = c("CM_1",   "CM_2",   "CM_3",   "CM_4-1", "CM_4-2", "CM_5",   "CM_6",   "CM_7",   "CM_8-1", "CM_8-2", "CM_8-3", "CM_9",   "CM_10", "CM_12",  "CM_14",  "CM_17",  "CM_18",  "CM_19-1", "CM_19-2",
                                              "EN_1",  "EN_2",  "EN_3",  "EN_5",  "EN_6",  "EN_7",  "EN_10", "EN_11", "EN_12", "EN_13", "EN_14", "EN_15", "EN_16", "EN_17", "EN_18", "EN_19", "EN_20", "EN_21",
                                              "FB_1",  "FB_2",  "FB_3",  "FB_4",  "FB_5",  "FB_6",  "FB_7",  "FB_8",  "FB_9",  "FB_11", "FB_12", "FB_13", "FB_14", "FB_15", "FB_17", "FB_18", "FB_19", "FB_20",
                                              "IN_1",  "IN_3",  "IN_4",  "IN_7",  "IN_8",  "IN_11", "IN_12", "IN_13", "IN_17", "IN_19",
                                              "HL_6",    "HL_8",    "HL_10",   "HL_15",   "HL_18_1", "HL_20",   "HL_22",   "HL_29",   "HL_34"
                                              ))

data_kept_cells <- SetIdent(data_kept_cells, value = "annot_dl")

p <- DotPlot(data_kept_cells, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_R2Q4_marker_genes_v2.pdf"), width = 20, height = 7)
  print(p)
dev.off()
```


### Specific gene over FB + HL_22 + HL_10 + HL_8 (Extended Fig. 2F)

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("PLA2G5", "APOE", "SEMA5A", "COL4A1", "HEYL", "COL4A2", "F5", "ITGA11", "GUCY1A1", "IGFBP7", "ITGA1", "ENPEP", "CCDC3", "EPS8", "HES4", "DLC1", "LAMC3", "MCAM", "ACTA2")

data_kept <- data_kept_cells[, data_kept_cells$annot_dl %in% c("FB_1",  "FB_2",  "FB_3",  "FB_4",  "FB_5",  "FB_6",  "FB_7",  "FB_8",  "FB_9",  "FB_11", "FB_12", "FB_13", "FB_14", "FB_15", "FB_17", "FB_18", "FB_19", "FB_20",
                                              "HL_8",    "HL_10", "HL_22")]

# reorder the cell types desired
data_kept$annot_dl <- factor(data_kept$annot_dl, 
                                   levels = c("FB_1",  "FB_2",  "FB_3",  "FB_4",  "FB_5",  "FB_6",  "FB_7",  "FB_8",  "FB_9",  "FB_11", "FB_12", "FB_13", "FB_14", "FB_15", "FB_17", "FB_18", "FB_19", "FB_20",
                                              "HL_8",    "HL_10", "HL_22"))

data_kept <- SetIdent(data_kept, value = "annot_dl")

p <- DotPlot(data_kept, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_markergenes_FB_HL22_HL10_HL8.pdf"), width = 8, height = 5)
  print(p)
dev.off()
```


### Specific gene over FB + HL_20 (Extended Fig. 4C)

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("WT1", "TBX18", "BNC1", "RARRES2", "UPK3B", "MSLN", "KRT19", "ITLN1", "PRG4", "SBSPON", "ALDH1A2", "EZR", "HAS1", "POSTN", "TWIST1", "SPARC", "CCBE1", "PDGFRA", "DPT", "IGFBP5", "RGS5", "FNDC1", "DCN", "COL1A1", "MOXD1", "MMP11", "CPB1", "VEGFC", "TCF21", "CENPF", "HMGB2")

data_kept <- data_kept_cells[, data_kept_cells$annot_dl %in% c("FB_1",  "FB_2",  "FB_3",  "FB_4",  "FB_5",  "FB_6",  "FB_7",  "FB_8",  "FB_9",  "FB_11", "FB_12", "FB_13", "FB_14", "FB_15", "FB_17", "FB_18", "FB_19", "FB_20",
                                              "HL_20")]

# reorder the cell types desired
data_kept$annot_dl <- factor(data_kept$annot_dl, 
                                   levels = c("FB_7", "FB_3", "FB_18", "FB_14", "FB_5", "FB_9", "FB_19", "FB_2", "FB_12", "FB_4", "FB_13", "FB_11", "FB_17", "FB_8", "FB_1", "FB_6", "FB_15", "FB_20",
                                              "HL_20"))

data_kept <- SetIdent(data_kept, value = "annot_dl")

p <- DotPlot(data_kept, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_markergenes_FB_HL20.pdf"), width = 8, height = 7)
  print(p)
dev.off()
```


### Specific gene over CM (Supp Fig. 4E)

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("TENM1", "TENM2", "TENM3", "TENM4", "ADGRL1", "ADGRL2", "ADGRL3", "ADGRL4", "LRRC4C", "NTNG1", "GDF10", "NRXN1", "NRXN3", "NECTIN2", "CLSTN1", "CLSTN2", "NXPH1", "SNAP25", "STX1A", "VAMP2")

data_kept <- data_kept_cells[, data_kept_cells$annot_dl %in% c("CM_1",   "CM_2",   "CM_3",   "CM_4-1", "CM_4-2", "CM_5",   "CM_6",   "CM_7",   "CM_8-1", "CM_8-2", "CM_8-3", "CM_9",   "CM_10", "CM_12",  "CM_14",  "CM_17",  "CM_18",  "CM_19-1", "CM_19-2")]

# reorder the cell types desired
data_kept$annot_dl <- factor(data_kept$annot_dl, 
                                   levels = c("CM_1",   "CM_2",   "CM_3",   "CM_4-1", "CM_4-2", "CM_5",   "CM_6",   "CM_7",   "CM_8-1", "CM_8-2", "CM_8-3", "CM_9",   "CM_10", "CM_12",  "CM_14",  "CM_17",  "CM_18",  "CM_19-1", "CM_19-2"))

data_kept <- SetIdent(data_kept, value = "annot_dl")

p <- DotPlot(data_kept, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_marker_genes_only_CM.pdf"), width = 8, height = 7)
  print(p)
dev.off()
```


### Specific gene over IN (Supp Fig. 4F)

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("TENM1", "TENM2", "TENM3", "TENM4", "ADGRL1", "ADGRL2", "ADGRL3", "ADGRL4", "LRRC4C", "NTNG1", "GDF10", "NRXN1", "NRXN3", "NECTIN2", "CLSTN1", "CLSTN2", "NXPH1", "SNAP25", "STX1A", "VAMP2")

data_kept <- data_kept_cells[, data_kept_cells$annot_dl %in% c("IN_1",  "IN_3",  "IN_4",  "IN_7",  "IN_8",  "IN_11", "IN_12", "IN_13", "IN_17", "IN_19")]

# reorder the cell types desired
data_kept$annot_dl <- factor(data_kept$annot_dl, 
                                   levels = c("IN_1",  "IN_3",  "IN_4",  "IN_7",  "IN_8",  "IN_11", "IN_12", "IN_13", "IN_17", "IN_19"))

data_kept <- SetIdent(data_kept, value = "annot_dl")

p <- DotPlot(data_kept, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_marker_genes_only_IN.pdf"), width = 6, height = 7)
  print(p)
dev.off()
```


### Specific gene over FB + HL_20 - R2Q3

```{r}
# specific list of marker to project the dotplot on
my_markers <- c("WT1", "TBX18", "BNC1", "RARRES2", "UPK3B", "MSLN", "KRT19", "ITLN1", "PRG4", "SBSPON", "ALDH1A2", "EZR", "HAS1", "POSTN", "TWIST1", "SPARC", "CCBE1", "PDGFRA", "DPT", "IGFBP5", "RGS5", "FNDC1", "DCN", "COL1A1", "MOXD1", "MMP11", "CPB1", "VEGFC", "TCF21", "CENPF", "HMGB2")

data_kept <- data_kept_cells[, data_kept_cells$annot_dl %in% c("FB_7", "FB_3", "HL_20")]

# reorder the cell types desired
data_kept$annot_dl <- factor(data_kept$annot_dl, 
                                   levels = c("FB_7", "FB_3", "HL_20"))

data_kept <- SetIdent(data_kept, value = "annot_dl")

p <- DotPlot(data_kept, features = my_markers) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8))

pdf(paste0(myfolder, "DL_dotplot_markergenes_FB_HL20_R2Q3.pdf"), width = 4, height = 7)
  print(p)
dev.off()
```


### DEG HL_20, FB_7, FB_3 (Extended Fig. 4D)

```{r}
cell_to_keep <- c("HL_20", "FB_7", "FB_3")
data_sel <- data_kept_cells[, data_kept_cells$annot_dl %in% cell_to_keep]

# reorder the cell types desired
data_sel$annot_dl <- factor(data_sel$annot_dl, levels = cell_to_keep)

data_sel <- SetIdent(data_sel, value = "annot_dl")


# The logfc threshold is set to 0.1 as compared to the default 0.25, which makes the detection of weaker signals possible (but it takes more computational power).
detable <- FindAllMarkers(data_sel,
                          only.pos = TRUE,
                          max.cells.per.ident = 300,
                          logfc.threshold = .1,
                          assay = "RNA",
                          min.pct = 0.05)

# Only keep DEG that have a p-value below 0.05.
detable <- detable[detable$p_val < 0.05, ]

# Adding two extra columns giving the difference (and log difference) between pct1 and pct2.
# pct1 is the percentage of cells in the specific cluster that express this gene.
# pct2 is the percentage of cells in all other clusters that express this gene.
detable$pct.diff <- detable$pct.1 - detable$pct.2
detable$log.pct.diff <- log2((detable$pct.1*99+1) / (detable$pct.2*99+1))


# Pick the 40 genes with the lowest p value for each cluster, then 20 highest based on pct.diff, then 5 highest based on avg_log2FC.
detable %>%
  group_by(cluster) %>%
  top_n(-40, p_val) %>%
  top_n(20, pct.diff) %>%
  top_n(10, avg_log2FC) ->
  top10

pdf(paste0(myfolder, "DEG_top10_HL_20_FB_7_FB_3.pdf"), width = 4.5, height = 7)
par(mfrow=c(1,1), mar=c(3,6,2,5))
DotPlot(data_sel, features = top10 %>% pull(gene) %>% unique()) +
  coord_flip() +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  scale_y_discrete() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
dev.off()

gc()
rm(detable, DGE_DATA, top10, DGE_cells, sample_size)
```


## Subcluster HL_20 (R2-Q3)

```{r}
hl20 <- data[, as.character(data$high_level_clusters_removed) == "20"] %>% 
  RunUMAP(dims = 1:50,
          reduction = "harmony_oi",
          n.neighbors = 30,
          n.epochs = 100,
          n.threads = 7,
          reduction.key = "subset",
          reduction.name = "umap_subset_hl20",
          seed.use = 42)

# Compute k nearest neighbours
set.seed(1)
nn <- CreateNNAdjacencyMatrix(data = hl20, reduction = "harmony_oi", k = 30, dims = 1:50)

# Cluster adjacency matrix
cl <- LouvainClustering(nn = nn, resolution = 0.6)
hl20$clusters_subset_hl20 <- factor(cl)

pdf(paste0(myfolder, "HL_subclusters_subset_hl20.pdf"), width = 10, height = 10)
DimPlot(hl20, reduction = "umap_subset_hl20", group.by = "clusters_subset_hl20", label = TRUE) + NoLegend()
dev.off()
```

```{r}
hl20 <- SetIdent(hl20, value = "clusters_subset_hl20")
de_markers_hl20 <- FindAllMarkers(hl20, only.pos = TRUE, logfc.threshold = 0.5) %>% 
  filter(p_val_adj < 0.01)

openxlsx::write.xlsx(de_markers_hl20 %>% filter(p_val_adj < 0.01),
                     file = paste0(myfolder, "detable_hl20.xlsx"))

p <- DotPlot(hl20, features = de_markers_hl20 %>% group_by(cluster) %>% 
          slice_head(n = 10) %>% pull(gene) %>% unique()) +
  coord_flip() +  # This flips the plot vertically
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  labs(x = "Genes", y = "Cluster")  # Flipped labels

pdf(paste0(myfolder, "dotplot_de_genes_hl20.pdf"), width = 4, height = 7)
print(p)
dev.off()
```

```{r}
# Create a single violin plot
pdf(file = paste0(myfolder, "Vln_nfeature_hl20.pdf"),  width = 5, height = 5)
VlnPlot(hl20, group.by = "clusters_subset_hl20", features = "nFeature_RNA", pt.size = 0) +
stat_summary(fun.y = median, geom='point', size = 25, colour = "black", shape = 95) +
theme(legend.position = 'none')
dev.off()
```

```{r}
# Create a single violin plot
pdf(file = paste0(myfolder, "Vln_nCount_hl20.pdf"),  width = 5, height = 5)
VlnPlot(hl20, group.by = "clusters_subset_hl20", features = "nCount_RNA", pt.size = 0) +
stat_summary(fun.y = median, geom='point', size = 25, colour = "black", shape = 95) +
theme(legend.position = 'none')
dev.off()
```


## Subcluster HL_20 (R2-Q3) - higher resolution

```{r}
# Cluster adjacency matrix
cl <- LouvainClustering(nn = nn, resolution = 0.7)
hl20$clusters_subset_hl20_2 <- factor(cl)

pdf(paste0(myfolder, "HL_subclusters_subset_hl20_2.pdf"), width = 10, height = 10)
DimPlot(hl20, reduction = "umap_subset_hl20", group.by = "clusters_subset_hl20_2", label = TRUE) + NoLegend()
dev.off()
```

```{r}
hl20 <- SetIdent(hl20, value = "clusters_subset_hl20_2")
de_markers_hl20_2 <- FindAllMarkers(hl20, only.pos = TRUE, logfc.threshold = 0.5) %>% 
  filter(p_val_adj < 0.01)

openxlsx::write.xlsx(de_markers_hl20_2 %>% filter(p_val_adj < 0.01),
                     file = paste0(myfolder, "detable_hl20_2.xlsx"))

p <- DotPlot(hl20, features = de_markers_hl20_2 %>% group_by(cluster) %>% 
          slice_head(n = 10) %>% pull(gene) %>% unique()) +
  coord_flip() +  # This flips the plot vertically
  theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5)) +
  scale_color_gradientn(colours = RColorBrewer::brewer.pal(n = 11, name = "RdBu") %>% rev()) +
  scale_x_discrete(limits = rev) +
  labs(x = "Genes", y = "Cluster")  # Flipped labels

pdf(paste0(myfolder, "dotplot_de_genes_hl20_2.pdf"), width = 4, height = 7)
print(p)
dev.off()
```

```{r}
# Create a single violin plot
pdf(file = paste0(myfolder, "Vln_nfeature_hl20_2.pdf"),  width = 5, height = 5)
VlnPlot(hl20, group.by = "clusters_subset_hl20_2", features = "nFeature_RNA", pt.size = 0) +
stat_summary(fun.y = median, geom='point', size = 25, colour = "black", shape = 95) +
theme(legend.position = 'none')
dev.off()
```

```{r}
# Create a single violin plot
pdf(file = paste0(myfolder, "Vln_nCount_hl20_2.pdf"),  width = 5, height = 5)
VlnPlot(hl20, group.by = "clusters_subset_hl20_2", features = "nCount_RNA", pt.size = 0) +
stat_summary(fun.y = median, geom='point', size = 25, colour = "black", shape = 95) +
theme(legend.position = 'none')
dev.off()
```

```{r}
hl20
table(data$high_level_clusters_removed)
table(hl20$clusters_subset_hl20_2)
table(hl20$clusters_subset_hl20_2, hl20$age)

```


## Pathway enrichments analysis

```{r}
library(gprofiler2)
```


### Load needed DEG

```{r}
# EN7 , EN17, EN18, FB4, FB11, FB13
detable_EN <- read.csv2(paste0(myfolder, "supplementary_tables/supplementary_table_5_DEG_EN.csv"))
detable_FB <- read.csv2(paste0(myfolder, "supplementary_tables/supplementary_table_6_DEG_FB.csv"))
```

```{r}
# EN7 = OF_VEC
# EN17 = IF_VEC
# EN18 = AtrSept_EC
# FB4 = VIC
# FB11 = Valve_MC_2
# FB13 = Valve_MC_1
```


### Pathway enrichment calculation

#### EN7 = OF_VEC

```{r}
input <- detable_EN[detable_EN$cluster %in% "OF_VEC", ] |>
  select(X, avg_log2FC, p_val_adj)

gene_list <- input$X
length(gene_list)
```

```{r}
# interested only in biological processes (GO:BP)
gostres_EN7 <- gost(query = gene_list, 
                organism = "hsapiens", ordered_query = TRUE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = TRUE, 
                user_threshold = 0.05, correction_method = "g_SCS", 
                domain_scope = "annotated", numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = FALSE)
```

```{r}
# results as dataframe
df <- as.data.frame(gostres_EN7[["result"]]) |> as.matrix()

# export dataframe
write.csv2(df, paste0(myfolder, "pathway_enrichment/EN7.csv"))
```


#### EN17 = IF_VEC

```{r}
input <- detable_EN[detable_EN$cluster %in% "IF_VEC", ] |>
  select(X, avg_log2FC, p_val_adj)

gene_list <- input$X
length(gene_list)
```

```{r}
# interested only in biological processes (GO:BP)
gostres <- gost(query = gene_list, 
                organism = "hsapiens", ordered_query = TRUE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = TRUE, 
                user_threshold = 0.05, correction_method = "g_SCS", 
                domain_scope = "annotated", numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = FALSE)
```

```{r}
# results as dataframe
df <- as.data.frame(gostres[["result"]]) |> as.matrix()

# export dataframe
write.csv2(df, paste0(myfolder, "pathway_enrichment/EN17.csv"))
```


#### EN18 = AtrSept_EC

```{r}
input <- detable_EN[detable_EN$cluster %in% "AtrSept_EC", ] |>
  select(X, avg_log2FC, p_val_adj)

gene_list <- input$X
length(gene_list)
```

```{r}
# interested only in biological processes (GO:BP)
gostres <- gost(query = gene_list, 
                organism = "hsapiens", ordered_query = TRUE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = TRUE, 
                user_threshold = 0.05, correction_method = "g_SCS", 
                domain_scope = "annotated", numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = FALSE)
```

```{r}
# results as dataframe
df <- as.data.frame(gostres[["result"]]) |> as.matrix()

# export dataframe
write.csv2(df, paste0(myfolder, "pathway_enrichment/EN18.csv"))
```


#### FB4 = VIC

```{r}
input <- detable_FB[detable_FB$cluster %in% "VIC", ] |>
  select(X, avg_log2FC, p_val_adj)

gene_list <- input$X
length(gene_list)
```

```{r}
# interested only in biological processes (GO:BP)
gostres_FB4 <- gost(query = gene_list, 
                organism = "hsapiens", ordered_query = TRUE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = TRUE, 
                user_threshold = 0.05, correction_method = "g_SCS", 
                domain_scope = "annotated", numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = FALSE)
```

```{r}
# results as dataframe
df <- as.data.frame(gostres_FB4[["result"]]) |> as.matrix()

# export dataframe
write.csv2(df, paste0(myfolder, "pathway_enrichment/FB4.csv"))
```


#### FB11 = Valve_MC_2

```{r}
input <- detable_FB[detable_FB$cluster %in% "Valve_MC_2", ] |>
  select(X, avg_log2FC, p_val_adj)

gene_list <- input$X
length(gene_list)
```

```{r}
# interested only in biological processes (GO:BP)
gostres_FB11 <- gost(query = gene_list, 
                organism = "hsapiens", ordered_query = TRUE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = TRUE, 
                user_threshold = 0.05, correction_method = "g_SCS", 
                domain_scope = "annotated", numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = FALSE)
```

```{r}
# results as dataframe
df <- as.data.frame(gostres_FB11[["result"]]) |> as.matrix()

# export dataframe
write.csv2(df, paste0(myfolder, "pathway_enrichment/FB11.csv"))
```


#### FB13 = Valve_MC_1

```{r}
input <- detable_FB[detable_FB$cluster %in% "Valve_MC_1", ] |>
  select(X, avg_log2FC, p_val_adj)

gene_list <- input$X
length(gene_list)
```

```{r}
# interested only in biological processes (GO:BP)
gostres <- gost(query = gene_list, 
                organism = "hsapiens", ordered_query = TRUE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = TRUE, 
                user_threshold = 0.05, correction_method = "g_SCS", 
                domain_scope = "annotated", numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = FALSE)
```

```{r}
# results as dataframe
df <- as.data.frame(gostres[["result"]]) |> as.matrix()

# export dataframe
write.csv2(df, paste0(myfolder, "pathway_enrichment/FB13.csv"))
```


### Pathway enrichement plotting
For selected clusters:

#### Load terms to highlight

```{r}
GO_highlight_EN7 <- read.xlsx("/home/rstudio/project/output/HDCA_heart_sc_analysis_docker/pathway_enrichment/pathway_enrichment.xlsx", sheet = 1, colNames = FALSE)
GO_highlight_FB4 <- read.xlsx("/home/rstudio/project/output/HDCA_heart_sc_analysis_docker/pathway_enrichment/pathway_enrichment.xlsx", sheet = 2, colNames = FALSE)
GO_highlight_FB11 <- read.xlsx("/home/rstudio/project/output/HDCA_heart_sc_analysis_docker/pathway_enrichment/pathway_enrichment.xlsx", sheet = 3, colNames = FALSE)
```

```{r}
#plotting
p <- gostplot(gostres_EN7, capped = FALSE, interactive = FALSE)
terms <- GO_highlight_EN7$X11
pp <- publish_gostplot(p, highlight_terms = terms, 
                       width = 10, height = 5, 
                       filename = paste0(myfolder, "pathway_enrichment/EN7.pdf"))
```

```{r}
#plotting
p <- gostplot(gostres_FB4, capped = FALSE, interactive = FALSE)
terms <- GO_highlight_FB4$X11

pp <- publish_gostplot(p, highlight_terms = terms, 
                       width = 10, height = 10, 
                       filename = paste0(myfolder, "pathway_enrichment/FB4.pdf"))
```

```{r}
#plotting
p <- gostplot(gostres_FB11, capped = FALSE, interactive = FALSE)
terms <- GO_highlight_FB11$X11
pp <- publish_gostplot(p, highlight_terms = terms, 
                       width = 10, height = 10, 
                       filename = paste0(myfolder, "pathway_enrichment/FB11.pdf"))
```


## LR selection clusters + selection LR pairs

Calculate mean function of Receptors (R) from SAN_CM with Ligands (L) from its direct neighbors (AVN_CM, Immat_CM_2, CALN1high, Aut_Neu_2, Ven_EC, Chrom_C) from niche network.

For each of the selected clusters, we keep genes that are differentially expressed in that cluster compared to the other DL clusters.

We assess all the pairs from the module of inputs from: https://static-content.springer.com/esm/art%3A10.1038%2Fs41586-023-06311-1/MediaObjects/41586_2023_6311_MOESM5_ESM.xlsx


### Load input module

```{r}
#load module of ligands-receptors
input_modules <- read.xlsx("/home/rstudio/project/data/metadata/41586_2023_6311_MOESM5_ESM.xlsx")

#keep only rows with values for protein_name_a (L) & protein_name_b (R)
input_curated <- input_modules |>
  filter(!is.na(protein_name_a) & protein_name_a != "") |>
  filter(!is.na(protein_name_b) & protein_name_b != "") |>
  select(protein_name_a, protein_name_b)

#specify list of LR to add to the module
LR_pairs_to_add <- list(c("TH", "ADRA1A"),
                        c("TH", "ADRA1B"),
                        c("TH", "ADRA1C"),
                        c("TH", "ADRA1D"),
                        c("TH", "ADRA2A"),
                        c("TH", "ADRA2B"),
                        c("TH", "ADRA2C"),
                        c("TH", "ADRB1"),
                        c("TH", "ADRB2"),
                        c("TH", "ADRB3"),
                        c("CHAT", "CHRM1"),
                        c("CHAT", "CHRM2"),
                        c("CHAT", "CHRM3"),
                        c("CHAT", "CHRM4"),
                        c("CHAT", "CHRM5"))

#add the specific pairs to the module
for (pairs in LR_pairs_to_add) {
  print(pairs)
  input_curated <- rbind(input_curated, pairs)
}

#create a specific list with 1) all ligands, 2) all receptors, 3) all ligands & receptors
ligands <- input_curated$protein_name_a
receptors <- input_curated$protein_name_b
ligands_receptors <- c(ligands, receptors)

rm(input_modules, LR_pairs_to_add)
gc()
```


### Load DEG from clusters of interest
SAN_CM = CM_19-2
AVN_CM = CM_19-1
Immat_CM_2 = CM_7
CALN1high_FB = FB_17
Aut_Neu_2 = IN_13
Ven_EC = EN_16
Chrom_C = IN_1

```{r}
#load DEG tables for each subclustering
detable_CM <- read.csv2(paste0(myfolder, "supplementary_tables/supplementary_table_3_DEG_CM.csv"))
detable_IN <- read.csv2(paste0(myfolder, "supplementary_tables/supplementary_table_4_DEG_IN.csv"))
detable_EN <- read.csv2(paste0(myfolder, "supplementary_tables/supplementary_table_5_DEG_EN.csv"))
detable_FB <- read.csv2(paste0(myfolder, "supplementary_tables/supplementary_table_6_DEG_FB.csv"))

#extract significant genes from each cluster of interest
SAN_CM_genes <- detable_CM |>filter(cluster == "SAN_CM") |> pull(gene)
AVN_CM_genes <- detable_CM |>filter(cluster == "AVN_CM") |> pull(gene)
Immat_CM_2_genes <- detable_CM |>filter(cluster == "Immat_CM_2") |> pull(gene)
CALN1high_FB_genes <- detable_FB |>filter(cluster == "CALN1high_FB") |> pull(gene)
Aut_Neu_2_genes <- detable_IN |>filter(cluster == "Aut_Neu_2") |> pull(gene)
Ven_EC_genes <- detable_EN |>filter(cluster == "Ven_EC") |> pull(gene)
Chrom_C_genes <- detable_IN |>filter(cluster == "Chrom_C") |> pull(gene)

#extract only the genes that are present in ligands_receptors
SAN_CM_genes_extract <- SAN_CM_genes[SAN_CM_genes %in% ligands_receptors == TRUE]
AVN_CM_genes_extract <- AVN_CM_genes[AVN_CM_genes %in% ligands_receptors == TRUE]
Immat_CM_2_genes_extract <- Immat_CM_2_genes[Immat_CM_2_genes %in% ligands_receptors == TRUE]
CALN1high_FB_genes_extract <- CALN1high_FB_genes[CALN1high_FB_genes %in% ligands_receptors == TRUE]
Aut_Neu_2_genes_extract <- Aut_Neu_2_genes[Aut_Neu_2_genes %in% ligands_receptors == TRUE]
Ven_EC_genes_extract <- Ven_EC_genes[Ven_EC_genes %in% ligands_receptors == TRUE]
Chrom_C_genes_extract <- Chrom_C_genes[Chrom_C_genes %in% ligands_receptors == TRUE]

#print ligands_receptors:
sort(SAN_CM_genes_extract)
sort(AVN_CM_genes_extract)
sort(Immat_CM_2_genes_extract)
sort(CALN1high_FB_genes_extract)
sort(Aut_Neu_2_genes_extract)
sort(Ven_EC_genes_extract)
sort(Chrom_C_genes_extract)

rm(detable_CM, detable_IN, detable_EN, detable_FB)
gc()
```


### ligands

```{r}
#extract Ligand for each cluster
#from SAN_CM_genes, extract only the genes that are present in ligands
SAN_CM_ligands_extract <- SAN_CM_genes[SAN_CM_genes %in% ligands == TRUE]

#from other clusters, extract only genes that are present in ligands
AVN_CM_ligands_extract <- AVN_CM_genes[AVN_CM_genes %in% ligands == TRUE]
Immat_CM_2_ligands_extract <- Immat_CM_2_genes[Immat_CM_2_genes %in% ligands == TRUE]
CALN1high_FB_ligands_extract <- CALN1high_FB_genes[CALN1high_FB_genes %in% ligands == TRUE]
Aut_Neu_2_ligands_extract <- Aut_Neu_2_genes[Aut_Neu_2_genes %in% ligands == TRUE]
Ven_EC_ligands_extract <- Ven_EC_genes[Ven_EC_genes %in% ligands == TRUE]
Chrom_C_ligands_extract <- Chrom_C_genes[Chrom_C_genes %in% ligands == TRUE]

#print ligands_receptors:
sort(SAN_CM_ligands_extract)

#print ligands_receptors:
sort(AVN_CM_ligands_extract)
sort(Immat_CM_2_ligands_extract)
sort(CALN1high_FB_ligands_extract)
sort(Aut_Neu_2_ligands_extract)
sort(Ven_EC_ligands_extract)
sort(Chrom_C_ligands_extract)
```


### receptors

```{r}
#extract Ligand for each cluster
#from SAN_CM_genes, extract only the genes that are present in receptors
SAN_CM_receptors_extract <- SAN_CM_genes[SAN_CM_genes %in% receptors == TRUE]

#from other clusters, extract only genes that are present in receptors
AVN_CM_receptors_extract <- AVN_CM_genes[AVN_CM_genes %in% receptors == TRUE]
Immat_CM_2_receptors_extract <- Immat_CM_2_genes[Immat_CM_2_genes %in% receptors == TRUE]
CALN1high_FB_receptors_extract <- CALN1high_FB_genes[CALN1high_FB_genes %in% receptors == TRUE]
Aut_Neu_2_receptors_extract <- Aut_Neu_2_genes[Aut_Neu_2_genes %in% receptors == TRUE]
Ven_EC_receptors_extract <- Ven_EC_genes[Ven_EC_genes %in% receptors == TRUE]
Chrom_C_receptors_extract <- Chrom_C_genes[Chrom_C_genes %in% receptors == TRUE]

#print ligands_receptors:
sort(SAN_CM_receptors_extract)

#print ligands_receptors:
sort(AVN_CM_receptors_extract)
sort(Immat_CM_2_receptors_extract)
sort(CALN1high_FB_receptors_extract)
sort(Aut_Neu_2_receptors_extract)
sort(Ven_EC_receptors_extract)
sort(Chrom_C_receptors_extract)
```


### Create list of vectors from the input_curated dataframe

```{r}
#create list of vector for LR pairs from the module
input_curated_list <- list()
for (row in seq_along(1:nrow(input_curated))) {
  print(row)
  i <- input_curated[row, ] |> pull(protein_name_a)
  j <- input_curated[row, ] |> pull(protein_name_b)
  vec <- c(i,j)
  input_curated_list <- append(input_curated_list, list(vec))
}
```


### Create a list of vectors for each potential L-R pairs

### from _genes_extract (no directionality kept)

#### SAN_CM & AVN_CM

```{r}
LR_SAN_AVN_list <- list()
for (i in SAN_CM_genes_extract){
  for (j in AVN_CM_genes_extract) {
    vec <- c(i,j)
    print(vec)
    LR_SAN_AVN_list <- append(LR_SAN_AVN_list, list(vec))
    
    vec2 <- c(j,i)
    print(vec2)
    LR_SAN_AVN_list <- append(LR_SAN_AVN_list, list(vec2))
  }
}
LR_SAN_AVN_list_common <- intersect(input_curated_list, LR_SAN_AVN_list)

rm(LR_SAN_AVN_list)
gc()
```


#### SAN_CM & Immat_CM_2

```{r}
LR_SAN_Immat_CM_list <- list()
for (i in SAN_CM_genes_extract){
  for (j in Immat_CM_2_genes_extract) {
    vec <- c(i,j)
    print(vec)
    LR_SAN_Immat_CM_list <- append(LR_SAN_Immat_CM_list, list(vec))
    
    vec2 <- c(j,i)
    print(vec2)
    LR_SAN_Immat_CM_list <- append(LR_SAN_Immat_CM_list, list(vec2))
  }
}
LR_SAN_Immat_CM_list_common <- intersect(input_curated_list, LR_SAN_Immat_CM_list)

rm(LR_SAN_Immat_CM_list)
gc()
```


#### SAN_CM & CALN1high_FB

```{r}
LR_SAN_CALN1high_FB_list <- list()
for (i in SAN_CM_genes_extract){
  for (j in CALN1high_FB_genes_extract) {
    vec <- c(i,j)
    print(vec)
    LR_SAN_CALN1high_FB_list <- append(LR_SAN_CALN1high_FB_list, list(vec))
    
    vec2 <- c(j,i)
    print(vec2)
    LR_SAN_CALN1high_FB_list <- append(LR_SAN_CALN1high_FB_list, list(vec2))
  }
}
LR_SAN_CALN1high_FB_list_common <- intersect(input_curated_list, LR_SAN_CALN1high_FB_list)

rm(LR_SAN_CALN1high_FB_list)
gc()
```


#### SAN_CM & Aut_Neu_2

```{r}
LR_SAN_Aut_Neu_2_list <- list()
for (i in SAN_CM_genes_extract){
  for (j in Aut_Neu_2_genes_extract) {
    vec <- c(i,j)
    print(vec)
    LR_SAN_Aut_Neu_2_list <- append(LR_SAN_Aut_Neu_2_list, list(vec))
    
    vec2 <- c(j,i)
    print(vec2)
    LR_SAN_Aut_Neu_2_list <- append(LR_SAN_Aut_Neu_2_list, list(vec2))

  }
}
LR_SAN_Aut_Neu_2_list_common <- intersect(input_curated_list, LR_SAN_Aut_Neu_2_list)

rm(LR_SAN_Aut_Neu_2_list)
gc()
```


#### SAN_CM & Ven_EC

```{r}
LR_SAN_Ven_EC_list <- list()
for (i in SAN_CM_genes_extract){
  for (j in Ven_EC_genes_extract) {
    vec <- c(i,j)
    print(vec)
    LR_SAN_Ven_EC_list <- append(LR_SAN_Ven_EC_list, list(vec))
    
    vec2 <- c(j,i)
    print(vec2)
    LR_SAN_Ven_EC_list <- append(LR_SAN_Ven_EC_list, list(vec2))

  }
}
LR_SAN_Ven_EC_list_common <- intersect(input_curated_list, LR_SAN_Ven_EC_list)

rm(LR_SAN_Ven_EC_list)
gc()
```


#### SAN_CM & Chrom_C

```{r}
LR_SAN_Chrom_C_list <- list()

for (i in SAN_CM_genes_extract){
  for (j in Chrom_C_genes_extract) {
    vec <- c(i,j)
    print(vec)
    LR_SAN_Chrom_C_list <- append(LR_SAN_Chrom_C_list, list(vec))

    vec2 <- c(j,i)
    print(vec2)
    LR_SAN_Chrom_C_list <- append(LR_SAN_Chrom_C_list, list(vec2))

  }
}
LR_SAN_Chrom_C_list_common <- intersect(input_curated_list, LR_SAN_Chrom_C_list)

rm(LR_SAN_Chrom_C_list)
gc()
```


### All catched LR pairs:

```{r}
LR_SAN_AVN_list_common
LR_SAN_Immat_CM_list_common
LR_SAN_CALN1high_FB_list_common
LR_SAN_Aut_Neu_2_list_common
LR_SAN_Ven_EC_list_common
LR_SAN_Chrom_C_list_common

all_LR_vector <- sort(unique(unlist(c(LR_SAN_AVN_list_common, LR_SAN_Immat_CM_list_common, LR_SAN_CALN1high_FB_list_common, LR_SAN_Aut_Neu_2_list_common, LR_SAN_Ven_EC_list_common, LR_SAN_Chrom_C_list_common))))
```


### from _ligands_extract & _receptors_extract (directionality kept)

#### SAN_CM & AVN_CM

```{r}
L_SAN_CM_R_AVN_CM_list <- list()
for (i in SAN_CM_ligands_extract){
  for (j in AVN_CM_receptors_extract) {
    vec <- c(i,j)
    print(vec)
    L_SAN_CM_R_AVN_CM_list <- append(L_SAN_CM_R_AVN_CM_list, list(vec))
  }
}
L_SAN_CM_R_AVN_CM_list_common <- intersect(input_curated_list, L_SAN_CM_R_AVN_CM_list)



L_AVN_CM_R_SAN_CM_list <- list()
for (i in  AVN_CM_ligands_extract){
  for (j in SAN_CM_receptors_extract) {
    vec <- c(i,j)
    print(vec)
    L_AVN_CM_R_SAN_CM_list <- append(L_AVN_CM_R_SAN_CM_list, list(vec))
  }
}
L_AVN_CM_R_SAN_CM_list_common <- intersect(input_curated_list, L_AVN_CM_R_SAN_CM_list)

rm(L_SAN_CM_R_AVN_CM_list, L_AVN_CM_R_SAN_CM_list)
gc()
```


#### SAN_CM & Immat_CM_2

```{r}
L_SAN_CM_R_Immat_CM_2_list <- list()
for (i in SAN_CM_ligands_extract){
  for (j in Immat_CM_2_receptors_extract) {
    vec <- c(i,j)
    print(vec)
    L_SAN_CM_R_Immat_CM_2_list <- append(L_SAN_CM_R_Immat_CM_2_list, list(vec))
  }
}
L_SAN_R_Immat_CM_list_common <- intersect(input_curated_list, L_SAN_CM_R_Immat_CM_2_list)



L_Immat_CM_2_R_SAN_CM_list <- list()
for (i in  Immat_CM_2_ligands_extract){
  for (j in SAN_CM_receptors_extract) {
    vec <- c(i,j)
    print(vec)
    L_Immat_CM_2_R_SAN_CM_list <- append(L_Immat_CM_2_R_SAN_CM_list, list(vec))
  }
}
L_Immat_CM_R_SAN_list_common <- intersect(input_curated_list, L_Immat_CM_2_R_SAN_CM_list)

rm(L_SAN_CM_R_Immat_CM_2_list, L_Immat_CM_2_R_SAN_CM_list)
gc()
```


#### SAN_CM & CALN1high_FB

```{r}
L_SAN_CM_R_CALN1high_FB_list <- list()
for (i in SAN_CM_ligands_extract){
  for (j in CALN1high_FB_receptors_extract) {
    vec <- c(i,j)
    print(vec)
    L_SAN_CM_R_CALN1high_FB_list <- append(L_SAN_CM_R_CALN1high_FB_list, list(vec))
  }
}
L_SAN_CM_R_CALN1high_FB_list_common <- intersect(input_curated_list, L_SAN_CM_R_CALN1high_FB_list)



L_CALN1high_FB_R_SAN_CM_list <- list()
for (i in  CALN1high_FB_ligands_extract){
  for (j in SAN_CM_receptors_extract) {
    vec <- c(i,j)
    print(vec)
    L_CALN1high_FB_R_SAN_CM_list <- append(L_CALN1high_FB_R_SAN_CM_list, list(vec))
  }
}
L_CALN1high_FB_R_SAN_CM_list_common <- intersect(input_curated_list, L_CALN1high_FB_R_SAN_CM_list)

rm(L_SAN_CM_R_CALN1high_FB_list, L_CALN1high_FB_R_SAN_CM_list)
gc()
```


#### SAN_CM & Aut_Neu_2

```{r}
L_SAN_CM_R_Aut_Neu_2_list <- list()
for (i in SAN_CM_ligands_extract){
  for (j in Aut_Neu_2_receptors_extract) {
    vec <- c(i,j)
    print(vec)
    L_SAN_CM_R_Aut_Neu_2_list <- append(L_SAN_CM_R_Aut_Neu_2_list, list(vec))
  }
}
L_SAN_CM_R_Aut_Neu_2_list_common <- intersect(input_curated_list, L_SAN_CM_R_Aut_Neu_2_list)



L_Aut_Neu_2_R_SAN_CM_list <- list()
for (i in  Aut_Neu_2_ligands_extract){
  for (j in SAN_CM_receptors_extract) {
    vec <- c(i,j)
    print(vec)
    L_Aut_Neu_2_R_SAN_CM_list <- append(L_Aut_Neu_2_R_SAN_CM_list, list(vec))
  }
}
L_Aut_Neu_2_R_SAN_CM_list_common <- intersect(input_curated_list, L_Aut_Neu_2_R_SAN_CM_list)

rm(L_SAN_CM_R_Aut_Neu_2_list, L_Aut_Neu_2_R_SAN_CM_list)
gc()
```


#### SAN_CM & Ven_EC

```{r}
L_SAN_CM_R_Ven_EC_list <- list()
for (i in SAN_CM_ligands_extract){
  for (j in Ven_EC_receptors_extract) {
    vec <- c(i,j)
    print(vec)
    L_SAN_CM_R_Ven_EC_list <- append(L_SAN_CM_R_Ven_EC_list, list(vec))
  }
}
L_SAN_CM_R_Ven_EC_list_common <- intersect(input_curated_list, L_SAN_CM_R_Ven_EC_list)



L_Ven_EC_R_SAN_CM_list <- list()
for (i in  Ven_EC_ligands_extract){
  for (j in SAN_CM_receptors_extract) {
    vec <- c(i,j)
    print(vec)
    L_Ven_EC_R_SAN_CM_list <- append(L_Ven_EC_R_SAN_CM_list, list(vec))
  }
}
L_Ven_EC_R_SAN_CM_list_common <- intersect(input_curated_list, L_Ven_EC_R_SAN_CM_list)

rm(L_SAN_CM_R_Ven_EC_list, L_Ven_EC_R_SAN_CM_list)
gc()
```


#### SAN_CM & Chrom_C

```{r}
L_SAN_CM_R_Chrom_C_list <- list()
for (i in SAN_CM_ligands_extract){
  for (j in Chrom_C_receptors_extract) {
    vec <- c(i,j)
    print(vec)
    L_SAN_CM_R_Chrom_C_list <- append(L_SAN_CM_R_Chrom_C_list, list(vec))
  }
}
L_SAN_CM_R_Chrom_C_list_common <- intersect(input_curated_list, L_SAN_CM_R_Chrom_C_list)



L_Chrom_C_R_SAN_CM_list <- list()
for (i in  Chrom_C_ligands_extract){
  for (j in SAN_CM_receptors_extract) {
    vec <- c(i,j)
    print(vec)
    L_Chrom_C_R_SAN_CM_list <- append(L_Chrom_C_R_SAN_CM_list, list(vec))
  }
}
L_Chrom_C_R_SAN_CM_list_common <- intersect(input_curated_list, L_Chrom_C_R_SAN_CM_list)

rm(L_SAN_CM_R_Chrom_C_list, L_Chrom_C_R_SAN_CM_list)
gc()
```


### All catched LR pairs:

```{r}
L_SAN_CM_R_AVN_CM_list_common
L_AVN_CM_R_SAN_CM_list_common
L_SAN_R_Immat_CM_list_common
L_Immat_CM_R_SAN_list_common
L_SAN_CM_R_CALN1high_FB_list_common
L_CALN1high_FB_R_SAN_CM_list_common
L_SAN_CM_R_Aut_Neu_2_list_common
L_Aut_Neu_2_R_SAN_CM_list_common
L_SAN_CM_R_Ven_EC_list_common
L_Ven_EC_R_SAN_CM_list_common
L_SAN_CM_R_Chrom_C_list_common
L_Chrom_C_R_SAN_CM_list_common


all_LR_vector <- sort(unique(unlist(c(L_SAN_CM_R_AVN_CM_list_common,
L_AVN_CM_R_SAN_CM_list_common,
L_SAN_R_Immat_CM_list_common,
L_Immat_CM_R_SAN_list_common,
L_SAN_CM_R_CALN1high_FB_list_common,
L_CALN1high_FB_R_SAN_CM_list_common,
L_SAN_CM_R_Aut_Neu_2_list_common,
L_Aut_Neu_2_R_SAN_CM_list_common,
L_SAN_CM_R_Ven_EC_list_common,
L_Ven_EC_R_SAN_CM_list_common,
L_SAN_CM_R_Chrom_C_list_common,
L_Chrom_C_R_SAN_CM_list_common))))
```


### Calculate LR scores

SAN_CM = CM_19-2
AVN_CM = CM_19-1
Immat_CM_2 = CM_7
CALN1high_FB = FB_17
Aut_Neu_2 = IN_13
Ven_EC = EN_16
Chrom_C = IN_1

```{r}
avg_exp <- AverageExpression(object = data_kept_cells, group.by = c('annot_dl'))$RNA
avg_exp <- avg_exp[rownames(avg_exp) %in% all_LR_vector, 
                   colnames(avg_exp) %in% c("CM_19-2", "CM_19-1", "CM_7", "FB_17", "IN_13", "EN_16", "IN_1")]
```


## Function L-->R score

```{r}
source("/home/rstudio/project/code/LR.R")
```


# No directionality 
## SAN to Others

### SAN(L) --> AVN(R)

```{r}
SAN_AVN <- LR_table(avg_exp,
                       Ligand_source = "CM_19-2",
                       Receptor_source = "CM_19-1",
                       list_LR_pairs = LR_SAN_AVN_list_common)
```

### SAN(L) --> Immat_CM_2(R)

```{r}
SAN_Immat_CM_2 <- LR_table(avg_exp,
                       Ligand_source = "CM_19-2",
                       Receptor_source = "CM_7",
                       list_LR_pairs = LR_SAN_Immat_CM_list_common)
```


### SAN(L) --> CALN1high_FB(R)

```{r}
SAN_CALN1high_FB <- LR_table(avg_exp,
                       Ligand_source = "CM_19-2",
                       Receptor_source = "FB_17",
                       list_LR_pairs = LR_SAN_CALN1high_FB_list_common)
```


### SAN(L) --> Aut_Neu_2(R)

```{r}
SAN_Aut_Neu_2 <- LR_table(avg_exp,
                       Ligand_source = "CM_19-2",
                       Receptor_source = "IN_13",
                       list_LR_pairs = LR_SAN_Aut_Neu_2_list_common)
#SAN_Aut_Neu_2 <- SAN_Aut_Neu_2 |> filter(Score != 0)
```


### SAN(L) --> Ven_EC(R)

```{r}
SAN_Ven_EC <- LR_table(avg_exp,
                       Ligand_source = "CM_19-2",
                       Receptor_source = "EN_16",
                       list_LR_pairs = LR_SAN_Ven_EC_list_common)
```


### SAN(L) --> Chrom_C(R)

```{r}
SAN_Chrom_C <- LR_table(avg_exp,
                       Ligand_source = "CM_19-2",
                       Receptor_source = "IN_1",
                       list_LR_pairs = LR_SAN_Chrom_C_list_common)
#SAN_Chrom_C <- SAN_Chrom_C |> filter(Score != 0)
```


### Merge data frames SAN (L) --> others (R)

```{r}
merged_dataframe <- bind_rows(SAN_AVN, SAN_CALN1high_FB, SAN_Aut_Neu_2, SAN_Ven_EC, SAN_Chrom_C) #remove SAN_Immat_CM_2 otherwise the merge does not work.
merged_dataframe$Score <- as.numeric(merged_dataframe$Score)
```


### Plot Heatmap

```{r}
pal <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "RdPu"))(25)

#plot
p <- ggplot(merged_dataframe, aes(merged_dataframe[,1], merged_dataframe[,2], fill = merged_dataframe[,3])) + 
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  labs(fill = "LR score",
       x = "Source cell types",
       y = "Ligand --> Receptor")

p


pdf(file = paste0(myfolder, "cellcellcommunication/Heatmap_SAN_to_others_v2.pdf"),
     width = 4, height = 25)
p
dev.off()
```


## Others to SAN

### AVN(L) --> SAN(R)

```{r}
AVN_SAN <- LR_table(avg_exp,
                       Ligand_source = "CM_19-1",
                       Receptor_source = "CM_19-2",
                       list_LR_pairs = LR_SAN_AVN_list_common)
```

### Immat_CM_2(L) --> SAN(R)

```{r}
Immat_CM_2_SAN <- LR_table(avg_exp,
                       Ligand_source = "CM_7",
                       Receptor_source = "CM_19-2",
                       list_LR_pairs = LR_SAN_Immat_CM_list_common)
```


### CALN1high_FB(L) --> SAN(R)

```{r}
CALN1high_FB_SAN <- LR_table(avg_exp,
                       Ligand_source = "FB_17",
                       Receptor_source = "CM_19-2",
                       list_LR_pairs = LR_SAN_CALN1high_FB_list_common)
```


### Aut_Neu_2(L) --> SAN(R)

```{r}
Aut_Neu_2_SAN <- LR_table(avg_exp,
                       Ligand_source = "IN_13",
                       Receptor_source = "CM_19-2",
                       list_LR_pairs = LR_SAN_Aut_Neu_2_list_common)
#Aut_Neu_2_SAN <- Aut_Neu_2_SAN |> filter(Score != 0)
```


### Ven_EC(L) --> SAN(R)

```{r}
Ven_EC_SAN <- LR_table(avg_exp,
                       Ligand_source = "EN_16",
                       Receptor_source = "CM_19-2",
                       list_LR_pairs = LR_SAN_Ven_EC_list_common)
```


### Chrom_C(L) --> SAN(R)

```{r}
Chrom_C_SAN <- LR_table(avg_exp,
                       Ligand_source = "IN_1",
                       Receptor_source = "CM_19-2",
                       list_LR_pairs = LR_SAN_Chrom_C_list_common)
#Chrom_C_SAN <- Chrom_C_SAN |> filter(Score != 0)
```


### Merge data frames Others (L) --> SAN (R)

```{r}
merged_dataframe <- bind_rows(AVN_SAN, CALN1high_FB_SAN, Aut_Neu_2_SAN, Ven_EC_SAN, Chrom_C_SAN)
merged_dataframe$Score <- as.numeric(merged_dataframe$Score)
```


### Plot Heatmap

```{r}
pal <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "RdPu"))(25)

#plot
p <- ggplot(merged_dataframe, aes(merged_dataframe[,1], merged_dataframe[,2], fill = merged_dataframe[,3])) + 
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  labs(fill = "LR score",
       x = "Source cell types",
       y = "Ligand --> Receptor")

p


pdf(file = paste0(myfolder, "cellcellcommunication/Heatmap_others_to_SAN_v2.pdf"),
     width = 4, height = 25)
p
dev.off()
```


### Merge data frames SAN_CM <-> Aut_Neu_2 // PTPRS-NTRK3 + TENM2-ADGRL1

```{r}
merged_dataframe <- bind_rows(Aut_Neu_2_SAN, SAN_Aut_Neu_2)
merged_dataframe$Score <- as.numeric(merged_dataframe$Score)

merged_dataframe <- merged_dataframe |> 
  filter(`Ligand -> Receptor` %in% c("PTPRS-NTRK3", "TENM2-ADGRL1"))
```


### Plot Heatmap

```{r}
pal <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "RdPu"))(25)

#plot
p <- ggplot(merged_dataframe, aes(merged_dataframe[,1], merged_dataframe[,2], fill = merged_dataframe[,3])) + 
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  labs(fill = "LR score",
       x = "Source cell types",
       y = "Ligand --> Receptor")

p

pdf(file = paste0(myfolder, "cellcellcommunication/Heatmap_Aut_Neu_2_and_SAN.pdf"),
     width = 4, height = 4)
p
dev.off()
```


# With directionality 

## SAN to Others

### SAN(L) --> AVN(R)

```{r}
SAN_AVN <- LR_table(avg_exp,
                       Ligand_source = "CM_19-2",
                       Receptor_source = "CM_19-1",
                       list_LR_pairs = L_SAN_CM_R_AVN_CM_list_common)
```

### SAN(L) --> Immat_CM_2(R)

```{r}
SAN_Immat_CM_2 <- LR_table(avg_exp,
                       Ligand_source = "CM_19-2",
                       Receptor_source = "CM_7",
                       list_LR_pairs = L_SAN_R_Immat_CM_list_common)
```


### SAN(L) --> CALN1high_FB(R)

```{r}
SAN_CALN1high_FB <- LR_table(avg_exp,
                       Ligand_source = "CM_19-2",
                       Receptor_source = "FB_17",
                       list_LR_pairs = L_SAN_CM_R_CALN1high_FB_list_common)
```


### SAN(L) --> Aut_Neu_2(R)

```{r}
SAN_Aut_Neu_2 <- LR_table(avg_exp,
                       Ligand_source = "CM_19-2",
                       Receptor_source = "IN_13",
                       list_LR_pairs = L_SAN_CM_R_Aut_Neu_2_list_common)
#SAN_Aut_Neu_2 <- SAN_Aut_Neu_2 |> filter(Score != 0)
```


### SAN(L) --> Ven_EC(R)

```{r}
SAN_Ven_EC <- LR_table(avg_exp,
                       Ligand_source = "CM_19-2",
                       Receptor_source = "EN_16",
                       list_LR_pairs = L_SAN_CM_R_Ven_EC_list_common)
```


### SAN(L) --> Chrom_C(R)

```{r}
SAN_Chrom_C <- LR_table(avg_exp,
                       Ligand_source = "CM_19-2",
                       Receptor_source = "IN_1",
                       list_LR_pairs = L_SAN_CM_R_Chrom_C_list_common)
#SAN_Chrom_C <- SAN_Chrom_C |> filter(Score != 0)
```


### Merge data frames SAN (L) --> others (R)

```{r}
merged_dataframe <- bind_rows(SAN_AVN, SAN_CALN1high_FB, SAN_Aut_Neu_2, SAN_Ven_EC) #remove SAN_Chrom_C, SAN_Immat_CM_2 otherwise the merge does not work.
merged_dataframe$Score <- as.numeric(merged_dataframe$Score)
```


### Plot Heatmap SAN to others (Fig. 7C)

```{r}
pal <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "RdPu"))(25)

#plot
p <- ggplot(merged_dataframe, aes(merged_dataframe[,1], merged_dataframe[,2], fill = merged_dataframe[,3])) + 
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  labs(fill = "LR score",
       x = "Source cell types",
       y = "Ligand --> Receptor")

p


pdf(file = paste0(myfolder, "cellcellcommunication/Heatmap_SAN_to_others_v3.pdf"),
     width = 4, height = 11)
p
dev.off()
```


## Others to SAN

### AVN(L) --> SAN(R)

```{r}
AVN_SAN <- LR_table(avg_exp,
                       Ligand_source = "CM_19-1",
                       Receptor_source = "CM_19-2",
                       list_LR_pairs = L_AVN_CM_R_SAN_CM_list_common)
```

### Immat_CM_2(L) --> SAN(R)

```{r}
Immat_CM_2_SAN <- LR_table(avg_exp,
                       Ligand_source = "CM_7",
                       Receptor_source = "CM_19-2",
                       list_LR_pairs = L_Immat_CM_R_SAN_list_common)
```


### CALN1high_FB(L) --> SAN(R)

```{r}
CALN1high_FB_SAN <- LR_table(avg_exp,
                       Ligand_source = "FB_17",
                       Receptor_source = "CM_19-2",
                       list_LR_pairs = L_CALN1high_FB_R_SAN_CM_list_common)
```


### Aut_Neu_2(L) --> SAN(R)

```{r}
Aut_Neu_2_SAN <- LR_table(avg_exp,
                       Ligand_source = "IN_13",
                       Receptor_source = "CM_19-2",
                       list_LR_pairs = L_Aut_Neu_2_R_SAN_CM_list_common)
#Aut_Neu_2_SAN <- Aut_Neu_2_SAN |> filter(Score != 0)
```


### Ven_EC(L) --> SAN(R)

```{r}
Ven_EC_SAN <- LR_table(avg_exp,
                       Ligand_source = "EN_16",
                       Receptor_source = "CM_19-2",
                       list_LR_pairs = L_Ven_EC_R_SAN_CM_list_common)
```


### Chrom_C(L) --> SAN(R)

```{r}
Chrom_C_SAN <- LR_table(avg_exp,
                       Ligand_source = "IN_1",
                       Receptor_source = "CM_19-2",
                       list_LR_pairs = L_Chrom_C_R_SAN_CM_list_common)
#Chrom_C_SAN <- Chrom_C_SAN |> filter(Score != 0)
```


### Merge data frames Others (L) --> SAN (R)

```{r}
merged_dataframe <- bind_rows(AVN_SAN, Aut_Neu_2_SAN, Ven_EC_SAN, Chrom_C_SAN, CALN1high_FB_SAN) #Immat_CM_2_SAN because empty
merged_dataframe$Score <- as.numeric(merged_dataframe$Score)
```


### Plot Heatmap others to SAN (Fig. 7C)

```{r}
pal <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "RdPu"))(25)

#plot
p <- ggplot(merged_dataframe, aes(merged_dataframe[,1], merged_dataframe[,2], fill = merged_dataframe[,3])) + 
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  labs(fill = "LR score",
       x = "Source cell types",
       y = "Ligand --> Receptor")

p


pdf(file = paste0(myfolder, "cellcellcommunication/Heatmap_others_to_SAN_v3.pdf"),
     width = 4, height = 8)
p
dev.off()
```



### Merge data frames IN_1+IN_13 (L) --> SAN (R) // CHAT-CHRM2 and TH-ADRB1 (Supp Fig. 4D)

```{r}
merged_dataframe <- bind_rows(Aut_Neu_2_SAN, Chrom_C_SAN)
merged_dataframe$Score <- as.numeric(merged_dataframe$Score)

merged_dataframe <- merged_dataframe |> 
  filter(`Ligand -> Receptor` %in% c("TH-ADRB1", "CHAT-CHRM2"))
```


### Plot Heatmap

```{r}
pal <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "RdPu"))(25)

#plot
p <- ggplot(merged_dataframe, aes(merged_dataframe[,1], merged_dataframe[,2], fill = merged_dataframe[,3])) + 
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  labs(fill = "LR score",
       x = "Source cell types",
       y = "Ligand --> Receptor")

p

pdf(file = paste0(myfolder, "cellcellcommunication/Heatmap_IN_to_SAN.pdf"),
     width = 4, height = 4)
p
dev.off()
```


###SAN_CM to IN_13

### Plot Heatmap IN_13 to SAN_CM

```{r}
Aut_Neu_2_SAN$Score <- as.numeric(Aut_Neu_2_SAN$Score)

pal <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "RdPu"))(25)

#plot
p <- ggplot(Aut_Neu_2_SAN, aes(Aut_Neu_2_SAN[,1], Aut_Neu_2_SAN[,2], fill = Aut_Neu_2_SAN[,3])) + 
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  labs(fill = "LR score",
       x = "Source cell types",
       y = "Ligand --> Receptor")

p

pdf(file = paste0(myfolder, "cellcellcommunication/Heatmap_Aut_Neu_2_SAN.pdf"),
     width = 3, height = 25)
p
dev.off()
```


### Plot Heatmap SAN_CM to IN_13

```{r}
SAN_Aut_Neu_2$Score <- as.numeric(SAN_Aut_Neu_2$Score)

pal <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "RdPu"))(25)

#plot
p <- ggplot(SAN_Aut_Neu_2, aes(SAN_Aut_Neu_2[,1], SAN_Aut_Neu_2[,2], fill = SAN_Aut_Neu_2[,3])) + 
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  labs(fill = "LR score",
       x = "Source cell types",
       y = "Ligand --> Receptor")

p

pdf(file = paste0(myfolder, "cellcellcommunication/Heatmap_SAN_Aut_Neu_2.pdf"),
     width = 3, height = 25)
p
dev.off()
```


### Plot Heatmap IN_13 to SAN_CM & SAN_CM to IN_13 (Supp Fig. 4G)

```{r}
merged_dataframe <- bind_rows(Aut_Neu_2_SAN, SAN_Aut_Neu_2)

pal <- colorRampPalette(RColorBrewer::brewer.pal(n = 8, name = "RdPu"))(25)

#plot
p <- ggplot(merged_dataframe, aes(merged_dataframe[,1], merged_dataframe[,2], fill = merged_dataframe[,3])) + 
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90)) +
  labs(fill = "LR score",
       x = "Source cell types",
       y = "Ligand --> Receptor")

p

pdf(file = paste0(myfolder, "cellcellcommunication/Heatmap_Aut_Neu_2_SAN_AND_SAN_Aut_Neu_2.pdf"),
     width = 3, height = 25)
p
dev.off()
```

```{r}

```







