---
title: "XX_HDCA_heart_innervation_trajectory_analysis"
author: "Raphaël Mauron"
output: html_document
editor_options: 
  chunk_output_type: console
---

# HDCA heart innervation trajectory analysis

## Settings

### Set library path

```{r setup, include=FALSE}
.libPaths("/home/rstudio/project/renv/library/R-4.3/aarch64-unknown-linux-gnu/")
knitr::opts_chunk$set(echo = TRUE)
```


### Set environment variables

```{r}
# Set the directory 
myfolder <- "/home/rstudio/project/output/HDCA_heart_sc_analysis_docker/"
```


### Load Libraries

```{r, message = FALSE}
library(Seurat)
library(Matrix)
library(dplyr)
library(harmony)
library(ggplot2)
library(patchwork)
library(openxlsx)
library(plotly)
library(slingshot)
library(RColorBrewer)
```


### Load data

```{r}
IN_subset <- readRDS(paste0(myfolder, "rds_objects/innervation_selection.rds"))

IN_subset$clusters_subset <- factor(IN_subset$clusters_subset, levels = c("1", "3", "4", "7", "8", "11", "12", "13", "17", "19")) 
table(IN_subset$clusters_subset)
```

```{r}
DimPlot(IN_subset, reduction = "umap_subset_sub", group.by = "clusters_subset", label = T)
```

```{r}
FeaturePlot(IN_subset, reduction = "umap_subset_sub", features = "FOSB", order = T)
```


## Unsuppervised

```{r}
# Convert Seurat object into a SingleCellExperiment object
IN_subset_sce <- as.SingleCellExperiment(x= IN_subset)

# Create Slingshot object (getLineages & getCurves)
sce <- slingshot(IN_subset_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB')

summary(sce$slingPseudotime_1)
summary(sce$slingPseudotime_2)


plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(sce), lwd=2, col = 'black')

pdf(file = paste0(myfolder, "trajectory/slingshot/IN_all_clusters_unsupervised.pdf"),  width = 10, height = 10)
plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(sce), lwd=2, type = 'lineages', col = 'black')
dev.off()

sce@colData@listData[["slingshot"]]@metadata[["lineages"]]
```


## Add start point

```{r}
lin1 <- getLineages(IN_subset_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB', start.clus = '4')
lin1
```


```{r}
plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(lin1), lwd = 2, col = 'black')

lin1@colData@listData[["slingshot"]]@metadata[["lineages"]]
```


## Add start and end points

```{r}
lin2 <- getLineages(IN_subset_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB', start.clus = c('7'), end.clus = c('13', '1'))
lin2
```

```{r}
pdf(file = paste0(myfolder, "trajectory/slingshot/IN_all_clusters_supervised.pdf"),  width = 10, height = 10)
plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(lin2), lwd = 2, col = 'black')
dev.off()

lin2@colData@listData[["slingshot"]]@metadata[["lineages"]]
```


## Clusters no 12

```{r}
IN_subset_no_12 <- IN_subset[, !IN_subset$clusters_subset == c(12)]
IN_subset_no_12$clusters_subset <- factor(IN_subset_no_12$clusters_subset, levels = c("1", "3", "4", "7", "8", "11",  "13", "17", "19")) 
table(IN_subset_no_12$clusters_subset)
```


### Unsuppervised

```{r}
# Convert Seurat object into a SingleCellExperiment object
IN_1_sce <- as.SingleCellExperiment(x= IN_subset_no_12)

# Create Slingshot object (getLineages & getCurves)
sce <- slingshot(IN_1_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB')

summary(sce$slingPseudotime_1)
summary(sce$slingPseudotime_2)


plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(sce), lwd=2, col = 'black')


plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(sce), lwd=2, type = 'lineages', col = 'black')
```


### Add start and end points

```{r}
lin2 <- getLineages(IN_1_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB', start.clus = '17', end.clus = c('13', '1'))
lin2

crv1 <- getCurves(lin2)
crv1
```

```{r}
pdf(file = paste0(myfolder, "trajectory/slingshot/IN_no_12_supervised.pdf"),  width = 10, height = 10)
plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(lin2), lwd = 3, col = 'black')
dev.off()

lin2@colData@listData[["slingshot"]]@metadata[["lineages"]]
```



## Clusters "1", "3", "13", "17"

```{r}
IN_1 <- IN_subset[, !IN_subset$clusters_subset == c(4, 7, 8, 11, 12, 19)]
IN_1$clusters_subset <- factor(IN_1$clusters_subset, levels = c("1", "3", "13", "17")) 
table(IN_1$clusters_subset)
```


### Unsuppervised

```{r}
# Convert Seurat object into a SingleCellExperiment object
IN_1_sce <- as.SingleCellExperiment(x= IN_1)

# Create Slingshot object (getLineages & getCurves)
sce <- slingshot(IN_1_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB')

summary(sce$slingPseudotime_1)
summary(sce$slingPseudotime_2)


plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(sce), lwd=2, col = 'black')

pdf(file = paste0(myfolder, "trajectory/slingshot/IN_1_3_13_17_unsupervised.pdf"),  width = 10, height = 10)
plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(sce), lwd=2, type = 'lineages', col = 'black')
dev.off()

sce@colData@listData[["slingshot"]]@metadata[["lineages"]]
```


### Add start and end points

```{r}
lin2 <- getLineages(IN_1_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB', start.clus = '17', end.clus = c('1'))
#lin2 <- getLineages(IN_1_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB', start.clus = '17')
lin2

crv1 <- getCurves(lin2)
crv1
```

```{r}
pdf(file = paste0(myfolder, "trajectory/slingshot/IN_1_3_13_17_supervised.pdf"),  width = 10, height = 10)
plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(crv1), lwd = 2, col = 'black')
dev.off()

lin2@colData@listData[["slingshot"]]@metadata[["lineages"]]
```



## Clusters "3", "13", "17"

```{r}
IN_1_1 <- IN_subset[, !IN_subset$clusters_subset == c(1, 4, 7, 8, 11, 12, 19)]
IN_1_1$clusters_subset <- factor(IN_1_1$clusters_subset, levels = c("3", "13", "17")) 
table(IN_1_1$clusters_subset)
```


### Unsuppervised

```{r}
# Convert Seurat object into a SingleCellExperiment object
IN_1_1_sce <- as.SingleCellExperiment(x= IN_1_1)

# Create Slingshot object (getLineages & getCurves)
sce <- slingshot(IN_1_1_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB')

summary(sce$slingPseudotime_1)
summary(sce$slingPseudotime_2)


plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(sce), lwd=2, col = 'black')

pdf(file = paste0(myfolder, "trajectory/slingshot/IN_3_13_17_unsupervised.pdf"),  width = 10, height = 10)
plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(sce), lwd=2, type = 'lineages', col = 'black')
dev.off()

sce@colData@listData[["slingshot"]]@metadata[["lineages"]]
```


### Add start and end points

```{r}
#lin2 <- getLineages(IN_1_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB', start.clus = '17', end.clus = '13')
lin2 <- getLineages(IN_1_1_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB', start.clus = '17')
lin2

crv1 <- getCurves(lin2)
crv1
```

```{r}
pdf(file = paste0(myfolder, "trajectory/slingshot/IN_3_13_17_supervised.pdf"),  width = 10, height = 10)
plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(crv1), lwd = 2, col = 'black', show.constraints = TRUE)
dev.off()

lin2@colData@listData[["slingshot"]]@metadata[["lineages"]]
```


## Clusters "4", "7", "8", "11", "19"


```{r}
IN_2 <- IN_subset[, !IN_subset$clusters_subset == c(1, 3, 12, 13, 17)]
IN_2$clusters_subset <- factor(IN_2$clusters_subset, levels = c("4", "7", "8", "11", "19")) 
table(IN_2$clusters_subset)
```


### Unsuppervised

```{r}
# Convert Seurat object into a SingleCellExperiment object
IN_2_sce <- as.SingleCellExperiment(x= IN_2)

# Create Slingshot object (getLineages & getCurves)
sce <- slingshot(IN_2_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB')

summary(sce$slingPseudotime_1)
summary(sce$slingPseudotime_2)


plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(sce), lwd=2, col = 'black')

pdf(file = paste0(myfolder, "trajectory/slingshot/IN_4_7_8_11_19_unsupervised.pdf"),  width = 10, height = 10)
plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(sce), lwd=2, type = 'lineages', col = 'black')
dev.off()

sce@colData@listData[["slingshot"]]@metadata[["lineages"]]
```


### Add start and end points

```{r}
lin2 <- getLineages(IN_2_sce, clusterLabels = 'clusters_subset', reducedDim = 'UMAP_SUBSET_SUB', start.clus = '7', end.clus = c('19'))
lin2

crv1 <- getCurves(lin2)
crv1
```

```{r}
pdf(file = paste0(myfolder, "trajectory/slingshot/IN_4_7_8_11_19_supervised.pdf"),  width = 10, height = 10)
plot(reducedDims(sce)$UMAP_SUBSET_SUB, col = brewer.pal(n = 10, name = "Paired")[sce$clusters_subset], pch=16, asp = 1)
lines(SlingshotDataSet(crv1), lwd = 2, col = 'black')
dev.off()

lin2@colData@listData[["slingshot"]]@metadata[["lineages"]]
```






