---
title: "12_HDCA_heart_export_data_sc"
author: "Raphaël Mauron"
output: html_document
editor_options: 
  chunk_output_type: console
---

# HDCA heart export data for viewer

NOTE: The path to load the data and the path to save the figures should be changed at your convenience.

Environment: **Docker**


## Settings

### Set library path

```{r setup, include=FALSE}
.libPaths("/home/rstudio/project/renv/library/R-4.3/aarch64-unknown-linux-gnu/")
knitr::opts_chunk$set(echo = TRUE)
```


### Set environment variables

```{r}
# Set the directory 
myfolder <- "/home/rstudio/project/output/HDCA_heart_sc_analysis_docker/"
```


### Load Libraries

```{r, message = FALSE}
library(Seurat)
library(Matrix)
library(future)
library(RcppHNSW)
library(igraph)
library(future.apply)
library(niceRplots)
library(DoubletFinder)
library(ggplot2)
library(ROCR)
library(KernSmooth)
library(dplyr)
library(patchwork)
library(harmony)
library(gprofiler2)
library(plotly)
library(gtools)
library(openxlsx)
library(dplyr)
library(tibble)
```


### Set seed

```{r}
set.seed(1)
```


### Export coarse-grained processed expression data

```{r}
data <- readRDS(paste0(myfolder, "rds_objects/HDCA_heart_sc_full_extended.rds"))
mat <- t(GetAssayData(object = data, assay = "RNA", slot = "data"))

rm(data)
gc()

write.csv(mat, file = paste0(myfolder, "export_data/export_sc/", "coarse_grained_gene_expr", ".csv"))
```


### Export coarse-grained processed expression raw

```{r}
data <- readRDS(paste0(myfolder, "rds_objects/HDCA_heart_sc_full_extended.rds"))
mat <-  t(GetAssayData(object = data, assay = "RNA", slot = "counts"))

rm(data)
gc()
#write.csv(mat, file = paste0(myfolder, "export_data/export_sc/", "coarse_grained_gene_expr_raw", ".csv"))

```

```{r}
write_chunked_csv <- function(mat, file, chunk_size = 1000) {
  for (i in seq(1, nrow(mat), by = chunk_size)) {
    chunk <- mat[i:min(i + chunk_size - 1, nrow(mat)), ]
    write.table(chunk, file = file, sep = ",", col.names = (i == 1), 
                append = (i != 1), row.names = TRUE, quote = FALSE)
  }
}

write_chunked_csv(mat, paste0(myfolder, "export_data/export_sc/", "coarse_grained_gene_expr_raw", ".csv"))
```



### Export fine-grained processed expression data

```{r}
#CM
CM <- readRDS(file = paste0(myfolder, "rds_objects/cardiomyocytes02.rds"))
mat <- t(GetAssayData(object = CM, assay = "RNA", slot = "data"))

rm(CM)
gc()

write.csv(mat, file = paste0(myfolder, "export_data/export_sc/", "CM_gene_expr", ".csv"))
```

```{r}
#EN
EN <- readRDS(file = paste0(myfolder, "rds_objects/endothelial.rds"))
mat <- t(GetAssayData(object = EN, assay = "RNA", slot = "data"))

rm(EN)
gc()

write.csv(mat, file = paste0(myfolder, "export_data/export_sc/", "EN_gene_expr", ".csv"))
```

```{r}
#FB
FB <- readRDS(paste0(myfolder, "rds_objects/fibroblasts.rds"))
mat <- t(GetAssayData(object = FB, assay = "RNA", slot = "data"))

rm(FB)
gc()

write.csv(mat, file = paste0(myfolder, "export_data/export_sc/", "FB_gene_expr", ".csv"))
```

```{r}
#IN
IN <- readRDS(paste0(myfolder, "rds_objects/innervation.rds"))
mat <- t(GetAssayData(object = IN, assay = "RNA", slot = "data"))

rm(IN)
gc()

write.csv(mat, file = paste0(myfolder, "export_data/export_sc/", "IN_gene_expr", ".csv"))
```


